### 2014-02-28 draft.tex
Article - 1er version

% !TEX TS-program = pdflatex
% !TEX encoding = UTF-8 Unicode

% This is a simple template for a LaTeX document using the "article" class.
% See "book", "report", "letter" for other types of document.

\documentclass{acm_proc_article-sp} % use larger type; default would be 10pt
%\usepackage[margin={1.5cm,3cm}]{geometry}
%\setlength{\columnsep}{1cm}
\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)

%%% Examples of Article customizations
% These packages are optional, depending whether you want the features they provide.
% See the LaTeX Companion or other references for full information.

%%% PAGE DIMENSIONS
%\usepackage{geometry} % to change the page dimensions
%\geometry{a4paper} % or letterpaper (US) or a5paper or....
% \geometry{margin=2in} % for example, change the margins to 2 inches all round
% \geometry{landscape} % set up the page for landscape
%   read geometry.pdf for detailed page layout information
\usepackage[hidelinks]{hyperref}
\usepackage[autostyle,french=guillemets,babel]{csquotes}
\usepackage{graphicx} % support the \includegraphics command and options
\usepackage[backend=biber]{biblatex}
\addbibresource{biblio.bib}
% \usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent

%%% PACKAGES
\usepackage{booktabs} % for much better looking tables
\usepackage{array} %
% for better arrays (eg matrices) in maths
\usepackage{paralist} % very flexible & customisable lists (eg. enumerate/itemize, etc.)
\usepackage{verbatim} % adds environment for commenting out blocks of text & for better verbatim
\usepackage{subfig} % make it possible to include more than one captioned figure/table in a single float
% These packages are all incorporated in the memoir class to one degree or another...

%%% HEADERS & FOOTERS
\usepackage{fancyhdr} % This should be set AFTER setting up the page geometry
\pagestyle{fancy} % options: empty , plain , fancy
\renewcommand{\headrulewidth}{0pt} % customise the layout...
\lhead{}\chead{}\rhead{}
\lfoot{}\cfoot{\thepage}\rfoot{}

%%% SECTION TITLE APPEARANCE
%\usepackage{sectsty}
%\allsectionsfont{\sffamily\mdseries\upshape} % (See the fntguide.pdf for font help)
%% (This matches ConTeXt defaults)

%%% ToC (table of contents) APPEARANCE
%\usepackage[nottoc,notlof,notlot]{tocbibind} % Put the bibliography in the ToC
%\usepackage[titles,subfigure]{tocloft} % Alter the style of the Table of Contents
%\renewcommand{\cftsecfont}{\rmfamily\mdseries\upshape}
%\renewcommand{\cftsecpagefont}{\rmfamily\mdseries\upshape} % No bold!

%%% END Article customizations

%%% The "real" document content comes below...

\title{A Sample {\ttlit ACM} SIG Proceedings Paper in LaTeX
Format\titlenote{(Does NOT produce the permission block, copyright information nor page numbering). For use with ACM\_PROC\_ARTICLE-SP.CLS. Supported by ACM.}}
\subtitle{[Extended Abstract]
\titlenote{A full version of this paper is available as
\textit{Author's Guide to Preparing ACM SIG Proceedings Using
\LaTeX$2_\epsilon$\ and BibTeX} at
\texttt{www.acm.org/eaddress.htm}}}
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\numberofauthors{1}
\alignauthor
Ben Trovato\titlenote{Dr.~Trovato insisted his name be first.}\\
       \affaddr{Institute for Clarity in Documentation}\\
       \affaddr{1932 Wallamaloo Lane}\\
       \affaddr{Wallamaloo, New Zealand}\\
       \email{trovato@corporation.com}
% 2nd. author
%\date{} % Activate to display a given date or no date (if empty),
         % otherwise the current date is printed 
}
\begin{document}
\maketitle

\abstract{
%Today, hypervisors have to find a complex trade-off between functionalities provided and security assurance. Those requirements are contrasting by design:
%adding new features will generally increases the LoC and consequently the attack surface. On the contrary, those code additions are required to properly fulfill market requests. 
%
%Nested virtualization is a technique capable to diversify hypervisor features. Exploiting an extra layer of virtualization, we will be able to decouple different requirements, avoiding the need to find delicate trade-off.
%
%In this paper, we will analyze the state of nested virtualization in modern hypervisor and evaulating the impact of utilization of microhypervisor to enforce a stronger security level.

}

\section{Introduction}

Cloud computing platform are a solid reality nowadays. Modern cloud platforms are able to provide elastic use of resources, optimizing their allocation in order to meet the user needings.
However, currently deployed cloud platforms present still several limitation.

First, it is still difficult to deliver cloud applications which are able to transparently leverage over several platform resources. The concept of SuperCloud \cite{art:blan, xclo:blank} addresses this kind of problems. Multi-platform coordination primarily for different standard and several format lock-in. Unfortunately, those lock-in are not only due to market strategical decisions but they also reflects the virtualization technology heterogeneity. For example, each cloud provider could use a different storage system or a different format to store VMs images. Different hypervisors have nowadays a low degree of compatibility between them and this generally prevents leveraging on some popular features like `` VM live migration'' between different platform. Moreover, this incompatibility could also be enlarged if we consider that several platforms adopts one or more different virtualization techniques. In addition, this situation induce a tight dependency between the customer and the cloud provider and hardens the possibility to guarantee the availability of service in presence of a certain provider or to put in place some geographical optimization due to cost or traffic origin. The result was that some data server failures in the past has been catastrophic, causing a prolonged outage of service for a plenty of applications.. In the remaining part of the work, we will refer to those issues as ``horizontal'' requirements, that shows straightforwardly their multi-platform nature.

The second class of issues concerns platform integrity and user data privacy. In parituclar, the security architecture of those platform normally do not provide high assurance of privacy. Morevoer, from the point of a cloud provider, it is difficult to observe and inspect deeply the functioning of the whoe platform, leveraging commodity general purpose hypervisors (GPH).
General Purpose Hypervisors are executed at the most privileged level and so, it is complicated to detect and intervene on them in presence of malicious attack or misbehavior, without service interruption for all VMs hosted on the server. Cloudvisor \cite{cloudvisor:zhang} analyzed the problem proposing a nested architecture, where the GPH is executed at a lower level of privilege and a tiny hypervisor, responsible to prevent data disclosure, is executed at Ring 0.Hence, this kind of requirements could be called vertical requirements, because they involves layers executed at different layer of the software architecture.

To sum up, nested virtualization has shown to be a interesting pattern for modern cloud platform, that could improve dramatically features and integrity of cloud platform. However, no architecture had been presented to address vertical and horizontal class of problems at the same time. XenBlanket does not investigate cloud security concerns, limiting the focus on new potential feature of a coordinated multi-provider cloud. Cloudvisor proposed a security architecture that leverages nested virtualization without allowing the possibility of executing more than one nested hypervisor at time, that is a crucial requirements of a multi-provider cloud platform.

In this paper, we will present a nested architecture 
% Name?
that address this two different horizontal and vertical kind of issues at the same time.

\section{Related Work}
Cloud computing feature set could be enlarged exploiting two layers of virtualization instead of just one. The presence of two different hypervisors allows the possibility to specialize hypervisors functions. Instead of having a general purpose hypervisor we could split task in two different layers, interconnected with a well defined interface.
Each layer will have to care about a subset of feautures, simplifying overall development.

Introducing nested virtualization, we have to identify the two main actors involved in the virtualization process: bare-metal hypervisor (L0) and nested hypervisor (L1). In addition, we will refer to the nested guest as L2 guest. 
However, virtualization has a computational overhead. It is small when we are dealing with single layer virtualization, but it grows exponentially when we try to add further levels \cite{rec:virt}. This huge performance overhead has prevented nested virtualization adoption during last years.

Nowadays, several works demonstrate that it is possible to put in place low overhead nested virtualization using special approaches \cite{turtle:ibm} \cite{cloudvisor:zhang}. Moreover, other works try to recover performance loss exploiting multi-level consolidation, introduced in nested virtualization \cite{art:blan}.
Therefore, we will consider those overhead not as consistent to prevent the possibility to exploit a similar approach.

\section{Nested Virtualization}

Recursive virtualization} refers to a system architecture with several layers of virtualization. It consists in a ``virtualized'' virtualization, where the guest system virtualize a nested guest. The concept could be generalized to an arbitrary number of nested guest layers.
More formally, Popek \& Golberg \cite{Popek:1974:FRV:361011.361073} discussed this subject, introducing the concept of ``recursively virtualizable'' machine and giving an initial definition of recursive virtual machines: If it is possible for a virtual machine to run under itself a copy of the VMM that could also exhibit all the properties of the VMM and this procedure could be repeated until the resources of the systems are consumed, then the original machine is recursively virtualizable.
Moreover, the authors practically defined the features needed by a machine in order to be recursive virtualizable, exploiting concepts of virtualizable architecture and VMM\label{par:form} and introducing a second theorem \cite{Popek:1974:FRV:361011.361073}.

Looking at hypervisor specialization, we match those feature set with the different role of  L0 and L1 in a nested virtualized infrastructure. 
L0 is charged of the vertical aspects of the platform. It has to achieve  the best possible performances and enforcing the strongest possible level of isolation. L0 represents the last line of defense of the platform and a privileged point of global status observation.
On the contrary, L1 is responsible for horizontal feature set. it has to provide the widest possible support to different platforms and the capacity to virtualize adopting different techniques ( para-virtualization, hardware-assisted, dynamic binary translation). In this sense, Williams et al. propose a patched version of Xen, Xen Blanket, to accomplish a L1 virtualization layer, designed to run over different platforms.

\section{Limits}

Nowadays, nested virtualization presents several limits related to the hardware support and software state of implementation. At first, to provide nested virtualization, hypervisors has to provide an exact copy of hardware virtualization facilities. The most exploited technique, Full virtualization with hardware assistance (HVM) relies on a quite complex dedicated support.
This set of facilities assist instruction emulation, memory address translation and device management. Both Xen and KVM, the most used open-source hypervisor propose a nested VMX (nVMX) feature, that consists in the possibility to efficiently handle nested hardware virtualization, in presence of a single-layer hardware facilities.
To address this performance problems, several hardware producers proposed new facilities, like Intel VMCS shadowing, that reduces the amount of nested VMEXIT when handling a nested guest VMEXIT.However, nested virtualization performance are still not good and the feature is not yet stable on modern hypervisor \cite{xen:test}.

Secondly, due to the monolithic architecture, each feature introduced in GPH enlarges the TCB. Therefore, leveraging GPH, a huge amount of code is running in the most privileged mode and has to be trust. Several works discuss and analyze the overall TCB of modern GPH \cite{nova, xmhf}, considering GPH TCB always bigger than 100 KLoC. 
%A brief comparison of TCB of widespread GPH is depicted in figure \ref{fig:tcbchart}. 
It cannot be underestimated that implementing another important feature increase the number of LoC, enlarging the attack surface. More importanly, this is particularly true looking at the monolithic approach of commercial hyper-visors, where any addition to the hypervisor body enlarges the TCB.

To sum up, current nested virtualization limits are represented by a certain lack of dedicated feature in modern hypervisor and the impossibility to rely on a tiny TCB.
The first one is being addressed by hypervisor developers. The second one could simply not be achieved with modern hypervisors. 
Therefore, to address this we propose a step further introducing microhypervisors.

\section{Microhypervisor}

%\begin{figure}
%\caption{GPH TCB comparison \cite{nova}}
%\label{fig:tcbchart} 
%\begin{center}
%\includegraphics[width=\columnwidth]{figures/TCBchart.png}
%\end{center}
%\end{figure}

% Xen and KVM TCB
Consequentially, several works focused this problem leveraging a micro-hypervisor architecture, with a traditional microkernel-like design that tries to consistently reduce the amount of code running with the highest privilege level. A micro-kernel traditionally consists in a tiny core running in Kernel Mode, offering only few fundamental services to other modules running (scheduling, IPC, memory management) in a less privileged domain. All other components, in particular device drivers, are moved out in user-space.
Microkernels had been questioned because of the wasteful performance of first generations, due to IPC intrinsic cost. This argument had been completed confuted by Liedtke in 1993: \cite{liedtke1993ImpIPCbyKerDes,liedtke1995micCon}. However, even if it was demonstrated that feasibility of high performance microkernels, it was necessary to carefully design IPC to obtain good performances \cite{Hartig:1997:P9S:269005.266660}.

Microkernels have the capacity of easily redesign the hierarchy of privileges. A user application has normally to trust the micro-kernel and the set of user-space software it uses. However, this set of trusted set of drivers and tools have not to be trusted by every application executed in the system. Each different user application could use its own set of tools, correctly registering and configuring them. Normally, this allows microkernels to implement different ``personalities''. User applications shares the system but could have to deal with completely different system services in a transparent way. However, ``personalities'' interest had been hugely reduced by the fact that system services, device drivers and even application had to be completely re-conceived to fit this new design pattern. Therefore, with the introduction of VMs, microkernels lost the battle in favor of hypervisors, in terms of popularity. However, the great advantage of tiny TCB, provided by microkernel, was not provided also by hypervisors. As we analyzed before, they present normally a monolithic architecture, like the great part of modern OS. This feature kept alive the interest in microkernels, in particular trying to couple microkernel benefits with virtualization.

Therefore, leveraging this concept of microkernels, a micro hypervisor (MH) is composed from a small kernel providing only essential functions, like scheduling, IPC, memory management and VMX extension. Several micro-hypervisor have been proposed in literature \cite{xmhf, nova}. They rely on the micro-kernel paradigm, extending it to virtualization.

In addition, MH inherits the concept of micro-kernel to implement different ``personalities''. Those ``personalities'' are concretely a set of ``processes'' handling specific functions running in user-space. As those functions are implemented outside the kernel, they are not unique but they are changeable and could coexist at the same time. This feature was seen with interest because of the possibility to move some virtualization-related from the kernelspace to user-space. This extraction could allow to strengthen the platform, considering that user-space crashes are generally not fatal for the whole system.  A possible example, that will be analyzed in next section, is the user-space VMM component of Nova, Vancouver, that resides in user-space \cite{nova}.

\section {MH issues}

Despite the positives, MH approaches had demonstrated some limitations. Firstly, the straightforward one is the initial lack of hardware support, lack of device drivers are still a relevant Achilles' heel for this solutions. However, this kind of issue could be addressed leveraging I/O passthrough techniques, although it will cause exclusive allocation issues.

Moreover, as reported in previous sections, several cloud computing issues are simply not addressable with general purpose hypervisor (GPH). However, they are commonly widely adopted and represent the bedrock of existing platform. Therefore, GPH exports wide and solid interfaces leveraged by a plenty of different software that is able to properly communicate and act through them on the virtualization layer. On the contrary, MH lack of cloud toolkit integration. Widespread GPH are tightly connected with cloud toolkit and could present solid interfaces to communicate with different tools.

Finally, for sake of keeping the TCB small, MH could generally lack of several virtualization features, commonly available on modern GPH hypervisor. For instance, XMHF does not support multiple guests. This kind of drawbacks could represent important issues to handle and limit the potential interest in MH, even they are able to introduce several important features.

\section{Possible architectures}

The adoption of an extra layer of virtualization is capable to introduce a remarkable set of new features but it does not reduce the dimension of Trusted Code Base (TCB). 
MH could represent a new approach to improve the potential benefits provided by the adoption of nested virtualization.  It could be possible to combine positive features of GPH and MH approaches, leveraging the new extra layer of virtualization given by the nesting process.
Next sections will present a brief analysis of all four possible combinations of GPH and MH, underlying potentialities, drawbacks and potential cost of implementation.

\subsection{GPH over GPH}

The first analyzed architecture concerns only GPH. At first, it is important to recall the most interesting achievable by such architectures. In particular, several remarkable features are live migration over different platform, transparent migration of virtualized clusters between different data-centers and improved isolation between customer instances and administration, improved control of the nested virtualization layer L1.
This approach allows us also to exploit different virtualization techniques. For instance, it could be possible to run at the same time PV and HVM nested instances, leveraging  Xen on KVM architecture ( if KVM nested VMX performs as expected).
The choice of KVM as L0 is due to the general smaller TCB, compared with standard Xen (without stub domains) and this is important due to the high privilege level of L0.

Supposing the completeness and the correctness of nested virtualization hardware facilities in widespread hypervisor, GPH over GPH represents the best achievable architecture leveraging mature hypervisors. To sum up, it could be possible to achieve live-migration between different platform and other features related to VM management flexibility but it could not still be possible to reduce the TCB. Analyzing MH approaches, the aim is try at most to retain to features, trying to introduce isolation and privacy improvements due to the lower TCB promised.

%Although this architectures presents an extra layer between the nested guest, normally controlled by the cloud user and the system critical part, L0, the latter is far from a satisfying security level. The TCB of KVM that is represented by the entire Linux Kernel is normally around 200 KLoCs, considering the medium size of a deployed Linux kernel. 
%Moreover, default KVM configuration does not provide a transparent protection mechanism between different VMs, so that compromising one instance of QEMU, that is executed in user-space, an attacker could compromise all the others VMs on the platform.

\section{MH over MH}

MH on MH in reported only for completeness. The interest on such architecture is limited because it will still present all cited issues of MH.
Moreover, the lack of nested VMX functionalities prevents us to concretely test this kind of architecture, because MH normally rely on VMX facilities in order to create HVM guests.
To conclude, nested virtualization is exploited in order to fill eco-systems gaps with GPH but MH over MH does not propose any improvement in this sense.

\section{MH over GPH}
\label{par:mog}

The second architecture, proposed in order to reduce security concerns, is built combining an GPH at L0 and a MH at L1. 
%Several MH are proposed in literature focusing on different .
%Therefore, this model tries to exploit microhypervisor as L1 hypervisor in order to enforce better isolation between L2 and L0. L1 will exploit the emulated nested VMX provided by GPH L0 and virtualize L2. 
%With this approach, we try to isolate the attacker in a VM virtualized on a secured layer, in order to prevent attacks to L0. The attacker would have to face with an essential virtualization layer and it would not be able to attack the back-end virtualization layer. 
An interesting point of this approach is the fact of being easily be deployed, without requiring any extra code addition to the GPH, L0.  Existing GPH-based platforms could potentially exports nested VMX features to those guests. Even if this is not so usual in modern cloud platform, this lack is considered temporary and related only to the immaturity of those features.
Secondly, the number of driver needed to properly support devices is significantly small. With a MH, the only I/O communication that the platform could be able to deliver, without the porting or addition of code, is Multi-level device assignment. This kind of approach is feasible only implementing in the L0 hypervisor software techniques like vIOMMU, that extends the IOMMU capabilities to several layers of virtualization.

In addition, the amount of code running in L0 kernel space is still as big as before and, therefore, no TCB reduction. 
Finally, adopting a MH as L1 prevents to use inter-platform enhanced features. As analyzed when dealing with MH design, the price of having a tiny core is normally paid with a lack of certain functionalities. For example, XMHF only supports one single guest. It could be easily noticed that this approach is not architecturally able to provide any new "horizontal" features, leveraging on nested virtualization. Therefore, the architecture is only able to provide the same features that are provided by today deployed ones, with an enhanced but not sufficient level of isolation.

\section{GPH over MH}
\label{par:gom}

In this configuration, the amount of code that have to be trusted is lower than other architectures, due to the modular structure of micro-kernels., executed at L0.
The possibility to exploit at L1 a GPH like Xen or KVM allows to deploy a similar architecture used in production today. The L1 GPH is normally integrated with modern cloud toolkits like OpenStack.
Therefore, the architecture presents a stronger isolation leveraging on nested virtualization, providing in particular a strong tiny core. The MH layer will be transparent from the user and it will not not any change in the functioning of the platform.

The first class of possible drawbacks of this architecture is represented by the necessity to expose to the L1 nested hypervisor the nested VMX extension.  This is experimentally done by GPH like Xen and KVM, but it is generally not included in the essential features provided by a minimal micro-hypervisor. As reported in XMHF paper \cite{xmhf}, the amount of lines of code for nested VMX is around 1KLOC. 
More generally, from an architectural point of view, the general MH minimalist attitude could represent a limitation to general preservation of existing platform functions. The constraint of privileged code size forces normally MH developers to drop some features, considered not essential. Therefore, several limitations could arise not only for nested virtualization proper features, but also for other class of features. For instance, XHMF is able only to virtualize one L1 instance at time. This is acceptable if the platform objective is limited to be provided of a secure core, but dramatically limits the possibility of having a platform model able to give the user the possibility to run its proper hypervisor as L1.

Another important class is the necessity to explicit hardware devices support. Device drivers represents a general Achilles' heel of MH, due the general cost of development. However, this class of problem could be addressed leveraging device assignment. L1 GPH hypervisor will handle the real device, using the proper driver now demoted to run in a lower level of privilege. It has to be underlined that this solution could represent a feasible approach only if it possible to overcome the exclusive allocation trouble (SR-IOV).

\section{Architecture comparison}

To conclude, this section provides a short comparison between different platforms, contrasting architecture features above the two main different types of benefits given by nested virtualization, as reported widely in literature \cite{art:blan, nested:berghmans,  cloudvisor:zhang, turtle:ibm}. The first could be identified in the augmented degree of isolation the architecture could enforce between host OS and the "customer" point of access, leveraging a trusted core. The second one is represented by the possibility to leverage multiple virtualization layers to easily define inter-platform policy and operations between different platforms.
In the above paragraphs, four different possible architectures were presented with the set of  substitute a GPH  with a MH, looking at the positive and negative new features. The starting point is GPH over GPH, a well analyzed approach in literature \cite{turtle:ibm, art:blan, }. We underlined features and limitations of this architecture in the first paragraphs and in the

To conclude, GPH over MH is the most interesting one in order to enlarge nested virtualization promised benefit. This architecture with a solid and secured MH as L0 is the only able to reduce the platform TCB and this could be a key point to obtain more protected infrastructure, .
Firstly, adopting a MH as architecture trusted core, we will have a secure layer guaranteeing that administrator could not lose the control of the system. Klein et Al. \cite{Klein:2009} have demonstrated that a kernel with less than 10KLoC could be formally verified, under several conditions. This is far from be achievable today with standard GPH over GPH approach where the core is not tiny and the TCB is an order of 
magnitude bigger than MH \cite{nova}.

However, the effort to set up a GPH over MH is strongly dependent on which technique leveraging for virtualizing L2. As stated in paragraph \ref{par:gom}, with L2 guest virtualized with the hardware assitance technique, the most popular one, the MH needs to export nVMX at least. 
On the contrary, adopting paravirtualization/DBT for L2 does not necessitate a particular hardware support and so it could not be necessary patch the MH to set up a platform. 
However, it has not be underestimated, however, that para-virtualization supports only a subset of OS, and could not be considered a general solution and DBT has shown important overhead.
%The last point of interest is related to device management. If MH act as L0 hypervisor, it has to deal with hardware devices, and this could necessitate of a significant code writing. In fact, there are several technologies, like Direct Assignment, allowing to bypass hypervisor control of a device. Those techniques are useful but they reduce the flexibility in the device management, requiring for example exclusive allocation of a device to a guest. To contrast this drawbacks, PCI-E propose the possibity to multiplex device with a set of virtual devices called virtual functions proposed to the L0 hypervisor, that will relax the contraint of exclusive allocation. Unfortunately this approach could not give the same flexibilty given by other techniques, like para-virtualization 
%With MH over GPH, MH could directly assigne devices to GPH that is able to implement  para-virtualize device, for example. 
%This will relief the MH of a part of devices management, giving to the GPH the responsibility to drive concretely the device but without losing interesting features. Therefore, we will need a limited amount of code to handle device without have to deal with each specific device.

To sum up,  the adoption of MH could  increase feature set introduced by nested virtualization adoption. The first impeding problem is the implementation of MH, able to propose to guest to virtualize with hardware assistance.

\section{Xen over NOVA}

\paragraph{Inter-platform features}
\paragraph{Tiny TCB}
\paragraph{Leveraging existing APIs}
\paragraph{Fail-safe components architecture}

%
%%
%%%%Improving the Status Quo
%Udo Steinberg NOVA 6
%Virtualization layer is critical. Make it as small as possible.
%Design Principles:
%1)Fine-grained functional decomposition
% Microhypervisor (privileged)
% Multiple user-level VMMs (unprivileged)
% User-level drivers, applications (unprivileged)
%2)Principle of least privilege among all components
% Capability-based authorization model
%Ideas adopted from the microkernel world
%Those still existing limitations to nested HVM, made we focus on nested PV.
%Leveraging on the new architecture presented we propose a security architecture based on Xen and NOVA. 
%
%Xen is a well-known and wide-used hypervisor that is the base of several of most important cloud platforms.  We focus in particularly on the Xen support for para-virtualization. 
%
%Available from the first version of Xen in 2003, PV does not requires any specific hardware facilities. Even if some features drop in Amd64 architecture leaded to a little performance loss for PV compared to HVM guests, the real limit of PV is represented by the impossibility to virtualize Windows and other proprietary operating systems.
%However this technique is currently still exploited in widespread cloud platform like Amazon EC2.
%
%NOVA is a tiny micro-hypervisor able to virtualize several guests. It is a recent but promising open-source project developed at Technische Universitat Dresden. The aim of project is double: minimize the amount of code running in kernel mode and try boost virtualization.
%
%NOVA implements some interesting features, in particular from security point of view. It adopts a capability-based authorization model in order to implement the principle of least privilege among components. It supports SMP guests, VT-d Direct device assignement and NPT/EPT.
%
%The modular architecture of NOVA allows also to enforce a increased separation between different L1 guest. Leveraging the modular architecture of NOVA, each L1 guest communicates directly with a separate instance of the Virtual Machine Monitor, component of the hypervisor executed in user space. 
%
%The concept is to boot and use Xen as L1 hypervisor, inside a NOVA VMs. Xen hypervisor will be able to virtualize complete PV guests in a secured Cloud Platform. The presence of NOVA is completely transparent for the user.

%Therefore, leveraging NOVA as L0 our platform could propose inter-platform features to the user. We have the opportunity to exploit a complete cloud toolkit with the Xen hypervisor implementing features like inter-platform migration and improved redundancy, as described in the Xen Blanket approach.

\section{Conclusion and future Works}

Future works will firstly concern the study of available MH, trying to identify which project could be suitable to act as L0 in a GPH over MH platform. As underlined previously, the key feature of L0 MH will be the capacity of having a tiny TCB. Moreover, additional requirements concerns the flexibility in terms VMs ( number of VM supported at the same time, CPU handling, I/O management). Finally, another criteria for choosing the L0 hypervisor will be the evaluated effort needed for integrating a nested HV extension the to the base project.
On the other hand, L1 choice is much simpler. Previous analysis identified Xen as the most suitable hypervisor for this role. The capacity to support different virtualization techniques added to the ability of running without HV support, makes Xen suitable to act as compatibility layer.

After this initial study, the first step will be deploying a minimal nested architecture testing ``out-of-the-box'' features and limitations. The following part will concern the conception of patch and extensions in order to adapt the general design of MH to properly act as L0.
The final goal is to provide an architecture able to leverage nested virtualization cloud platforms improvements, having a smaller TCB and allowing customers to leverage flexible migrations and scheduling.

%\section{Conclusion}
\printbibliography
\end{document}

### 2014-03-03 draft.tex
Design principles extraction

@@
 The second class of issues concerns platform integrity and user data privacy.
-In parituclar, the security architecture of those platform normally do not provide high assurance of privacy.
-Morevoer, from the point of a cloud provider, it is difficult to observe and inspect deeply the functioning of the whoe platform, leveraging commodity general purpose hypervisors (GPH).
+At first, the security architecture of those platform normally does not provide high assurance of privacy.
+Morevoer, from the point of a cloud provider, it is difficult to observe and inspect deeply the functioning of the whole platform, leveraging commodity general purpose hypervisors (GPH).
 General Purpose Hypervisors are executed at the most privileged level and so, it is complicated to detect and intervene on them in presence of malicious attack or misbehavior, without service interruption for all VMs hosted on the server.
@@
 To sum up, nested virtualization has shown to be a interesting pattern for modern cloud platform, that could improve dramatically features and integrity of cloud platform.
-However, no architecture had been presented to address vertical and horizontal class of problems at the same time.
-XenBlanket does not investigate cloud security concerns, limiting the focus on new potential feature of a coordinated multi-provider cloud.
-Cloudvisor proposed a security architecture that leverages nested virtualization without allowing the possibility of executing more than one nested hypervisor at time, that is a crucial requirements of a multi-provider cloud platform.
+From this initial analysis it is possible to extract several design principles, reported in table \ref{int:des}
 
-In this paper, we will present a nested architecture 
-% Name?
-that address this two different horizontal and vertical kind of issues at the same time.
+\begin{table}
+\label{int:des}
+\begin{tabular}{llll}
+\toprule
+Id & Functional design Principle & Class &\\
+\midrule
+   DS1 & User-Data privacy & Horizontal & \\
+   DS2 & Inter-platform support & Horizontal & \\
+   DS3 & Small TCB & Vertical & \\
+   DS4 & Fail-Safe modular architecture &  & \\
+\bottomrule
+      \end{tabular}
+\end{table}
+
+\paragraph{Inter-platform support} The cloud architecture has to be designed in order to support directly different platforms and service providers. 
+\paragraph{Fail-Safe modular architecture} In order to recover soft failures, the archtecure is supposed to be fail-safe, being able of automatically migrate instances or user-data in presence of a menace, but also reactivating the correct functioning in presence of a software failure.
+In order to satisfy this property, the amount of safety-critical components has to be as lowest as possible and the overall design has to be kept as modular as possible
+\paragraph{Small TCB} The amount of lines of code executed at the highest level of privilege is normally proportional to the number of potential failures and bugs in the platform.
+According to this, an architecture presenting a tiny TCB will improve its safety and integrity properties by design.
+\paragraph{User-Data privacy} The architecture's administrator would not have the right to inspect and analyze user data and instances, without the explicit consensus with the user above mutually trusted services.
+User-data privacy is a major challenge of modern cloud platforms.
+It is important to identify a trade-off that allows providers to identify and intervene on possible menaces, but also guarantee the impossibility for a curious or malicious operator to access to them without restrictions.
+
+Moreover, beside this functional design principles specifying which features have to be achieved by the platform, another set of non-functional requirements could be identified.
+Those requirements concern not directly features or requirements, but the feasibility of the architecture.
+For example, a major issue could be represented by the amount of code lines that have to be added and maintained to an existing architecture to meet the functional requirement.
+Even it does not represent a technical limitation to a certain proposal, it could become a ``show-stopper'' factor for its utilization.
 
@@
 Each layer will have to care about a subset of feautures, simplifying overall development.
+However, no architecture had been presented to address vertical and horizontal class of problems at the same time.
+XenBlanket does not investigate cloud security concerns, limiting the focus on new potential feature of a coordinated multi-provider cloud.
+Cloudvisor proposed a security architecture that leverages nested virtualization without allowing the possibility of executing more than one nested hypervisor at time, that is a crucial requirements of a multi-provider cloud platform.
+In fact, in this scenario, each user has to run its own L1 hypervisor instance to enforce a compatibility layer, removing incompatibilities between different providers.
 
@@
 
-\section{Nested Virtualization}
+\section{Classic nested virtualization approach}
 
-Recursive virtualization} refers to a system architecture with several layers of virtualization.
+Recursive virtualization refers to a system architecture with several layers of virtualization.
 It consists in a ``virtualized'' virtualization, where the guest system virtualize a nested guest.
@@
 
-\section{Limits}
+\subsection{Limits}
 
@@
 
-\section {MH issues}
+\subsection {MH open issues}
 
@@
 
-\section{Possible architectures}
+\section{Design Principles}
 
@@
 
-\section{MH over MH}
+\subsection{MH over MH}
 
@@
 
-\section{MH over GPH}
+\subsection{MH over GPH}
 \label{par:mog}
@@
 
-\section{GPH over MH}
+\subsection{GPH over MH}
 \label{par:gom}
@@
 
-\section{Architecture comparison}
+\subsection{Architecture comparison}
 
@@
 
-\section{Xen over NOVA}
-
+\section{Implementation}
+\subsection{Xen over NOVA}
 \paragraph{Inter-platform features}
@@
 \paragraph{Fail-safe components architecture}
+\subsection{}
 

### 2014-03-05 draft.tex
DS added + Comparison table

@@
 % These packages are all incorporated in the memoir class to one degree or another...
+\usepackage{amsfonts}
 
@@
 \section{Introduction}
-
+\label{ref:intro}
 Cloud computing platform are a solid reality nowadays. Modern cloud platforms are able to provide elastic use of resources, optimizing their allocation in order to meet the user needings.
@@
 The second class of issues concerns platform integrity and user data privacy.
-At first, the security architecture of those platform normally does not provide high assurance of privacy.
-Morevoer, from the point of a cloud provider, it is difficult to observe and inspect deeply the functioning of the whole platform, leveraging commodity general purpose hypervisors (GPH).
-General Purpose Hypervisors are executed at the most privileged level and so, it is complicated to detect and intervene on them in presence of malicious attack or misbehavior, without service interruption for all VMs hosted on the server.
-Cloudvisor \cite{cloudvisor:zhang} analyzed the problem proposing a nested architecture, where the GPH is executed at a lower level of privilege and a tiny hypervisor, responsible to prevent data disclosure, is executed at Ring 0.Hence, this kind of requirements could be called vertical requirements, because they involves layers executed at different layer of the software architecture.
+On one hand, from an user-centric point of view, the user would like to preserve privacy of its data when it stores data on the cloud platform.
+However, hypervisor is normally the most privileged task executed in the platform and therefore a curious or malicious administrator could easily retrieve those data \cite{cloudvisor:zhang}.
+On the other hand, from a user-centric point of view, it is difficult to observe and inspect deeply the functioning of the whole platform, leveraging commodity general purpose hypervisors (GPH).General Purpose Hypervisors are executed at the most privileged level and so, it is complicated to detect and intervene on them in presence of malicious attack or misbehavior, without service interruption for all VMs hosted on the server.
+To sum up, both visions collide with the execution of the hypervisor at the highest level of privileged.
+This attitude prevents the platform administrator from a real monitoring of the whole platform, but also the possibility for users to concretely inspects actions that platform administrators do on user data.
+In the following paragraphs, we will make reference to this class of problem as ``Vertical'' requirements.
 
-To sum up, nested virtualization has shown to be a interesting pattern for modern cloud platform, that could improve dramatically features and integrity of cloud platform.
-From this initial analysis it is possible to extract several design principles, reported in table \ref{int:des}
+To resume, in order to present a complete view of the architecture, we identified five key design principles, that a cloud computing platform has to implement in order to solve the cited problems.
+In the following paragraphs, those principles will represent a grid of analysis of different architectures.
+
+\paragraph{Inter-platform support} The cloud architecture has to be designed in order to support directly different platforms and service providers. 
+\paragraph{Fail-Safe modular architecture} In order to recover soft failures, the archtecure is supposed to be fail-safe, being able of automatically migrate instances or user-data in presence of a menace, but also reactivating the correct functioning in presence of a software failure.
+In order to satisfy this property, the amount of safety-critical components has to be as lowest as possible and the overall design has to be kept as modular as possible
+\paragraph{Small TCB} The amount of lines of code executed at the highest level of privilege is normally proportional to the number of potential failures and bugs in the platform.
+According to this, an architecture presenting a tiny TCB will improve its safety and integrity properties by design.
+\paragraph{User-Data privacy} A curious or malicious administrator ( or another platform user) would not have the right to inspect and analyze user data and instances, without the explicit consensus with the user above mutually trusted services.
+User-data privacy is a major challenge of modern cloud platforms.
+It is important to identify a trade-off that allows providers to identify and intervene on possible menaces, but also guarantee the impossibility for a curious or malicious operator to access to them without restrictions.
+\paragraph{3rd party compatibility} Avoid disruption of existing 3rd part applications compatibility could represent a major principle to consider.
+Platform interfaces has to be kept as close as possible to the existing one, preventing application rewrite.
+
+Moreover, beside this functional design principles specifying which features have to be achieved by the platform, another set of non-design principles could be identified.
+Those principles does not concern directly features or aspects of the architecture of the platform, but the effective feasibility of the architecture.
+\paragraph{Minimize code addition} a major issue could be represented by the amount of code lines that have to be added and maintained to an existing architecture to meet the functional requirement.
+Even it does not represent strictly a technical limitation to a certain proposal, it could become a ``show-stopper'' factor for its utilization.
+
+In our analysis, it will be kept in consideration not only the overall suitability of the architecture, but also a quantification of the effort needed to develop and deploy it.
+In table \ref{int:des} it is possible to observe a resume of previous cited principles.
 
@@
 \midrule
-   DS1 & User-Data privacy & Horizontal & \\
+   DS1 & User-Data privacy & Vertical & \\
    DS2 & Inter-platform support & Horizontal & \\
    DS3 & Small TCB & Vertical & \\
-   DS4 & Fail-Safe modular architecture &  & \\
+   DS4 & Fail-Safe modular architecture & Vertical & \\ 
+   DS5 & Backwards 3rd party compatibility & Horizontal \\
+   NDS1 & Minimize code addition & $\times$& \\
 \bottomrule
@@
 
-\paragraph{Inter-platform support} The cloud architecture has to be designed in order to support directly different platforms and service providers. 
-\paragraph{Fail-Safe modular architecture} In order to recover soft failures, the archtecure is supposed to be fail-safe, being able of automatically migrate instances or user-data in presence of a menace, but also reactivating the correct functioning in presence of a software failure.
-In order to satisfy this property, the amount of safety-critical components has to be as lowest as possible and the overall design has to be kept as modular as possible
-\paragraph{Small TCB} The amount of lines of code executed at the highest level of privilege is normally proportional to the number of potential failures and bugs in the platform.
-According to this, an architecture presenting a tiny TCB will improve its safety and integrity properties by design.
-\paragraph{User-Data privacy} The architecture's administrator would not have the right to inspect and analyze user data and instances, without the explicit consensus with the user above mutually trusted services.
-User-data privacy is a major challenge of modern cloud platforms.
-It is important to identify a trade-off that allows providers to identify and intervene on possible menaces, but also guarantee the impossibility for a curious or malicious operator to access to them without restrictions.
-
-Moreover, beside this functional design principles specifying which features have to be achieved by the platform, another set of non-functional requirements could be identified.
-Those requirements concern not directly features or requirements, but the feasibility of the architecture.
-For example, a major issue could be represented by the amount of code lines that have to be added and maintained to an existing architecture to meet the functional requirement.
-Even it does not represent a technical limitation to a certain proposal, it could become a ``show-stopper'' factor for its utilization.
-
 \section{Related Work}
+% Intro + performance issues
 Cloud computing feature set could be enlarged exploiting two layers of virtualization instead of just one.
 The presence of two different hypervisors allows the possibility to specialize hypervisors functions.
-Instead of having a general purpose hypervisor we could split task in two different layers, interconnected with a well defined interface.
-Each layer will have to care about a subset of feautures, simplifying overall development.
-However, no architecture had been presented to address vertical and horizontal class of problems at the same time.
-XenBlanket does not investigate cloud security concerns, limiting the focus on new potential feature of a coordinated multi-provider cloud.
-Cloudvisor proposed a security architecture that leverages nested virtualization without allowing the possibility of executing more than one nested hypervisor at time, that is a crucial requirements of a multi-provider cloud platform.
-In fact, in this scenario, each user has to run its own L1 hypervisor instance to enforce a compatibility layer, removing incompatibilities between different providers.
+Instead of having a general purpose hypervisor we could split the hypervisor role in two different layers, interconnected with a well defined interface.
+Each layer will have to care about a subset of features, simplifying overall development.
+However, nested virtualization on x86 architecture faced impeding performance issues \cite{rec:virt}.
+
+% Perfomance issuess adressed
+However, several works demonstrate that it is possible to put in place low overhead nested virtualization using special approaches \cite{turtle:ibm} .
+Moreover, other works try to recover performance loss exploiting multi-level consolidation, introduced in nested virtualization \cite{art:blan}.
+In conclusion, nested virtualization slowdown could be considered at now an acceptable price to achieve advanced features and, in addition, hardware manifacturers are developing and proposing new hardware facilities to reduce this overhead \cite{vmcs:nakajima}.
+
+To sum up, no architecture had been presented to address vertical and horizontal class of problems at the same time until now.
+XenBlanket does not investigate cloud architectural security concerns (DS3,DS4), limiting the focus on new potential feature of a coordinated multi-provider cloud.
+Cloudvisor proposed a security architecture that leverages nested virtualization without allowing the possibility of executing more than one nested hypervisor at time, that is a crucial requirements of a multi-provider cloud platform (DS2).
+In a SuperCloud scenario, each user has to run its own L1 hypervisor instance to enforce a compatibility layer, overcoming incompatibilities between different providers.
+
+Other approaches leveraging only one layer of virtualization does not achieve the whole design-principle satisfaction.
+Self-service cloud approach (SSC) \cite{ssc:art} design a complete security architecture to overcome major security and privacy issues of currently deployed cloud infrastructures, in order to alleviate the different needing user-centric and provider-centric vision of cloud computing presented in \ref{par:intro}.
+Even if this architecture reproduces an approach similar to the SuperCloud without leveraging an extra layer of virtualization ( DS
+
+\section{Classic nested virtualization approach}
+
+Nested virtualization refers to a system architecture with two layers of virtualization.
+It consists in a ``virtualized'' virtualization, where the guest system virtualize a nested guest.
+The concept could be generalized to an arbitrary number of nested guest layers, introducing recursive virtualization.
+%More formally, Popek \& Golberg \cite{Popek:1974:FRV:361011.361073} discussed this subject, introducing the concept of ``recursively virtualizable'' machine and giving an initial definition of recursive virtual machines: If it is possible for a virtual machine to run under itself a copy of the VMM that could also exhibit all the properties of the VMM and this procedure could be repeated until the resources of the systems are consumed, then the original machine is recursively virtualizable.
 
@@
 In addition, we will refer to the nested guest as L2 guest.
-However, virtualization has a computational overhead.
+However, virtualization introduces a computational overhead.
 It is small when we are dealing with single layer virtualization, but it grows exponentially when we try to add further levels \cite{rec:virt}.
-This huge performance overhead has prevented nested virtualization adoption during last years.
+This huge performance overhead has prevented nested virtualization adoption during last years and had been finally addressed by Ben Yehuda et Al \cite{turtles:ibm} proposing an architecture with acceptable overhead.
 
-Nowadays, several works demonstrate that it is possible to put in place low overhead nested virtualization using special approaches \cite{turtle:ibm} \cite{cloudvisor:zhang}.
-Moreover, other works try to recover performance loss exploiting multi-level consolidation, introduced in nested virtualization \cite{art:blan}.
-Therefore, we will consider those overhead not as consistent to prevent the possibility to exploit a similar approach.
-
-\section{Classic nested virtualization approach}
-
-Recursive virtualization refers to a system architecture with several layers of virtualization.
-It consists in a ``virtualized'' virtualization, where the guest system virtualize a nested guest.
-The concept could be generalized to an arbitrary number of nested guest layers.
-More formally, Popek \& Golberg \cite{Popek:1974:FRV:361011.361073} discussed this subject, introducing the concept of ``recursively virtualizable'' machine and giving an initial definition of recursive virtual machines: If it is possible for a virtual machine to run under itself a copy of the VMM that could also exhibit all the properties of the VMM and this procedure could be repeated until the resources of the systems are consumed, then the original machine is recursively virtualizable.
-Moreover, the authors practically defined the features needed by a machine in order to be recursive virtualizable, exploiting concepts of virtualizable architecture and VMM\label{par:form} and introducing a second theorem \cite{Popek:1974:FRV:361011.361073}.
-
-Looking at hypervisor specialization, we match those feature set with the different role of  L0 and L1 in a nested virtualized infrastructure. 
-L0 is charged of the vertical aspects of the platform.
+Nested virtualization feasibility opened a plenty of new possibilities and perspectives, that have been largely discussed in literature \cite{rec:virt,turtles:ibm,art:blan}.
+The first aspect in common among different works proposed in literature is to diversify hypervisor functions, introducing new features but keeping the existing ones.
+In a general analysis of existing works, it could be possible to identify the general role of each different layer.
+On one hand, L0 is charged of the vertical aspects of the platform.
 It has to achieve  the best possible performances and enforcing the strongest possible level of isolation.
 L0 represents the last line of defense of the platform and a privileged point of global status observation.
-On the contrary, L1 is responsible for horizontal feature set.
+
+On the other hand, L1 is responsible for horizontal feature set.
 it has to provide the widest possible support to different platforms and the capacity to virtualize adopting different techniques ( para-virtualization, hardware-assisted, dynamic binary translation).
@@
 propose a patched version of Xen, Xen Blanket, to accomplish a L1 virtualization layer, designed to run over different platforms.
+This concretely allows to run over different physical platform the same VM, and live-migrate it at the occurrence.
 
-\subsection{Limits}
+\subsection{Classical nested virtualization limits}
 
@@
 At first, to provide nested virtualization, hypervisors has to provide an exact copy of hardware virtualization facilities.
-The most exploited technique, Full virtualization with hardware assistance (HVM) relies on a quite complex dedicated support.
-This set of facilities assist instruction emulation, memory address translation and device management.
-Both Xen and KVM, the most used open-source hypervisor propose a nested VMX (nVMX) feature, that consists in the possibility to efficiently handle nested hardware virtualization, in presence of a single-layer hardware facilities.
-To address this performance problems, several hardware producers proposed new facilities, like Intel VMCS shadowing, that reduces the amount of nested VMEXIT when handling a nested guest VMEXIT.However, nested virtualization performance are still not good and the feature is not yet stable on modern hypervisor \cite{xen:test}.
+The most exploited technique, full virtualization with hardware assistance (HV), relies on a quite complex dedicated support.
+This set of facilities assists instruction emulation, memory address translation and device management.
+Both Xen and KVM propose nested VMX/SVN features, that consists in the possibility to efficiently handle nested hardware virtualization, in presence of a single-layer hardware facilities.
+%To address this performance problems, several hardware producers proposed new facilities, like Intel VMCS shadowing, that reduces the amount of nested VMEXIT when handling a nested guest VMEXIT.
+Although nested virtualization performance are still not good and the feature is not yet stable on modern hypervisor \cite{xen:test}, the feature is under heavy development and it will be hopefully in a usable state soon.
 
@@
 Several works discuss and analyze the overall TCB of modern GPH \cite{nova, xmhf}, considering GPH TCB always bigger than 100 KLoC.
+This approach directly collides with one design principles introduced in the introduction paragraph (DS3), that identified a small TCB as a major requirement for a solid architecture.
 %A brief comparison of TCB of widespread GPH is depicted in figure \ref{fig:tcbchart}. 
-It cannot be underestimated that implementing another important feature increase the number of LoC, enlarging the attack surface.
-More importanly, this is particularly true looking at the monolithic approach of commercial hyper-visors, where any addition to the hypervisor body enlarges the TCB.
+%More importanly, this is particularly true looking at the monolithic approach of commercial hyper-visors, where any addition to the hypervisor body enlarges the TCB.
 
-To sum up, current nested virtualization limits are represented by a certain lack of dedicated feature in modern hypervisor and the impossibility to rely on a tiny TCB.
+Finally, the above architecture does not improve consistently the overall stability of the infrastructure.
+The underlying core, L0, has the same problem of single-layer GPH architecture.
+Infact, the consistent TCB implies also that a failure experienced in one of the critical components will simply produce a generalized failure of the whole platform, not satysfing the DS4.
+This kind of architecture, in particular, does not leverage completely the possibility to hide behind a compatible layer, a really modular architecture with only a tiny critical core.
+
+To sum up, current nested virtualization limits are represented by a certain lack of dedicated feature in modern hypervisor and the impossibility to rely on a tiny TCB (DS3,DS4).
 The first one is being addressed by hypervisor developers. The second one could simply not be achieved with modern hypervisors.
-Therefore, to address this we propose a step further introducing microhypervisors.
 
@@
 This argument had been completed confuted by Liedtke in 1993: \cite{liedtke1993ImpIPCbyKerDes,liedtke1995micCon}.
-However, even if it was demonstrated that feasibility of high performance microkernels, it was necessary to carefully design IPC to obtain good performances \cite{Hartig:1997:P9S:269005.266660}.
 
@@
 They rely on the micro-kernel paradigm, extending it to virtualization.
+MH naturally satisfy vertical design principles (DS1,DS3, DS4).
+The tiny core could be considered the only system-critical componente of the architecture and if it possible to assure its correctness it could represent the last line of defense, where it could be possible to rebuild the architecture.
 
@@
 As those functions are implemented outside the kernel, they are not unique but they are changeable and could coexist at the same time.
-This feature was seen with interest because of the possibility to move some virtualization-related from the kernelspace to user-space.
-This extraction could allow to strengthen the platform, considering that user-space crashes are generally not fatal for the whole system.
+This feature was seen with interest because of the possibility to move some virtualization-related from the kernelspace to user-space and to strictly separate piece of software executed for different users.
+This extraction process drives to a more solid platform, considering that user-space crashes are generally not fatal for the whole system.
 A possible example, that will be analyzed in next section, is the user-space VMM component of Nova, Vancouver, that resides in user-space \cite{nova}.
@@
 Despite the positives, MH approaches had demonstrated some limitations.
-Firstly, the straightforward one is the initial lack of hardware support, lack of device drivers are still a relevant Achilles' heel for this solutions.
-However, this kind of issue could be addressed leveraging I/O passthrough techniques, although it will cause exclusive allocation issues.
+Firstly, the straightforward one is the initial lack of hardware support, lack of device drivers are still a relevant Achilles' heel for this solutions, inherited from its ancestors microkernels.
+Fortunately, this kind of issue could be addressed leveraging I/O passthrough techniques, although it will cause exclusive allocation issues.
 
@@
 Therefore, GPH exports wide and solid interfaces leveraged by a plenty of different software that is able to properly communicate and act through them on the virtualization layer.
-On the contrary, MH lack of cloud toolkit integration.
-Widespread GPH are tightly connected with cloud toolkit and could present solid interfaces to communicate with different tools.
+On the contrary, MH lack of cloud toolkit integration, requiring an explicit support to be written, not satysfing 3rd party compatibility requirements (DS5).
+GPH are tightly connected with cloud toolkit and could present solid interfaces to communicate with different tools.
 
@@
 
-\section{Design Principles}
+\section{Architecture comparison}
 
@@
 MH could represent a new approach to improve the potential benefits provided by the adoption of nested virtualization.
-It could be possible to combine positive features of GPH and MH approaches, leveraging the new extra layer of virtualization given by the nesting process.
-Next sections will present a brief analysis of all four possible combinations of GPH and MH, underlying potentialities, drawbacks and potential cost of implementation.
+In the following subsections, it will presented the four possible architectures achievable ``nesting'' of GPH and MH, underlying potentialities, drawbacks and potential cost of implementation.
 
@@
 At first, it is important to recall the most interesting achievable by such architectures.
-In particular, several remarkable features are live migration over different platform, transparent migration of virtualized clusters between different data-centers and improved isolation between customer instances and administration, improved control of the nested virtualization layer L1.
+In particular, several remarkable features are live migration over different platform, transparent migration of virtualized clusters between different data-centers and improved isolation between customer instances and administration, improved control of the nested virtualization layer L1 (horizontal requirements DS.
 This approach allows us also to exploit different virtualization techniques.
@@
 
+\begin{table*}
+\centering
+\begin{tabular}{llllll}
+\toprule
+Id & Functional design Principle & GPH over GPH & MH over MH & MH over GPH & GPH over MH\\
+\midrule
+DS1 & User-Data privacy & $\times$  & $\times$ & $\times$ & $\times$ \\
+   DS2 & Inter-platform support & $\times$  & $\times$ & $\times$ & $\times$ \\
+   DS3 & Small TCB & $\times$  & $\times$ & $\times$ & $\times$ \\
+   DS4 & Fail-Safe modular architecture & $\times$  & $\times$ & $\times$ & $\times$ \\
+   DS5 & Backwards 3rd party compatibility & $\times$  & $\times$ & $\times$ & $\times$ \\
+   NDS1 & Minimize code addition & $\times$  & $\times$ & $\times$ & $\times$ \\
+\bottomrule
+      \end{tabular}
+\end{table*}
+
 In this configuration, the amount of code that have to be trusted is lower than other architectures, due to the modular structure of micro-kernels., executed at L0.
@@
 
-\subsection{Architecture comparison}
+\subsection{Conclusion}
 

### 2014-03-07 paper/draft.tex
First revision

--- 2014-03-05 draft.tex
@@
 % \usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent
-
 %%% PACKAGES
@@
 }
+
+%\setlist[itemize]{leftmargin=*}
 \begin{document}
@@
 However, hypervisor is normally the most privileged task executed in the platform and therefore a curious or malicious administrator could easily retrieve those data \cite{cloudvisor:zhang}.
+Moreover, in multi-tenancy cloud platform, a user shares the same physical resources with other users.
+Several works \ref{sec:you,bot:att} showed the potential threat and risks of attacks issued by other users located on the same physical resource.
+
 On the other hand, from a user-centric point of view, it is difficult to observe and inspect deeply the functioning of the whole platform, leveraging commodity general purpose hypervisors (GPH).General Purpose Hypervisors are executed at the most privileged level and so, it is complicated to detect and intervene on them in presence of malicious attack or misbehavior, without service interruption for all VMs hosted on the server.
@@
 In the following paragraphs, those principles will represent a grid of analysis of different architectures.
-
-\paragraph{Inter-platform support} The cloud architecture has to be designed in order to support directly different platforms and service providers. 
-\paragraph{Fail-Safe modular architecture} In order to recover soft failures, the archtecure is supposed to be fail-safe, being able of automatically migrate instances or user-data in presence of a menace, but also reactivating the correct functioning in presence of a software failure.
+\begin{itemize}[]
+\item \textbf{Inter-platform support} The cloud architecture has to be designed in order to support directly different platforms and service providers.
+The same user virtual instances have to be executed ( and potentially migrated) unmodified over different platforms.
+\item \textbf{Fail-Safe modular architecture} In order to recover soft failures, the archtecure is supposed to be fail-safe, being able of automatically migrate instances or user-data in presence of a menace, but also reactivating the correct functioning in presence of a software failure.
 In order to satisfy this property, the amount of safety-critical components has to be as lowest as possible and the overall design has to be kept as modular as possible
-\paragraph{Small TCB} The amount of lines of code executed at the highest level of privilege is normally proportional to the number of potential failures and bugs in the platform.
+\item \textbf{Small TCB} The amount of lines of code executed at the highest level of privilege is normally proportional to the number of potential failures and bugs in the platform.
 According to this, an architecture presenting a tiny TCB will improve its safety and integrity properties by design.
-\paragraph{User-Data privacy} A curious or malicious administrator ( or another platform user) would not have the right to inspect and analyze user data and instances, without the explicit consensus with the user above mutually trusted services.
+\item \textbf{User-Data privacy} A curious or malicious administrator ( or another platform user) would not have the right to inspect and analyze user data and instances, without the explicit consensus with the user above mutually trusted services.
 User-data privacy is a major challenge of modern cloud platforms.
 It is important to identify a trade-off that allows providers to identify and intervene on possible menaces, but also guarantee the impossibility for a curious or malicious operator to access to them without restrictions.
-\paragraph{3rd party compatibility} Avoid disruption of existing 3rd part applications compatibility could represent a major principle to consider.
+In order to respect this principle when a platform administrator manages the platform it has not the highest level of privilege and each action on users object that he wants to deliver has to be validated by a mutually-trusted core, executed at the highest level of privilege.
+Moreover, this mutually-trust core has also to apply memory page cipher/uncipher in order to prevent memory page snooping when a page is swapped out.
+\item \textbf{3rd part compatibility} Avoid disruption of existing 3rd part applications compatibility could represent a major principle to consider.
 Platform interfaces has to be kept as close as possible to the existing one, preventing application rewrite.
-
+\end{itemize}
 Moreover, beside this functional design principles specifying which features have to be achieved by the platform, another set of non-design principles could be identified.
 Those principles does not concern directly features or aspects of the architecture of the platform, but the effective feasibility of the architecture.
-\paragraph{Minimize code addition} a major issue could be represented by the amount of code lines that have to be added and maintained to an existing architecture to meet the functional requirement.
+\begin{itemize}[]
+\item \textbf{Minimize code addition} a major issue could be represented by the amount of code lines that have to be added and maintained to an existing architecture to meet the functional requirement.
 Even it does not represent strictly a technical limitation to a certain proposal, it could become a ``show-stopper'' factor for its utilization.
-
+\end{itemize}
 In our analysis, it will be kept in consideration not only the overall suitability of the architecture, but also a quantification of the effort needed to develop and deploy it.
@@
 \midrule
-   DS1 & User-Data privacy & Vertical & \\
-   DS2 & Inter-platform support & Horizontal & \\
-   DS3 & Small TCB & Vertical & \\
-   DS4 & Fail-Safe modular architecture & Vertical & \\ 
-   DS5 & Backwards 3rd party compatibility & Horizontal \\
-   NDS1 & Minimize code addition & $\times$& \\
+   DP1 & User-Data privacy & Vertical & \\
+   DP2 & Fail-Safe modular architecture & Vertical & \\ 
+   DP3 & Small TCB & Vertical & \\
+   DP4 & Inter-platform support & Horizontal & \\
+   DP5 & Backwards 3rd part compatibility & Horizontal \\
+   NDP1 & Minimize code addition & & \\
 \bottomrule
@@
 To sum up, no architecture had been presented to address vertical and horizontal class of problems at the same time until now.
-XenBlanket does not investigate cloud architectural security concerns (DS3,DS4), limiting the focus on new potential feature of a coordinated multi-provider cloud.
-Cloudvisor proposed a security architecture that leverages nested virtualization without allowing the possibility of executing more than one nested hypervisor at time, that is a crucial requirements of a multi-provider cloud platform (DS2).
+XenBlanket does not investigate cloud architectural security concerns (DP3,DP2), limiting the focus on new potential feature of a coordinated multi-provider cloud.
+Cloudvisor proposed a security architecture that leverages nested virtualization without allowing the possibility of executing more than one nested hypervisor at time, that is a crucial requirements of a multi-provider cloud platform (DP4).
 In a SuperCloud scenario, each user has to run its own L1 hypervisor instance to enforce a compatibility layer, overcoming incompatibilities between different providers.
@@
 Self-service cloud approach (SSC) \cite{ssc:art} design a complete security architecture to overcome major security and privacy issues of currently deployed cloud infrastructures, in order to alleviate the different needing user-centric and provider-centric vision of cloud computing presented in \ref{par:intro}.
-Even if this architecture reproduces an approach similar to the SuperCloud without leveraging an extra layer of virtualization ( DS
+Even if this architecture reproduces an approach similar to the SuperCloud without leveraging an extra layer of virtualization (DP4), the approach changes the behavior where the user acts introducing new components that have to be directly managed by the user.
+Therefore, the main drawback of this solution is to change the interface with the user, preventing him to straightforwardly move to the new platform without having to change its applications (DP5).
 
@@
 Several works discuss and analyze the overall TCB of modern GPH \cite{nova, xmhf}, considering GPH TCB always bigger than 100 KLoC.
-This approach directly collides with one design principles introduced in the introduction paragraph (DS3), that identified a small TCB as a major requirement for a solid architecture.
+This approach directly collides with one design principles introduced in the introduction paragraph (DP3), that identified a small TCB as a major requirement for a solid architecture.
 %A brief comparison of TCB of widespread GPH is depicted in figure \ref{fig:tcbchart}. 
@@
 
-Finally, the above architecture does not improve consistently the overall stability of the infrastructure.
-The underlying core, L0, has the same problem of single-layer GPH architecture.
-Infact, the consistent TCB implies also that a failure experienced in one of the critical components will simply produce a generalized failure of the whole platform, not satysfing the DS4.
+Finally, the above architecture does not improve the overall stability of the infrastructure.
+The underlying core, L0, has the same problem of single-layer GPH architecture: the large TCB implies also that a failure experienced in one of the critical components will simply produce a generalized failure of the whole platform, not satysfing DP2.
+Moreover, in GPH standard architecture, the administrator has the entire control of the platform and could potentially snoop and inspect user VM status and data.
+
 This kind of architecture, in particular, does not leverage completely the possibility to hide behind a compatible layer, a really modular architecture with only a tiny critical core.
-
-To sum up, current nested virtualization limits are represented by a certain lack of dedicated feature in modern hypervisor and the impossibility to rely on a tiny TCB (DS3,DS4).
+To sum up, current nested virtualization limits are represented by a certain lack of dedicated feature in modern hypervisor and the impossibility to rely on a tiny TCB (DP3,DP4).
 The first one is being addressed by hypervisor developers. The second one could simply not be achieved with modern hypervisors.
@@
 They rely on the micro-kernel paradigm, extending it to virtualization.
-MH naturally satisfy vertical design principles (DS1,DS3, DS4).
+MH naturally satisfy vertical design principles (DP1,DP3, DP4).
 The tiny core could be considered the only system-critical componente of the architecture and if it possible to assure its correctness it could represent the last line of defense, where it could be possible to rebuild the architecture.
@@
 
-\subsection {MH open issues}
+\subsection {MH general issues}
 
@@
 Therefore, GPH exports wide and solid interfaces leveraged by a plenty of different software that is able to properly communicate and act through them on the virtualization layer.
-On the contrary, MH lack of cloud toolkit integration, requiring an explicit support to be written, not satysfing 3rd party compatibility requirements (DS5).
+On the contrary, MH lack of cloud toolkit integration, requiring an explicit support to be written, not satysfing 3rd part compatibility requirements (DP5).
 GPH are tightly connected with cloud toolkit and could present solid interfaces to communicate with different tools.
@@
 For instance, XMHF does not support multiple guests.
-This kind of drawbacks could represent important issues to handle and limit the potential interest in MH, even they are able to introduce several important features.
+This kind of drawbacks could limits the potential interest in MH, even they are able to introduce several important features.
 
@@
 At first, it is important to recall the most interesting achievable by such architectures.
-In particular, several remarkable features are live migration over different platform, transparent migration of virtualized clusters between different data-centers and improved isolation between customer instances and administration, improved control of the nested virtualization layer L1 (horizontal requirements DS.
-This approach allows us also to exploit different virtualization techniques.
-For instance, it could be possible to run at the same time PV and HVM nested instances, leveraging  Xen on KVM architecture ( if KVM nested VMX performs as expected).
-The choice of KVM as L0 is due to the general smaller TCB, compared with standard Xen (without stub domains) and this is important due to the high privilege level of L0.
+In particular, several remarkable features are live migration over different platform, transparent migration of virtualized clusters between different data-centers and improved isolation between customer instances and administration, improved control of the nested virtualization layer L1 (horizontal requirements DP1).
+Supposing the completeness and the correctness of nested virtualization hardware facilities in widespread hypervisor, GPH over GPH represents the best achievable architecture to respect horizontal design principles, but it lacks in vertical DPs as the size of TCB is the same of currently deployed architectures.
 
-Supposing the completeness and the correctness of nested virtualization hardware facilities in widespread hypervisor, GPH over GPH represents the best achievable architecture leveraging mature hypervisors.
 To sum up, it could be possible to achieve live-migration between different platform and other features related to VM management flexibility but it could not still be possible to reduce the TCB.
@@
 
-MH on MH in reported only for completeness. The interest on such architecture is limited because it will still present all cited issues of MH.
-Moreover, the lack of nested VMX functionalities prevents us to concretely test this kind of architecture, because MH normally rely on VMX facilities in order to create HVM guests.
-To conclude, nested virtualization is exploited in order to fill eco-systems gaps with GPH but MH over MH does not propose any improvement in this sense.
+MH on MH in reported only to see the orthogonality between this approach and GPH over GPH.
+A theoretical implementation of MH over MH will provide the vertical requirements that are missing in the previous approach.
+The tiny core could be able to overcome monolithic problems due to privilege level of adminitration tasks (DP1), the possibility to improve the resilience of the platform to a large set of software failures (DP2) and finally reduce the possibility of the vulnerabilities of the most privilege part of the software (DP3).
+However, the interest on such architecture is limited because it is not able to overcome any cited issues of single-layer MH.
+Moreover, several horizontal properties are theoretically achievable but they will require a huge effort in developing functions of interest.
+Moreover, the lack of nested VMX functionalities prevents to concretely test this kind of architecture, because MH normally rely on VMX facilities in order to create a nested HVM guests.
+To conclude, nested virtualization is exploited in order to fill eco-systems gaps with GPH but MH over MH does not provide any improvement in this sense.
 
@@
 
-The second architecture, proposed in order to reduce security concerns, is built combining an GPH at L0 and a MH at L1. 
+The third architecture, proposed in order to reduce security concerns, is built combining an GPH at L0 and a MH at L1. 
 %Several MH are proposed in literature focusing on different .
@@
 Existing GPH-based platforms could potentially exports nested VMX features to those guests.
-Even if this is not so usual in modern cloud platform, this lack is considered temporary and related only to the immaturity of those features.
-Secondly, the number of driver needed to properly support devices is significantly small.
-With a MH, the only I/O communication that the platform could be able to deliver, without the porting or addition of code, is Multi-level device assignment.
-This kind of approach is feasible only implementing in the L0 hypervisor software techniques like vIOMMU, that extends the IOMMU capabilities to several layers of virtualization.
+Even if this is not so usual in modern cloud platform, this lack is considered temporary and related only to the experimental status of those features.
 
-In addition, the amount of code running in L0 kernel space is still as big as before and, therefore, no TCB reduction. 
+However, the amount of code running in L0 kernel space is still as big as before and, therefore, no TCB reduction. 
 Finally, adopting a MH as L1 prevents to use inter-platform enhanced features.
@@
 \begin{table*}
+\label{fin:conf}
 \centering
-\begin{tabular}{llllll}
+\begin{tabular}{llcllll}
 \toprule
-Id & Functional design Principle & GPH over GPH & MH over MH & MH over GPH & GPH over MH\\
+Id & Functional design Principle & Class & GPH over GPH & MH over MH & MH over GPH & GPH over MH\\
 \midrule
-DS1 & User-Data privacy & $\times$  & $\times$ & $\times$ & $\times$ \\
-   DS2 & Inter-platform support & $\times$  & $\times$ & $\times$ & $\times$ \\
-   DS3 & Small TCB & $\times$  & $\times$ & $\times$ & $\times$ \\
-   DS4 & Fail-Safe modular architecture & $\times$  & $\times$ & $\times$ & $\times$ \\
-   DS5 & Backwards 3rd party compatibility & $\times$  & $\times$ & $\times$ & $\times$ \\
-   NDS1 & Minimize code addition & $\times$  & $\times$ & $\times$ & $\times$ \\
+DP1 & User-Data privacy & Vertical & $\times$  & $\checkmark$ & $\times $ & $\checkmark$ \\
+      DP2 & Fail-Safe modular architecture & Vertical & $\times$  & $\checkmark$ & $\times$ & $\checkmark$ \\
+   DP3 & Small TCB & Vertical &  $\times$& $\checkmark$ & $\times$ & $\checkmark$ \\
+   DP4 & Inter-platform support & Horizontal & $\checkmark$  & $\times$ & $\times$ & $\checkmark$ \\
+   DP5 & 3rd part compatibility & Horizontal & $\checkmark$  & $\times$ & $\times$ & $\checkmark$ \\
+   NDP1 & Minimize code addition & - & $\checkmark$  & $\times$ & $\times$ & $\checkmark$ \\
 \bottomrule
@@
 
-In this configuration, the amount of code that have to be trusted is lower than other architectures, due to the modular structure of micro-kernels., executed at L0.
-The possibility to exploit at L1 a GPH like Xen or KVM allows to deploy a similar architecture used in production today. The L1 GPH is normally integrated with modern cloud toolkits like OpenStack.
-Therefore, the architecture presents a stronger isolation leveraging on nested virtualization, providing in particular a strong tiny core.
+In this configuration, the TCB is smaller than other architectures, due to the modular structure of micro-kernels., executed at L0.
+Leveraging as L1 a GPH, like Xen or KVM, provides to the user an environment apparently  identical to the one used in production today, satisfying DP5.
+The L1 GPH is normally integrated with modern cloud toolkits like OpenStack and will support 3rd part applications.
+Therefore, the architecture presents a stronger isolation leveraging on nested virtualization, providing in particular a strong tiny core(DP3).
 The MH layer will be transparent from the user and it will not not any change in the functioning of the platform.
 
-The first class of possible drawbacks of this architecture is represented by the necessity to expose to the L1 nested hypervisor the nested VMX extension.
+The first class of possible drawbacks of this architecture is represented by the necessity to expose to the L1 nested hypervisor the nested VMX extension to let him ``nest'' another HVM guest.
 This is experimentally done by GPH like Xen and KVM, but it is generally not included in the essential features provided by a minimal micro-hypervisor.
@@
 To conclude, this section provides a short comparison between different platforms, contrasting architecture features above the two main different types of benefits given by nested virtualization, as reported widely in literature \cite{art:blan, nested:berghmans,  cloudvisor:zhang, turtle:ibm}.
-The first could be identified in the augmented degree of isolation the architecture could enforce between host OS and the "customer" point of access, leveraging a trusted core.
+The first could be identified in the augmented degree of isolation the architecture could enforce between host OS and the "customer" point of access, allowing to redesign the security architecture of the whole platform.
 The second one is represented by the possibility to leverage multiple virtualization layers to easily define inter-platform policy and operations between different platforms.
-In the above paragraphs, four different possible architectures were presented with the set of  substitute a GPH  with a MH, looking at the positive and negative new features.
-The starting point is GPH over GPH, a well analyzed approach in literature \cite{turtle:ibm, art:blan, }.
-We underlined features and limitations of this architecture in the first paragraphs and in the
-
+In table \ref{fin:conf}, it is possible a summary of the previous analysis.
 To conclude, GPH over MH is the most interesting one in order to enlarge nested virtualization promised benefit.
@@
 As stated in paragraph \ref{par:gom}, with L2 guest virtualized with the hardware assitance technique, the most popular one, the MH needs to export nVMX at least.
-On the contrary, adopting paravirtualization/DBT for L2 does not necessitate a particular hardware support and so it could not be necessary patch the MH to set up a platform. 
+On the contrary, adopting paravirtualization/Dynamic binary translation for L2 does not necessitate a particular hardware support and so it could not be necessary patch the MH to set up a platform. 
 However, it has not be underestimated, however, that para-virtualization supports only a subset of OS, and could not be considered a general solution and DBT has shown important overhead.
@@
 \section{Implementation}
-\subsection{Xen over NOVA}
-\paragraph{Inter-platform features}
-\paragraph{Tiny TCB}
-\paragraph{Leveraging existing APIs}
-\paragraph{Fail-safe components architecture}
-\subsection{}
+In the previous sections we proposed a functional analysis of cloud architecture in order to respect design principles that we identified in the introduction.
+In this section, we propose a GPH over MH implementation, based on Xen over NOVA.
+
+\begin{figure}
+\caption{Xen over NOVA architecture}
+\label{fig:io}
+\begin{center}
+\includegraphics[width=0.6\columnwidth]{article.png}
+\end{center}
+\end{figure}
+
+\begin{itemize}[]
+\item \textbf{Inter-platform features} Executed as L1, Xen would virtualize its nested guest initially using PV delivering the compatibility layer.
+\item \textbf{Tiny TCB} The NOVA kernel will represent the only part of the platform executed at the highest level of privilege.
+Even the VMM, Vancouver, will be executed in the L0 user-space.
+The micro-kernel in L0 enforces a strictly modular approach to L0.
+Therefore, the VMM will present distinct processes instances for each L1 guest, enforcing a context division.
+\item \textbf{3rd part compatibility} From a user point of view the architecture still exports the same interfaces, allowing the user to keep its applications unchanged.
+\item \textbf{Inter-platform support}
+\item \textbf{Fail-safe components architecture}  Moreover, the tiny core is the only safety-critical software component.
+The system could potentially be able to recover any crash in different user-space components.
+\end{itemize}
 
@@
 
-Future works will firstly concern the study of available MH, trying to identify which project could be suitable to act as L0 in a GPH over MH platform.
+%Future works will firstly concern the study of available MH, trying to identify which project could be suitable to act as L0 in a GPH over MH platform.
 As underlined previously, the key feature of L0 MH will be the capacity of having a tiny TCB.
@@
 Finally, another criteria for choosing the L0 hypervisor will be the evaluated effort needed for integrating a nested HV extension the to the base project.
-On the other hand, L1 choice is much simpler.
+%On the other hand, L1 choice is much simpler.
 Previous analysis identified Xen as the most suitable hypervisor for this role.
 The capacity to support different virtualization techniques added to the ability of running without HV support, makes Xen suitable to act as compatibility layer.
-
-After this initial study, the first step will be deploying a minimal nested architecture testing ``out-of-the-box'' features and limitations.
+%
+%After this initial study, the first step will be deploying a minimal nested architecture testing ``out-of-the-box'' features and limitations.
 The following part will concern the conception of patch and extensions in order to adapt the general design of MH to properly act as L0.
-The final goal is to provide an architecture able to leverage nested virtualization cloud platforms improvements, having a smaller TCB and allowing customers to leverage flexible migrations and scheduling.
+%The final goal is to provide an architecture able to leverage nested virtualization cloud platforms improvements, having a smaller TCB and allowing customers to leverage flexible migrations and scheduling.
 

### 2014-03-07 paper/draft.tex
First Revision

@@
 Moreover, in multi-tenancy cloud platform, a user shares the same physical resources with other users.
-Several works \ref{sec:you,bot:att} showed the potential threat and risks of attacks issued by other users located on the same physical resource.
-
-On the other hand, from a user-centric point of view, it is difficult to observe and inspect deeply the functioning of the whole platform, leveraging commodity general purpose hypervisors (GPH).General Purpose Hypervisors are executed at the most privileged level and so, it is complicated to detect and intervene on them in presence of malicious attack or misbehavior, without service interruption for all VMs hosted on the server.
+Several works \cite{sec:you, bot:att} showed the potential threat and risks of attacks issued by other users located on the same physical resource.
+On the other hand, from a user-centric point of view, it is difficult to observe and inspect deeply the functioning of the whole platform, leveraging commodity general purpose hypervisors (GPHs).GPHs are executed at the most privileged level and so, it is complicated to detect and intervene on them in presence of malicious attack or misbehavior, without service interruption for all VMs hosted on the server.
 To sum up, both visions collide with the execution of the hypervisor at the highest level of privileged.
 This attitude prevents the platform administrator from a real monitoring of the whole platform, but also the possibility for users to concretely inspects actions that platform administrators do on user data.
+A major challenge of new generation cloud platform is to alleviate this different tension, finding an acceptable trade-off.
 In the following paragraphs, we will make reference to this class of problem as ``Vertical'' requirements.
@@
 \begin{itemize}[]
-\item \textbf{Inter-platform support} The cloud architecture has to be designed in order to support directly different platforms and service providers.
-The same user virtual instances have to be executed ( and potentially migrated) unmodified over different platforms.
-\item \textbf{Fail-Safe modular architecture} In order to recover soft failures, the archtecure is supposed to be fail-safe, being able of automatically migrate instances or user-data in presence of a menace, but also reactivating the correct functioning in presence of a software failure.
-In order to satisfy this property, the amount of safety-critical components has to be as lowest as possible and the overall design has to be kept as modular as possible
-\item \textbf{Small TCB} The amount of lines of code executed at the highest level of privilege is normally proportional to the number of potential failures and bugs in the platform.
-According to this, an architecture presenting a tiny TCB will improve its safety and integrity properties by design.
-\item \textbf{User-Data privacy} A curious or malicious administrator ( or another platform user) would not have the right to inspect and analyze user data and instances, without the explicit consensus with the user above mutually trusted services.
+\item \textbf{DP1 User-Data privacy} A curious or malicious administrator ( or another platform user) would not have the right to inspect and analyze user data and instances, without the explicit consensus with the user above mutually trusted services.
 User-data privacy is a major challenge of modern cloud platforms.
@@
 Moreover, this mutually-trust core has also to apply memory page cipher/uncipher in order to prevent memory page snooping when a page is swapped out.
-\item \textbf{3rd part compatibility} Avoid disruption of existing 3rd part applications compatibility could represent a major principle to consider.
+\item \textbf{DP2 Fail-Safe modular architecture} In order to recover soft failures, the archtecure is supposed to be fail-safe, being able of automatically migrate instances or user-data in presence of a menace, but also reactivating the correct functioning in presence of a software failure.
+In order to satisfy this property, the amount of safety-critical components has to be as lowest as possible and the overall design has to be kept as modular as possible
+\item \textbf{DP3 Small TCB} The amount of lines of code executed at the highest level of privilege is normally proportional to the number of potential failures and bugs in the platform.
+According to this, an architecture presenting a tiny TCB will improve its safety and integrity properties by design.
+\item \textbf{DP4 Inter-platform support} The cloud architecture has to be designed in order to support directly different platforms and service providers.
+The same user virtual instances have to be executed ( and potentially migrated) unmodified over different platforms.
+\item \textbf{DP5 3rd part compatibility} Avoid disruption of existing 3rd part applications compatibility could represent a major principle to consider.
 Platform interfaces has to be kept as close as possible to the existing one, preventing application rewrite.
@@
 \begin{itemize}[]
-\item \textbf{Minimize code addition} a major issue could be represented by the amount of code lines that have to be added and maintained to an existing architecture to meet the functional requirement.
+\item \textbf{NDP1 Minimize code addition} a major issue could be represented by the amount of code lines that have to be added and maintained to an existing architecture to meet the functional requirement.
 Even it does not represent strictly a technical limitation to a certain proposal, it could become a ``show-stopper'' factor for its utilization.
@@
 % Intro + performance issues
+Several works presented in literature integrates these a subset of this design principles in their architecture, leveraging nested virtualization \cite{turtle:ibm,art:blan, cloudvisor:zhang}
 Cloud computing feature set could be enlarged exploiting two layers of virtualization instead of just one.
@@
 Even if this architecture reproduces an approach similar to the SuperCloud without leveraging an extra layer of virtualization (DP4), the approach changes the behavior where the user acts introducing new components that have to be directly managed by the user.
-Therefore, the main drawback of this solution is to change the interface with the user, preventing him to straightforwardly move to the new platform without having to change its applications (DP5).
+Therefore, the main drawback of this solution is to change the interface with the user, preventing the possibility to straightforwardly move to the new platform without having to change its applications (DP5).
 
@@
 it has to provide the widest possible support to different platforms and the capacity to virtualize adopting different techniques ( para-virtualization, hardware-assisted, dynamic binary translation).
-In this sense, Williams et al.
-propose a patched version of Xen, Xen Blanket, to accomplish a L1 virtualization layer, designed to run over different platforms.
+According to these requirements, Williams et al.
+\cite{art:blan} propose a patched version of Xen, Xen Blanket, to accomplish a L1 virtualization layer, designed to run over different platforms.
 This concretely allows to run over different physical platform the same VM, and live-migrate it at the occurrence.
@@
 Several works discuss and analyze the overall TCB of modern GPH \cite{nova, xmhf}, considering GPH TCB always bigger than 100 KLoC.
-This approach directly collides with one design principles introduced in the introduction paragraph (DP3), that identified a small TCB as a major requirement for a solid architecture.
+This approach directly collides with one design principles introduced in the introduction paragraph (DP3),
+%that identified a small TCB as a major requirement for a solid architecture.
 %A brief comparison of TCB of widespread GPH is depicted in figure \ref{fig:tcbchart}. 
@@
 Microkernels have the capacity of easily redesign the hierarchy of privileges.
-A user application has normally to trust the micro-kernel and the set of user-space software it uses.
+A user application has normally to trust the tiny micro-kernel and only the set of user-space software it uses.
 However, this set of trusted set of drivers and tools have not to be trusted by every application executed in the system.
-Each different user application could use its own set of tools, correctly registering and configuring them.
 Normally, this allows microkernels to implement different ``personalities''.
@@
 MH naturally satisfy vertical design principles (DP1,DP3, DP4).
-The tiny core could be considered the only system-critical componente of the architecture and if it possible to assure its correctness it could represent the last line of defense, where it could be possible to rebuild the architecture.
+The tiny core could be considered the only system-critical componente of the architecture and, due to its relative small dimension and the consequent minor attitude to vulnerabilities,it could represent the ``last line of defense''.
 
@@
 
-Finally, for sake of keeping the TCB small, MH could generally lack of several virtualization features, commonly available on modern GPH hypervisor.
+Finally, microkernel modularization introduce a major challenge in architecture design.
+Due the increased complexity, for sake of kernel size, MH could generally lack of several virtualization features, commonly available on modern GPH hypervisor.
 For instance, XMHF does not support multiple guests.
@@
 At first, it is important to recall the most interesting achievable by such architectures.
-In particular, several remarkable features are live migration over different platform, transparent migration of virtualized clusters between different data-centers and improved isolation between customer instances and administration, improved control of the nested virtualization layer L1 (horizontal requirements DP1).
-Supposing the completeness and the correctness of nested virtualization hardware facilities in widespread hypervisor, GPH over GPH represents the best achievable architecture to respect horizontal design principles, but it lacks in vertical DPs as the size of TCB is the same of currently deployed architectures.
+In particular, several remarkable features are live migration over different platform, transparent migration of virtualized clusters between different data-centers (horizontal requirements DP1).
+Supposing the completeness and the correctness of nested virtualization hardware facilities in widespread hypervisor, GPH over GPH satisfies all horizontal design principles, but it lacks in vertical DPs as the size of TCB is the same of currently deployed architectures.
 
@@
 The first impeding problem is the implementation of MH, able to propose to guest to virtualize with hardware assistance.
+In next section, we present an implementation of GPH over MH, that try to circumvent this limitation leveraging paravirtualization.
 
@@
 \begin{itemize}[]
-\item \textbf{Inter-platform features} Executed as L1, Xen would virtualize its nested guest initially using PV delivering the compatibility layer.
+\item \textbf{Inter-platform support} Executed as L1, Xen would virtualize its nested guest initially using PV delivering the compatibility layer.
 \item \textbf{Tiny TCB} The NOVA kernel will represent the only part of the platform executed at the highest level of privilege.
@@
 Therefore, the VMM will present distinct processes instances for each L1 guest, enforcing a context division.
-\item \textbf{3rd part compatibility} From a user point of view the architecture still exports the same interfaces, allowing the user to keep its applications unchanged.
-\item \textbf{Inter-platform support}
+\item \textbf{3rd part compatibility} From a user point of view the architecture still exports the same interfaces provided by the Xen hypervisor, allowing the user to keep its applications unchanged.
 \item \textbf{Fail-safe components architecture}  Moreover, the tiny core is the only safety-critical software component.

### 2014-03-07 paper/draft.tex
First Revision

@@
 As those functions are implemented outside the kernel, they are not unique but they are changeable and could coexist at the same time.
-This feature was seen with interest because of the possibility to move some virtualization-related from the kernelspace to user-space and to strictly separate piece of software executed for different users.
-This extraction process drives to a more solid platform, considering that user-space crashes are generally not fatal for the whole system.
+This feature was seen with interest because of the possibility to move some virtualization-related from the kernel space to user-space and to strictly separate piece of software executed for different users.
+This extraction process drives to a more solid platform, considering that user-space crashes are generally not fatal for the entire system integrity.
 A possible example, that will be analyzed in next section, is the user-space VMM component of Nova, Vancouver, that resides in user-space \cite{nova}.

### 2014-03-13 paper/draft.tex
Added announce for the article structure

@@
 \end{table}
+
+The reste of the paper in organized as follows.
+In Section 2 we start analyzing related work.
+In section 3, we analyze current limitations of classical nested virtualization approach and in section 4 we analyze the microvisor approach.
+In section 5 we analyzes cloud architectures propositions, combining nested virtualization and microvisors, leveraging the design principles grid introduced before.
+Furthermore, in section 6 we define a security architecture for multi-cloud scenario, with an analysis of necessary software components to satisfy vertical design principles.
+In section 8, we conclude the paper with future works in section  in section 6 we define a security architecture for multi-cloud scenario, with an analysis of necessary software components to satisfy vertical design principles.
+In section 8, we conclude the paper with future works in section 8.
 
@@
 In next section, we present an implementation of GPH over MH, that try to circumvent this limitation leveraging paravirtualization.
-
  \section{Implementation}

### 2014-03-13 paper/draft.tex
First commit...

@@
 
-\documentclass{acm_proc_article-sp} % use larger type; default would be 10pt
+\documentclass{sig-alternate}
 %\usepackage[margin={1.5cm,3cm}]{geometry}
@@
 %   read geometry.pdf for detailed page layout information
-\usepackage[hidelinks]{hyperref}
-\usepackage[autostyle,french=guillemets,babel]{csquotes}
+%\usepackage[hidelinks]{hyperref}
+%\usepackage[autostyle,french=guillemets,babel]{csquotes}
 \usepackage{graphicx} % support the \includegraphics command and options
-\usepackage[backend=biber]{biblatex}
-\addbibresource{biblio.bib}
+%\usepackage[backend=biber]{biblatex}
+%\addbibresource{biblio.bib}
 % \usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent
@@
 \usepackage{verbatim} % adds environment for commenting out blocks of text & for better verbatim
-\usepackage{subfig} % make it possible to include more than one captioned figure/table in a single float
+%\usepackage{subfig} % make it possible to include more than one captioned figure/table in a single float
 % These packages are all incorporated in the memoir class to one degree or another...
@@
 
-\title{A Sample {\ttlit ACM} SIG Proceedings Paper in LaTeX
-Format\titlenote{(Does NOT produce the permission block, copyright information nor page numbering). For use with ACM\_PROC\_ARTICLE-SP.CLS. Supported by ACM.}}
-\subtitle{[Extended Abstract]
-\titlenote{A full version of this paper is available as
-\textit{Author's Guide to Preparing ACM SIG Proceedings Using
-\LaTeX$2_\epsilon$\ and BibTeX} at
-\texttt{www.acm.org/eaddress.htm}}}
+\title{Towards Bridging the Gap... [TBD]}
+
 \author{
@@
 % 1st. author
-\numberofauthors{1}
+\numberofauthors{3}
 \alignauthor
-Ben Trovato\titlenote{Dr.~Trovato insisted his name be first.}\\
-       \affaddr{Institute for Clarity in Documentation}\\
-       \affaddr{1932 Wallamaloo Lane}\\
-       \affaddr{Wallamaloo, New Zealand}\\
-       \email{trovato@corporation.com}
-% 2nd. author
-%\date{} % Activate to display a given date or no date (if empty),
-         % otherwise the current date is printed 
+Alex Palesandro\\
+\affaddr{Orange Labs, France}\\
+\texttt{palexster@gmail.com}
+\alignauthor
+Aurélien Wailly\\
+\affaddr{Orange Labs, France}\\
+\texttt{aurelien.wailly@orange.com}
+\alignauthor
+Marc Lacoste\\
+\affaddr{Orange Labs, France}
+\texttt{marc.lacoste@orange.com}
 }
-
 %\setlist[itemize]{leftmargin=*}
@@
 %In this paper, we will analyze the state of nested virtualization in modern hypervisor and evaulating the impact of utilization of microhypervisor to enforce a stronger security level.
-
 }
@@
 \begin{figure}
-\caption{Xen over NOVA architecture}
-\label{fig:io}
 \begin{center}
@@
 \end{center}
+\caption{Xen over NOVA architecture}
+\label{fig:io}
 \end{figure}
@@
 %\section{Conclusion}
-\printbibliography
+%{\scriptsize
+%\bibliographystyle{abbrv}
+%\bibliography{biblio}
+%}
 \end{document}

### 2014-03-14 paper/draft.tex
Merge
Added Related Works paragraph

@@
 \usepackage{graphicx} % support the \includegraphics command and options
-%\usepackage[backend=biber]{biblatex}
-%\addbibresource{biblio.bib}
+\usepackage[backend=biber]{biblatex}
+\addbibresource{biblio.bib}
 % \usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent
@@
 
-\title{Towards Bridging the Gap... [TBD]}
+\title{}
 
@@
 \affaddr{Orange Labs, France}\\
-\texttt{palexster@gmail.com}
+\texttt{alex.palesandro@gmail.com}
 \alignauthor
@@
 \item \textbf{DP4 Inter-platform support} The cloud architecture has to be designed in order to support directly different platforms and service providers.
-The same user virtual instances have to be executed ( and potentially migrated) unmodified over different platforms.
+The same user virtual instances have to be executed (and potentially migrated) unmodified over different platforms.
 \item \textbf{DP5 3rd part compatibility} Avoid disruption of existing 3rd part applications compatibility could represent a major principle to consider.
@@
 Moreover, other works try to recover performance loss exploiting multi-level consolidation, introduced in nested virtualization \cite{art:blan}.
-In conclusion, nested virtualization slowdown could be considered at now an acceptable price to achieve advanced features and, in addition, hardware manifacturers are developing and proposing new hardware facilities to reduce this overhead \cite{vmcs:nakajima}.
+In conclusion, nested virtualization slowdown could be considered nowadays an acceptable price to achieve advanced features and, in addition, hardware manifacturers are developing and proposing new hardware facilities to reduce this overhead \cite{vmcs:nakajima}.
 
@@
 Therefore, the main drawback of this solution is to change the interface with the user, preventing the possibility to straightforwardly move to the new platform without having to change its applications (DP5).
+
+Furthermore, DeHype \cite{wu2013TamHosHypMosDepExe} presented an dissected version of KVM hypervisor, trying to reduce the TCB of this popular hypervisor.
+The approach consisted in demoting a great part of the hypervisor codebase to a user-level execution.
+DeHype satisfies all vertical design principles listed before (DS1,DS2,DS3), reinterpreting the microkernel approach of  demoting the greatest part of code to userspace.
+DeHype increases a lot the robustness of a cloud platform leveraging KVM, limiting for example the possibility of attack in multi-tenancy platform, but it does not address horizontal design principles about a multi-provider service.
 
@@
 Finally, the above architecture does not improve the overall stability of the infrastructure.
-The underlying core, L0, has the same problem of single-layer GPH architecture: the large TCB implies also that a failure experienced in one of the critical components will simply produce a generalized failure of the whole platform, not satysfing DP2.
+The underlying core, L0, has the same problem of single-layer GPH architecture: the large TCB implies also that a failure experienced in one of the critical components will simply produce a generalized failure of the whole platform, not satisfying DP2.
 Moreover, in GPH standard architecture, the administrator has the entire control of the platform and could potentially snoop and inspect user VM status and data.
@@
 Those ``personalities'' are concretely a set of ``processes'' handling specific functions running in user-space.
-As those functions are implemented outside the kernel, they are not unique but they are changeable and could coexist at the same time.
-This feature was seen with interest because of the possibility to move some virtualization-related from the kernel space to user-space and to strictly separate piece of software executed for different users.
-This extraction process drives to a more solid platform, considering that user-space crashes are generally not fatal for the entire system integrity.
+As those functions are implemented outside the kernel, each user could rely on a different instances of the hypervisor user-space process, that could coexist at the same time.
+This process separation between user could enforce a straightforward mechanism to neutralize VM-to-VM attack from a malicious user.
+Moreover, user-space processes are easier to debug with commodity tools (GDB, valgrind).
+Finally, this extraction process drives to a more solid platform, considering that user-space crashes are generally not fatal for the entire system integrity.
 A possible example, that will be analyzed in next section, is the user-space VMM component of Nova, Vancouver, that resides in user-space \cite{nova}.
+In addition, this approach is not far from the one proposed in DeHype \cite{wu2013TamHosHypMosDepExe}, where a consistent part of the KVM hypervisor has been demoted in user-space.
 
@@
 \section{Architecture comparison}
+
+\begin{figure*}
+\begin{center}
+\includegraphics[width=2\columnwidth]{globalcut.png}
+\end{center}
+\caption{}
+\label{fig:io}
+\end{figure*}
 
@@
 \subsection{GPH over MH}
+
 \label{par:gom}
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=0.6\columnwidth]{article.png}
-\end{center}
-\caption{Xen over NOVA architecture}
-\label{fig:io}
-\end{figure}
-
 \begin{itemize}[]

### 2014-03-17 paper/draft.tex
Added Related Works paragraph
Premier abstract
Extraction du problème, de la contrib et de l'implem
Il faut intégrer le commentaire au-dessus (notamment le trade-off qui
n'apparait pas).

@@
 
-\title{}
+\title{ Bridging [TBD]}
 
@@
 %In this paper, we will analyze the state of nested virtualization in modern hypervisor and evaulating the impact of utilization of microhypervisor to enforce a stronger security level.
+The problem of homogeneity over multiple IaaS providers is receiving considerable attention with the evolution toward nested virtualization.
+These new architectures unveil user-centric opportunities such as efficient resources management and VM migration.
+However the inherent drawbacks hinder its global adoption with a larger TCB and the complexity to monitor another virtualization layer.
+Thus, we argue that the promising architectural view of micro hypervisors approach is the key to the separation of hypervisor duties and reducing the attack surface.
+
+Furthermore, our architecture mushrooms commodity and micro hypervisors to leverage nested virtualization in terms of user-data privacy, modularity, TCB size and compatibility.
+These design principles are evaluated for each architecture configuration and compared to establish a reference security architecture.
+The resulting combination is built over common hypervisors empowered with micro hypervisors features.
+A reference implementation proven the architecture to be viable and will be extended to meet broader hypervisor scopes, enabling migration over heterogeneous cloud servers.
 }

### 2014-03-18 paper/draft.tex
Modification plan intro
- Petites fautes
- Ajout références \label{}

@@
 \label{ref:intro}
-Cloud computing platform are a solid reality nowadays. Modern cloud platforms are able to provide elastic use of resources, optimizing their allocation in order to meet the user needings.
+Cloud computing platform are a solid reality nowadays.
+Modern cloud platforms are able to provide elastic use of resources, optimizing their allocation in order to meet the user needings.
 However, currently deployed cloud platforms present still several limitation.
@@
 First, it is still difficult to deliver cloud applications which are able to transparently leverage over several platform resources.
-The concept of SuperCloud \cite{art:blan, xclo:blank} addresses this kind of problems.
+The concept of SuperCloud~\cite{art:blan, xclo:blank} addresses this kind of problems.
 Multi-platform coordination primarily for different standard and several format lock-in.
@@
 On one hand, from an user-centric point of view, the user would like to preserve privacy of its data when it stores data on the cloud platform.
-However, hypervisor is normally the most privileged task executed in the platform and therefore a curious or malicious administrator could easily retrieve those data \cite{cloudvisor:zhang}.
+However, hypervisor is normally the most privileged task executed in the platform and therefore a curious or malicious administrator could easily retrieve those data~\cite{cloudvisor:zhang}.
 Moreover, in multi-tenancy cloud platform, a user shares the same physical resources with other users.
-Several works \cite{sec:you, bot:att} showed the potential threat and risks of attacks issued by other users located on the same physical resource.
+Several works~\cite{sec:you, bot:att} showed the potential threat and risks of attacks issued by other users located on the same physical resource.
 On the other hand, from a user-centric point of view, it is difficult to observe and inspect deeply the functioning of the whole platform, leveraging commodity general purpose hypervisors (GPHs).GPHs are executed at the most privileged level and so, it is complicated to detect and intervene on them in presence of malicious attack or misbehavior, without service interruption for all VMs hosted on the server.
@@
 
-The reste of the paper in organized as follows.
-In Section 2 we start analyzing related work.
-In section 3, we analyze current limitations of classical nested virtualization approach and in section 4 we analyze the microvisor approach.
-In section 5 we analyzes cloud architectures propositions, combining nested virtualization and microvisors, leveraging the design principles grid introduced before.
-Furthermore, in section 6 we define a security architecture for multi-cloud scenario, with an analysis of necessary software components to satisfy vertical design principles.
-In section 8, we conclude the paper with future works in section  in section 6 we define a security architecture for multi-cloud scenario, with an analysis of necessary software components to satisfy vertical design principles.
+This paper is organized as following.
+In Section~\ref{sec:rw} we analyze the nested virtualization ecosystem and position existing solutions regarding our design principles.
+In Section~\ref{classic}, we details limitations of classical nested virtualization approach and in Section~\ref{sec:mh} we present the microvisor approach.
+In Section~\ref{sec:archcomp} we compare cloud architectures propositions, combining nested virtualization and microvisors, leveraging the design principles grid introduced before.
+%Furthermore, in Section 6 we define a security architecture for multi-cloud scenario, with an analysis of necessary software components to satisfy vertical design principles.
+In Section~\ref{sec:conclu}, we conclude the paper with future works and our vision of the SuperCloud.
+% in section in section 6 we define a security architecture for multi-cloud scenario, with an analysis of necessary software components to satisfy vertical design principles.
 In section 8, we conclude the paper with future works in section 8.
@@
 \section{Related Work}
+\label{sec:rw}
 % Intro + performance issues
@@
 \section{Classic nested virtualization approach}
+\label{sec:classic}
 
@@
 \section{Microhypervisor}
+\label{sec:mh}
 
@@
 \section{Architecture comparison}
+\label{sec:archcomp}
 
@@
 \section{Conclusion and future Works}
+\label{sec:conclu}
 

### 2014-03-18 paper/draft.tex
Relecture NDSS

@@
 
-Furthermore, DeHype \cite{wu2013TamHosHypMosDepExe} presented an dissected version of KVM hypervisor, trying to reduce the TCB of this popular hypervisor.
+% Integration NDSS Taming
+Furthermore, DeHype \cite{wu2013TamHosHypMosDepExe} presented a dissected version of KVM hypervisor, trying to reduce the TCB of this popular hypervisor.
 The approach consisted in demoting a great part of the hypervisor codebase to a user-level execution.

### 2014-03-19 paper/draft.tex
Added draft of conclusion

@@
 The approach consisted in demoting a great part of the hypervisor codebase to a user-level execution.
-DeHype satisfies all vertical design principles listed before (DS1,DS2,DS3), reinterpreting the microkernel approach of demoting the greatest part of code to userspace.
+DeHype satisfies all vertical design principles listed before (DS1,DS2,DS3), reinterpreting the microkernel approach defining  the greatest part of code to userspace.
 DeHype increases a lot the robustness of a cloud platform leveraging KVM, limiting for example the possibility of attack in multi-tenancy platform, but it does not address horizontal design principles about a multi-provider service.
@@
 The concept could be generalized to an arbitrary number of nested guest layers, introducing recursive virtualization.
+Virtualization is able to execute in a deprivileged context  that normally requires a privileged execution.
+With nested virtualization we could apply this approach even to virtualization itself, executing a deprivileged hypervisor and leveraging code reutilization.
 %More formally, Popek \& Golberg \cite{Popek:1974:FRV:361011.361073} discussed this subject, introducing the concept of ``recursively virtualizable'' machine and giving an initial definition of recursive virtual machines: If it is possible for a virtual machine to run under itself a copy of the VMM that could also exhibit all the properties of the VMM and this procedure could be repeated until the resources of the systems are consumed, then the original machine is recursively virtualizable.
@@
 However, virtualization introduces a computational overhead.
-It is small when we are dealing with single layer virtualization, but it grows exponentially when we try to add further levels \cite{rec:virt}.
+It is normally acceptable when we are dealing with single layer virtualization, but it grows exponentially when we try to add further levels \cite{rec:virt}.
 This huge performance overhead has prevented nested virtualization adoption during last years and had been finally addressed by Ben Yehuda et Al \cite{turtles:ibm} proposing an architecture with acceptable overhead.
@@
 
-\subsection{Conclusion}
+\subsection{Lessons learned}
 
@@
 In next section, we present an implementation of GPH over MH, that try to circumvent this limitation leveraging paravirtualization.
+\section{Security Architecture}
+\label{ref:arch}
+\section{Vertical properties}
+
 \section{Implementation}
 In the previous sections we proposed a functional analysis of cloud architecture in order to respect design principles that we identified in the introduction.
-In this section, we propose a GPH over MH implementation, based on Xen over NOVA.
+In this section, we propose a GPH over MH implementation, based on Xen over NOVA \cite{nova}.
 
@@
 
-%Future works will firstly concern the study of available MH, trying to identify which project could be suitable to act as L0 in a GPH over MH platform.
-As underlined previously, the key feature of L0 MH will be the capacity of having a tiny TCB.
-Moreover, additional requirements concerns the flexibility in terms VMs ( number of VM supported at the same time, CPU handling, I/O management).
-Finally, another criteria for choosing the L0 hypervisor will be the evaluated effort needed for integrating a nested HV extension the to the base project.
+In previous sections we provided a global analysis of modern challenges of cloud computing  and how those could be globally addressed addressed leveraging nested virtualization and different hypervisor architectures.
+
+Future works will be focused on two complementary parts.
+Firstly, they will concern the completion of the architecture of security in order to concretely meet horizontal design principles, identifying key components, similarly as done for the vertical properties.
+At the same time, a major effort will be directed on enlarging the initial implementation.
+
 %On the other hand, L1 choice is much simpler.

### 2014-03-19 paper/draft.tex
Updated bibliography.

--- 2014-03-18 paper/draft.tex
@@
-% !TEX TS-program = pdflatex
-% !TEX encoding = UTF-8 Unicode
-
-% This is a simple template for a LaTeX document using the "article" class.
-% See "book", "report", "letter" for other types of document.
-
 \documentclass{sig-alternate}
-%\usepackage[margin={1.5cm,3cm}]{geometry}
-%\setlength{\columnsep}{1cm}
-\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)
-
-%%% Examples of Article customizations
-% These packages are optional, depending whether you want the features they provide.
-% See the LaTeX Companion or other references for full information.
-
-%%% PAGE DIMENSIONS
-%\usepackage{geometry} % to change the page dimensions
-%\geometry{a4paper} % or letterpaper (US) or a5paper or....
-% \geometry{margin=2in} % for example, change the margins to 2 inches all round
-% \geometry{landscape} % set up the page for landscape
-%   read geometry.pdf for detailed page layout information
+\usepackage[utf8]{inputenc} 
 %\usepackage[hidelinks]{hyperref}
@@
 \usepackage{graphicx} % support the \includegraphics command and options
-\usepackage[backend=biber]{biblatex}
-\addbibresource{biblio.bib}
+%\usepackage[backend=biber]{biblatex}
+%\addbibresource{biblio.bib}
 % \usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent
-%%% PACKAGES
-\usepackage{booktabs} % for much better looking tables
-\usepackage{array} %
-% for better arrays (eg matrices) in maths
-\usepackage{paralist} % very flexible & customisable lists (eg. enumerate/itemize, etc.)
-\usepackage{verbatim} % adds environment for commenting out blocks of text & for better verbatim
-%\usepackage{subfig} % make it possible to include more than one captioned figure/table in a single float
-% These packages are all incorporated in the memoir class to one degree or another...
+\usepackage{booktabs}
+\usepackage{array} 
+\usepackage{paralist} 
+\usepackage{verbatim} 
 \usepackage{amsfonts}
+\usepackage{fancyhdr} 
+%\pagestyle{fancy} 
+%\renewcommand{\headrulewidth}{0pt} % customise the layout...
+%\lhead{}\chead{}\rhead{}
+%\lfoot{}\cfoot{\thepage}\rfoot{}
 
-%%% HEADERS & FOOTERS
-\usepackage{fancyhdr} % This should be set AFTER setting up the page geometry
-\pagestyle{fancy} % options: empty , plain , fancy
-\renewcommand{\headrulewidth}{0pt} % customise the layout...
-\lhead{}\chead{}\rhead{}
-\lfoot{}\cfoot{\thepage}\rfoot{}
-
-%%% SECTION TITLE APPEARANCE
-%\usepackage{sectsty}
-%\allsectionsfont{\sffamily\mdseries\upshape} % (See the fntguide.pdf for font help)
-%% (This matches ConTeXt defaults)
-
-%%% ToC (table of contents) APPEARANCE
-%\usepackage[nottoc,notlof,notlot]{tocbibind} % Put the bibliography in the ToC
-%\usepackage[titles,subfigure]{tocloft} % Alter the style of the Table of Contents
-%\renewcommand{\cftsecfont}{\rmfamily\mdseries\upshape}
-%\renewcommand{\cftsecpagefont}{\rmfamily\mdseries\upshape} % No bold!
-
-%%% END Article customizations
-
-%%% The "real" document content comes below...
-
-\title{ Bridging [TBD]}
+\title{Bridging [TBD]}
 
 \author{
-% You can go ahead and credit any number of authors here,
-% e.g. one 'row of three' or two rows (consisting of one row of three
-% and a second row of one, two or three).
-%
-% The command \alignauthor (no curly braces needed) should
-% precede each author name, affiliation/snail-mail address and
-% e-mail address. Additionally, tag each line of
-% affiliation/address with \affaddr, and tag the
-% e-mail address with \email.
-%
-% 1st. author
 \numberofauthors{3}
@@
 }
-%\setlist[itemize]{leftmargin=*}
+
 \begin{document}
@@
 \section{Introduction}
-\label{ref:intro}
+\label{sec:intro}
+
 Cloud computing platform are a solid reality nowadays.
@@
 In Section~\ref{sec:rw} we analyze the nested virtualization ecosystem and position existing solutions regarding our design principles.
-In Section~\ref{classic}, we details limitations of classical nested virtualization approach and in Section~\ref{sec:mh} we present the microvisor approach.
+In Section~\ref{sec:classic}, we details limitations of classical nested virtualization approach and in Section~\ref{sec:mh} we present the microvisor approach.
 In Section~\ref{sec:archcomp} we compare cloud architectures propositions, combining nested virtualization and microvisors, leveraging the design principles grid introduced before.
@@
 \label{sec:rw}
+
 % Intro + performance issues
@@
 Other approaches leveraging only one layer of virtualization does not achieve the whole design-principle satisfaction.
-Self-service cloud approach (SSC) \cite{ssc:art} design a complete security architecture to overcome major security and privacy issues of currently deployed cloud infrastructures, in order to alleviate the different needing user-centric and provider-centric vision of cloud computing presented in \ref{par:intro}.
+Self-service cloud approach (SSC) \cite{ssc:art} design a complete security architecture to overcome major security and privacy issues of currently deployed cloud infrastructures, in order to alleviate the different needing user-centric and provider-centric vision of cloud computing presented in \ref{sec:intro}.
 Even if this architecture reproduces an approach similar to the SuperCloud without leveraging an extra layer of virtualization (DP4), the approach changes the behavior where the user acts introducing new components that have to be directly managed by the user.
@@
 It is small when we are dealing with single layer virtualization, but it grows exponentially when we try to add further levels \cite{rec:virt}.
-This huge performance overhead has prevented nested virtualization adoption during last years and had been finally addressed by Ben Yehuda et Al \cite{turtles:ibm} proposing an architecture with acceptable overhead.
+This huge performance overhead has prevented nested virtualization adoption during last years and had been finally addressed by Ben Yehuda et Al \cite{turtle:ibm} proposing an architecture with acceptable overhead.
 
-Nested virtualization feasibility opened a plenty of new possibilities and perspectives, that have been largely discussed in literature \cite{rec:virt,turtles:ibm,art:blan}.
+Nested virtualization feasibility opened a plenty of new possibilities and perspectives, that have been largely discussed in literature \cite{rec:virt,turtle:ibm,art:blan}.
 The first aspect in common among different works proposed in literature is to diversify hypervisor functions, introducing new features but keeping the existing ones.
@@
 
-%\section{Conclusion}
-%{\scriptsize
-%\bibliographystyle{abbrv}
-%\bibliography{biblio}
-%}
+{\scriptsize
+\bibliographystyle{abbrv}
+\bibliography{biblio}
+}
 \end{document}

### 2014-03-19 paper/draft.tex

@@
 \usepackage{fancyhdr}
-%\pagestyle{fancy} 
+\usepackage{times}
+%\pagestyle{fancy} % options: empty , plain , fancy
 %\renewcommand{\headrulewidth}{0pt} % customise the layout...
@@
 
-\title{Bridging [TBD]}
+\title{ Bridging [TBD]}
 
@@
 Modern cloud platforms are able to provide elastic use of resources, optimizing their allocation in order to meet the user needings.
-However, currently deployed cloud platforms present still several limitation.
+However, currently deployed cloud platforms present still several limitations.
 
@@
 Several works~\cite{sec:you, bot:att} showed the potential threat and risks of attacks issued by other users located on the same physical resource.
-On the other hand, from a user-centric point of view, it is difficult to observe and inspect deeply the functioning of the whole platform, leveraging commodity general purpose hypervisors (GPHs).GPHs are executed at the most privileged level and so, it is complicated to detect and intervene on them in presence of malicious attack or misbehavior, without service interruption for all VMs hosted on the server.
+On the other hand, from a user-centric point of view, it is difficult to observe and inspect deeply the functioning of the whole platform, leveraging commodity general purpose hypervisors (GPHs).
+GPHs are executed at the most privileged level and so, it is complicated to detect and intervene on them in presence of malicious attack or misbehavior, without service interruption for all VMs hosted on the server.
 To sum up, both visions collide with the execution of the hypervisor at the highest level of privileged.
@@
 \begin{itemize}[]
-\item \textbf{DP1 User-Data privacy} A curious or malicious administrator ( or another platform user) would not have the right to inspect and analyze user data and instances, without the explicit consensus with the user above mutually trusted services.
+\item \textbf{DP1 User-Data privacy} A curious or malicious administrator (or another platform user) would not have the right to inspect and analyze user data and instances, without the explicit consensus with the user above mutually trusted services.
 User-data privacy is a major challenge of modern cloud platforms.
@@
 \item \textbf{DP2 Fail-Safe modular architecture} In order to recover soft failures, the archtecure is supposed to be fail-safe, being able of automatically migrate instances or user-data in presence of a menace, but also reactivating the correct functioning in presence of a software failure.
-In order to satisfy this property, the amount of safety-critical components has to be as lowest as possible and the overall design has to be kept as modular as possible
+In order to satisfy this property, the amount of safety-critical components has to be as lowest as possible and the overall design has to be kept as modular as possible.
 \item \textbf{DP3 Small TCB} The amount of lines of code executed at the highest level of privilege is normally proportional to the number of potential failures and bugs in the platform.
@@
 However, virtualization introduces a computational overhead.
-It is small when we are dealing with single layer virtualization, but it grows exponentially when we try to add further levels \cite{rec:virt}.
+It is normally acceptable when we are dealing with single layer virtualization, but it grows exponentially when we try to add further levels \cite{rec:virt}.
 This huge performance overhead has prevented nested virtualization adoption during last years and had been finally addressed by Ben Yehuda et Al \cite{turtle:ibm} proposing an architecture with acceptable overhead.
@@
 On the other hand, L1 is responsible for horizontal feature set.
-it has to provide the widest possible support to different platforms and the capacity to virtualize adopting different techniques ( para-virtualization, hardware-assisted, dynamic binary translation).
+It has to provide the widest possible support to different platforms and the capacity to virtualize adopting different techniques (para-virtualization, hardware-assisted, dynamic binary translation).
 According to these requirements, Williams et al.
@@
 This kind of architecture, in particular, does not leverage completely the possibility to hide behind a compatible layer, a really modular architecture with only a tiny critical core.
-To sum up, current nested virtualization limits are represented by a certain lack of dedicated feature in modern hypervisor and the impossibility to rely on a tiny TCB (DP3,DP4).
+To sum up, current nested virtualization limits are represented by a certain lack of dedicated feature in modern hypervisor and the impossibility to rely on a tiny TCB (DP3, DP4).
 The first one is being addressed by hypervisor developers. The second one could simply not be achieved with modern hypervisors.
@@
 A theoretical implementation of MH over MH will provide the vertical requirements that are missing in the previous approach.
-The tiny core could be able to overcome monolithic problems due to privilege level of adminitration tasks (DP1), the possibility to improve the resilience of the platform to a large set of software failures (DP2) and finally reduce the possibility of the vulnerabilities of the most privilege part of the software (DP3).
+The tiny core could be able to overcome monolithic problems due to privilege level of administration tasks (DP1), the possibility to improve the resilience of the platform to a large set of software failures (DP2) and finally reduce the possibility of the vulnerabilities of the most privilege part of the software (DP3).
 However, the interest on such architecture is limited because it is not able to overcome any cited issues of single-layer MH.
@@
 \subsection{GPH over MH}
-
 \label{par:gom}
@@
 In next section, we present an implementation of GPH over MH, that try to circumvent this limitation leveraging paravirtualization.
+
+\section{Security Architecture}
+\label{ref:arch}
+
+\section{Vertical properties}
+
 \section{Implementation}
 In the previous sections we proposed a functional analysis of cloud architecture in order to respect design principles that we identified in the introduction.
-In this section, we propose a GPH over MH implementation, based on Xen over NOVA.
+In this section, we propose a GPH over MH implementation, based on Xen over NOVA \cite{nova}.
 
@@
 
+%\section{Conclusion}
+%{\scriptsize
+%\bibliographystyle{abbrv}
+%\bibliography{biblio}
+%}
+
 {\scriptsize

### 2014-03-19 paper/draft.tex
Updated intro.

@@
 
-Cloud computing platform are a solid reality nowadays.
-Modern cloud platforms are able to provide elastic use of resources, optimizing their allocation in order to meet the user needings.
-However, currently deployed cloud platforms present still several limitations.
+\noindent Cloud computing is now a solid reality.
+Modern cloud platforms are able to provide elastic use of resources, optimizing their allocation to meet user needs.
+However, currently deployed cloud infrastructures present several limitations.
 
-First, it is still difficult to deliver cloud applications which are able to transparently leverage over several platform resources.
-The concept of SuperCloud~\cite{art:blan, xclo:blank} addresses this kind of problems.
-Multi-platform coordination primarily for different standard and several format lock-in.
-Unfortunately, those lock-in are not only due to market strategical decisions but they also reflects the virtualization technology heterogeneity.
-For example, each cloud provider could use a different storage system or a different format to store VMs images.
-Different hypervisors have nowadays a low degree of compatibility between them and this generally prevents leveraging on some popular features like `` VM live migration'' between different platform.
-Moreover, this incompatibility could also be enlarged if we consider that several platforms adopts one or more different virtualization techniques.
-In addition, this situation induce a tight dependency between the customer and the cloud provider and hardens the possibility to guarantee the availability of service in presence of a certain provider or to put in place some geographical optimization due to cost or traffic origin.
-The result was that some data server failures in the past has been catastrophic, causing a prolonged outage of service for a plenty of applications..
-In the remaining part of the work, we will refer to those issues as ``horizontal'' requirements, that shows straightforwardly their multi-platform nature.
+First, it is still difficult to deploy cloud applications able to transparently leverage resources of several heterogeneous platforms.
+The concept of super-cloud~\cite{art:blan, xclo:blank} tries notably to overcome such issues.
+Multi-platform coordination is hard, primarily due to lock-in to different standards such as VM formats.
+Unfortunately, such lock-in is not only due to market strategic decisions, but also reflects heterogeneity of virtualization technologies.
+For example, each cloud provider may use different storage systems or different formats to store VMs images.
+Different hypervisors have nowadays a low degree of compatibility between them.
+This generally prevents leveraging some popular features like VM live migration between different platforms.
+Incompatibility is even greater if we consider that each platform may adopt several types of virtualization techniques.
+This situation induces a tight dependency between the customer and the cloud provider.
+It may prevent guaranteeing service availability by relying on alternative providers, or to set up geographical optimizations due to cost or origin of traffic.
+This has resulted in some dramatic data server failures in recent years, causing prolonged outage of service for many applications.
+In what follows, we refer to this class of challenges as \textit{horizontal}, to highlight their multi-platform nature.
 

### 2014-03-19 paper/draft.tex
Updated intro.

@@
 
-\noindent Cloud computing is now a solid reality.
+\noindent Distributed cloud computing is now becoming a reality.
 Modern cloud platforms are able to provide elastic use of resources, optimizing their allocation to meet user needs.
-However, currently deployed cloud infrastructures present several limitations.
+However, currently deployed distributed cloud infrastructures present several limitations.
 
-First, it is still difficult to deploy cloud applications able to transparently leverage resources of several heterogeneous platforms.
+First, deploying cloud applications able to transparently leverage resources of several heterogeneous platforms is still difficult.
 The concept of super-cloud~\cite{art:blan, xclo:blank} tries notably to overcome such issues.
@@
 This has resulted in some dramatic data server failures in recent years, causing prolonged outage of service for many applications.
-In what follows, we refer to this class of challenges as \textit{horizontal}, to highlight their multi-platform nature.
+In what follows, we refer to this class of multi-platform issues as \textit{horizontal} challenges.
 
-The second class of issues concerns platform integrity and user data privacy.
-On one hand, from an user-centric point of view, the user would like to preserve privacy of its data when it stores data on the cloud platform.
-However, hypervisor is normally the most privileged task executed in the platform and therefore a curious or malicious administrator could easily retrieve those data~\cite{cloudvisor:zhang}.
-Moreover, in multi-tenancy cloud platform, a user shares the same physical resources with other users.
-Several works~\cite{sec:you, bot:att} showed the potential threat and risks of attacks issued by other users located on the same physical resource.
-On the other hand, from a user-centric point of view, it is difficult to observe and inspect deeply the functioning of the whole platform, leveraging commodity general purpose hypervisors (GPHs).
-GPHs are executed at the most privileged level and so, it is complicated to detect and intervene on them in presence of malicious attack or misbehavior, without service interruption for all VMs hosted on the server.
-To sum up, both visions collide with the execution of the hypervisor at the highest level of privileged.
-This attitude prevents the platform administrator from a real monitoring of the whole platform, but also the possibility for users to concretely inspects actions that platform administrators do on user data.
-A major challenge of new generation cloud platform is to alleviate this different tension, finding an acceptable trade-off.
-In the following paragraphs, we will make reference to this class of problem as ``Vertical'' requirements.
+Second, a number of issues have been raised about user data privacy and platform integrity.
+From the user perspective, users would like to preserve privacy of their data stored on the cloud platform.
+However, due to overprivileged hypervisors, a curious or malicious administrator could easily retrieve such data~\cite{cloudvisor:zhang}.
+In a multi-tenant cloud platform, physical resource sharing between users may lead not only to data leakage or but also attacks against the IaaS~\cite{bot:att, sec:you}.
+From the provider perspective, a key objective is to thus guarantee platform integrity against rogue VMs.
+However, commodity general-purpose hypervisors (GPHs) make it difficult to observe deeply all layers of a running IaaS platform.
+Indeed, running GPHs at the highest level of privilege hampers detection and reaction on the hypervisor itself in presence of faulty or malicious behavior -- unless to cause interruption of service for all hosted VMs.
+Thus, both user-centric and provider-centric visions of the cloud have a problem with overprivileged hypervisors.
+%: the platform administrator cannot perform in-depth infrastructure monitoring nor can users concretely inspect administrator actions on their data. 
+A major challenge for new generations of cloud infrastructures is thus to find acceptable trade-offs between such conflicting requirements.
+In what follows, we refer to this class of layer-oriented issues as \textit{vertical} challenges.
 
@@
 
-\section{Classic nested virtualization approach}
+\section{Tradidional Nested Virtualization}
 \label{sec:classic}

### 2014-03-20 paper/draft.tex
Updated title.
PV-FV clarification

@@
 
-\title{ Bridging [TBD]}
+\title{From Nested Virtualization Architecture \\To 2D Security Management for Distributed Clouds}
 
@@
 
-\section{Tradidional Nested Virtualization}
+\section{Current Nested Virtualization}
 \label{sec:classic}
@@
 The concept could be generalized to an arbitrary number of nested guest layers, introducing recursive virtualization.
-Virtualization is able to execute in a deprivileged context  that normally requires a privileged execution.
-With nested virtualization we could apply this approach even to virtualization itself, executing a deprivileged hypervisor and leveraging code reutilization.
+Virtualization is able to execute tasks in a deprivileged context that normally requires a privileged execution.
+With nested virtualization we could extend this approach to virtualization itself, executing a guest hypervisor with nested guest VMs, without requiring any change to any of them.
 %More formally, Popek \& Golberg \cite{Popek:1974:FRV:361011.361073} discussed this subject, introducing the concept of ``recursively virtualizable'' machine and giving an initial definition of recursive virtual machines: If it is possible for a virtual machine to run under itself a copy of the VMM that could also exhibit all the properties of the VMM and this procedure could be repeated until the resources of the systems are consumed, then the original machine is recursively virtualizable.
@@
 
-\subsection{Classical nested virtualization limits}
+\subsection{Limits}
 
@@
 Both Xen and KVM propose nested VMX/SVN features, that consists in the possibility to efficiently handle nested hardware virtualization, in presence of a single-layer hardware facilities.
+Full Virtualization is currently the only virtualization technique supported for L1 virtualization, due to its capacity of running unmodified guest OS and therefore hypervisors.
+On the contrary, PV would require an ad hoc modified hypervisor to act as L1 (not satisfying NDP1).
 %To address this performance problems, several hardware producers proposed new facilities, like Intel VMCS shadowing, that reduces the amount of nested VMEXIT when handling a nested guest VMEXIT.
@@
 
-\section{Microhypervisor}
+\section{Micro-hypervisors}
 \label{sec:mh}
@@
 
-\subsection {MH general issues}
+\subsection {MH General Issues}
 
@@
 
-\section{Architecture comparison}
+\section{Architecture Comparison}
 \label{sec:archcomp}
@@
 
-\subsection{GPH over GPH}
+\subsection{GPH Over GPH}
 
@@
 
-\subsection{MH over MH}
+\subsection{MH Over MH}
 
@@
 
-\subsection{MH over GPH}
+\subsection{MH Over GPH}
 \label{par:mog}
@@
 
-\subsection{GPH over MH}
+\subsection{GPH Over MH}
 \label{par:gom}
@@
 
-\subsection{Lessons learned}
+\subsection{Lessons Learned}
 
@@
 
-\section{Vertical properties}
+\section{Vertical Properties}
 
@@
 
-\section{Conclusion and future Works}
+\section{Conclusion and Future Work}
 \label{sec:conclu}

### 2014-03-20 paper/draft.tex
Updated intro.

@@
 
-To resume, in order to present a complete view of the architecture, we identified five key design principles, that a cloud computing platform has to implement in order to solve the cited problems.
-In the following paragraphs, those principles will represent a grid of analysis of different architectures.
+We identified five key design principles that a IaaS architecture should implement to meet the previous challenges.
+In what follows, those principles will serve as reading template to analyze different architectures.
 \begin{itemize}[]
-\item \textbf{DP1 User-Data privacy} A curious or malicious administrator (or another platform user) would not have the right to inspect and analyze user data and instances, without the explicit consensus with the user above mutually trusted services.
-User-data privacy is a major challenge of modern cloud platforms.
-It is important to identify a trade-off that allows providers to identify and intervene on possible menaces, but also guarantee the impossibility for a curious or malicious operator to access to them without restrictions.
-In order to respect this principle when a platform administrator manages the platform it has not the highest level of privilege and each action on users object that he wants to deliver has to be validated by a mutually-trusted core, executed at the highest level of privilege.
-Moreover, this mutually-trust core has also to apply memory page cipher/uncipher in order to prevent memory page snooping when a page is swapped out.
-\item \textbf{DP2 Fail-Safe modular architecture} In order to recover soft failures, the archtecure is supposed to be fail-safe, being able of automatically migrate instances or user-data in presence of a menace, but also reactivating the correct functioning in presence of a software failure.
+\item \textbf{DP1 User-Data Privacy} A curious or malicious administrator (or another platform user) should not be authorized to inspect and analyze user data and VM instances without explicit user consent relying on mutually trusted services.
+A trade-off should be found to allow providers to detect and react to potential threats, but also to prevent unlimited platform and user data access in such inappropriate operator behaviors.
+Thus, a platform administrator should not manage the infrastructure directly using the highest level of privilege.
+Instead, his actions on user data objects should be validated by a mutually-trusted kernel run in the most privileged mode.
+This kernel should also enforce memory page encryption to prevent snooping when pages are swapped out.
+\item \textbf{DP2 Fail-Safe Modular Architecture} In order to recover soft failures, the archtecure is supposed to be fail-safe, being able of automatically migrate instances or user-data in presence of a menace, but also reactivating the correct functioning in presence of a software failure.
 In order to satisfy this property, the amount of safety-critical components has to be as lowest as possible and the overall design has to be kept as modular as possible.
@@
 According to this, an architecture presenting a tiny TCB will improve its safety and integrity properties by design.
-\item \textbf{DP4 Inter-platform support} The cloud architecture has to be designed in order to support directly different platforms and service providers.
+\item \textbf{DP4 Inter-Platform Support} The cloud architecture has to be designed in order to support directly different platforms and service providers.
 The same user virtual instances have to be executed (and potentially migrated) unmodified over different platforms.
-\item \textbf{DP5 3rd part compatibility} Avoid disruption of existing 3rd part applications compatibility could represent a major principle to consider.
+\item \textbf{DP5 Compatibility with Legacy} Avoid disruption of existing 3rd part applications compatibility could represent a major principle to consider.
 Platform interfaces has to be kept as close as possible to the existing one, preventing application rewrite.
@@
 The concept could be generalized to an arbitrary number of nested guest layers, introducing recursive virtualization.
-Virtualization is able to execute tasks in a deprivileged context that normally requires a privileged execution.
-With nested virtualization we could extend this approach to virtualization itself, executing a guest hypervisor with nested guest VMs, without requiring any change to any of them.
+Virtualization is able to execute in a deprivileged context  that normally requires a privileged execution.
+With nested virtualization we could apply this approach even to virtualization itself, executing a deprivileged hypervisor and leveraging code reutilization.
 %More formally, Popek \& Golberg \cite{Popek:1974:FRV:361011.361073} discussed this subject, introducing the concept of ``recursively virtualizable'' machine and giving an initial definition of recursive virtual machines: If it is possible for a virtual machine to run under itself a copy of the VMM that could also exhibit all the properties of the VMM and this procedure could be repeated until the resources of the systems are consumed, then the original machine is recursively virtualizable.
@@
 Both Xen and KVM propose nested VMX/SVN features, that consists in the possibility to efficiently handle nested hardware virtualization, in presence of a single-layer hardware facilities.
-Full Virtualization is currently the only virtualization technique supported for L1 virtualization, due to its capacity of running unmodified guest OS and therefore hypervisors.
-On the contrary, PV would require an ad hoc modified hypervisor to act as L1 (not satisfying NDP1).
 %To address this performance problems, several hardware producers proposed new facilities, like Intel VMCS shadowing, that reduces the amount of nested VMEXIT when handling a nested guest VMEXIT.

### 2014-03-21 paper/draft.tex
Updated requirements.

@@
 
+This paper is organized as follows.
+In Section~\ref{sec:rw} we analyze the nested virtualization ecosystem and position existing solutions regarding our design principles.
+In Section~\ref{sec:classic}, we details limitations of classical nested virtualization approach and in Section~\ref{sec:mh} we present the microvisor approach.
+In Section~\ref{sec:archcomp} we compare cloud architectures propositions, combining nested virtualization and microvisors, leveraging the design principles grid introduced before.
+%Furthermore, in Section 6 we define a security architecture for multi-cloud scenario, with an analysis of necessary software components to satisfy vertical design principles.
+In Section~\ref{sec:conclu}, we conclude the paper with future works and our vision of the SuperCloud.
+% in section in section 6 we define a security architecture for multi-cloud scenario, with an analysis of necessary software components to satisfy vertical design principles.
+In section 8, we conclude the paper with future works in section 8.
+
+\section{Challenges and Design Principles}
+
 First, deploying cloud applications able to transparently leverage resources of several heterogeneous platforms is still difficult.
@@
 
-We identified five key design principles that a IaaS architecture should implement to meet the previous challenges.
+We identified five key design principles (DP) that a IaaS architecture should implement to meet the previous challenges.
 In what follows, those principles will serve as reading template to analyze different architectures.
 \begin{itemize}[]
-\item \textbf{DP1 User-Data Privacy} A curious or malicious administrator (or another platform user) should not be authorized to inspect and analyze user data and VM instances without explicit user consent relying on mutually trusted services.
+\item \textbf{DP1 User Data Privacy} A curious or malicious administrator (or another platform user) should not be authorized to inspect and analyze user data and VM instances without explicit user consent relying on mutually trusted services.
 A trade-off should be found to allow providers to detect and react to potential threats, but also to prevent unlimited platform and user data access in such inappropriate operator behaviors.
@@
 This kernel should also enforce memory page encryption to prevent snooping when pages are swapped out.
-\item \textbf{DP2 Fail-Safe Modular Architecture} In order to recover soft failures, the archtecure is supposed to be fail-safe, being able of automatically migrate instances or user-data in presence of a menace, but also reactivating the correct functioning in presence of a software failure.
-In order to satisfy this property, the amount of safety-critical components has to be as lowest as possible and the overall design has to be kept as modular as possible.
-\item \textbf{DP3 Small TCB} The amount of lines of code executed at the highest level of privilege is normally proportional to the number of potential failures and bugs in the platform.
-According to this, an architecture presenting a tiny TCB will improve its safety and integrity properties by design.
-\item \textbf{DP4 Inter-Platform Support} The cloud architecture has to be designed in order to support directly different platforms and service providers.
-The same user virtual instances have to be executed (and potentially migrated) unmodified over different platforms.
-\item \textbf{DP5 Compatibility with Legacy} Avoid disruption of existing 3rd part applications compatibility could represent a major principle to consider.
-Platform interfaces has to be kept as close as possible to the existing one, preventing application rewrite.
+\item \textbf{DP2 Fail-Safe Modular Architecture} The architecture should be fail-safe: for instance, to enable automatically migration of instances or user-data in presence of a threat and to recover from software failures.
+To reach this goal, the amount of safety-critical components should be as low as possible to mimize the attack surface.
+Similarly, the overall design should be as modular as possible to enable platform reconfiguration.
+\item \textbf{DP3 Small TCB} The number of potential failures and bugs in the platform is directly linked with the code size executed at the highest level of privilege.
+Thus, an architecture with a tiny TCB (Trusted Computing Base) will improve safety and integrity by design.
+\item \textbf{DP4 Inter-Platform Support} The cloud architecture should allow supporting directly different platforms and service providers.
+The same user virtual instances should be allowed to be run (and potentially migrated) unmodified over different IaaS platforms.
+\item \textbf{DP5 Compatibility with Legacy} Evolution in the IaaS architecture should avoiding disrupting existing third party applications.
+Thus, platform interfaces should be kept as close as possible to existing ones, to prevent code rewriting notably for applications.
 \end{itemize}
-Moreover, beside this functional design principles specifying which features have to be achieved by the platform, another set of non-design principles could be identified.
-Those principles does not concern directly features or aspects of the architecture of the platform, but the effective feasibility of the architecture.
+In addition to the above functional design principles that specify which features have to be achieved by the platform architecture, another set of non-design principles (NDP) should be considered regarding effective feasibility of implementing the architecture.
+We just give an example in terms of platform usability:
 \begin{itemize}[]
-\item \textbf{NDP1 Minimize code addition} a major issue could be represented by the amount of code lines that have to be added and maintained to an existing architecture to meet the functional requirement.
-Even it does not represent strictly a technical limitation to a certain proposal, it could become a ``show-stopper'' factor for its utilization.
+\item \textbf{NDP1 Minimal Additional Code} A major issue could be represented by the amount of code to be added and maintained to an existing architecture to meet functional requirements.
+Despite not being a strictly a technical limitation of the architecture, it could become a show-stopper to its practical use.
 \end{itemize}
-In our analysis, it will be kept in consideration not only the overall suitability of the architecture, but also a quantification of the effort needed to develop and deploy it.
-In table \ref{int:des} it is possible to observe a resume of previous cited principles.
+In our analysis, we will consider not only the overall suitability of the architecture, but also the amount of effort needed to develop and deploy it.
+Table \ref{int:des} summarizes the previous principles and which class of requirements they address.
 
 \begin{table}
-\label{int:des}
-\begin{tabular}{llll}
+\caption{Design Principles for a Distributed IaaS Platform.\label{int:des}}
+\begin{tabular}{lll}
 \toprule
-Id & Functional design Principle & Class &\\
+\textbf{Id} & \textbf{Design Principle} & \textbf{Class of Requirement} \\
 \midrule
-DP1 & User-Data privacy & Vertical & \\
-DP2 & Fail-Safe modular architecture & Vertical & \\ 
-DP3 & Small TCB & Vertical & \\
-DP4 & Inter-platform support & Horizontal & \\
-DP5 & Backwards 3rd part compatibility & Horizontal \\
-NDP1 & Minimize code addition & & \\
+DP1 & User Data Privacy & Vertical \\
+DP2 & Fail-Safe Modular Architecture & Vertical \\ 
+DP3 & Small TCB & Vertical \\
+DP4 & Inter-Platform Support & Horizontal \\
+DP5 & Compatibility with Legacy & Horizontal \\
+NDP1 & Minimal Additional Code & -- \\
 \bottomrule
@@
 \end{table}
-
-This paper is organized as following.
-In Section~\ref{sec:rw} we analyze the nested virtualization ecosystem and position existing solutions regarding our design principles.
-In Section~\ref{sec:classic}, we details limitations of classical nested virtualization approach and in Section~\ref{sec:mh} we present the microvisor approach.
-In Section~\ref{sec:archcomp} we compare cloud architectures propositions, combining nested virtualization and microvisors, leveraging the design principles grid introduced before.
-%Furthermore, in Section 6 we define a security architecture for multi-cloud scenario, with an analysis of necessary software components to satisfy vertical design principles.
-In Section~\ref{sec:conclu}, we conclude the paper with future works and our vision of the SuperCloud.
-% in section in section 6 we define a security architecture for multi-cloud scenario, with an analysis of necessary software components to satisfy vertical design principles.
-In section 8, we conclude the paper with future works in section 8.
 

### 2014-03-23 paper/draft.tex
Updating nv section

@@
 
-Nested virtualization refers to a system architecture with two layers of virtualization.
-It consists in a ``virtualized'' virtualization, where the guest system virtualize a nested guest.
-The concept could be generalized to an arbitrary number of nested guest layers, introducing recursive virtualization.
-Virtualization is able to execute tasks in a deprivileged context that normally requires a privileged execution.
-With nested virtualization we could extend this approach to virtualization itself, executing a guest hypervisor with nested guest VMs, without requiring any change to any of them.
+\subsection{Principle}
+
+\noindent \textit{Nested virtualization (NV)} refers to a system architecture with two layers of virtualization~\cite{turtle:ibm}.
+It consists in ``virtualized'' virtualization: the guest system virtualizes a nested guest.
+The concept may be generalized to an arbitrary number of nested guest layers, leading to recursive virtualization~\cite{rec:virt,Popek:1974:FRV:361011.361073}.
+(Full) virtualization is able to execute tasks in a deprivileged context that normally requires a privileged execution.
+With NV, this approach may be extended to virtualization itself: a guest hypervisor may be run with nested guest VMs without requiring any change to any of them.
 %More formally, Popek \& Golberg \cite{Popek:1974:FRV:361011.361073} discussed this subject, introducing the concept of ``recursively virtualizable'' machine and giving an initial definition of recursive virtual machines: If it is possible for a virtual machine to run under itself a copy of the VMM that could also exhibit all the properties of the VMM and this procedure could be repeated until the resources of the systems are consumed, then the original machine is recursively virtualizable.
 
-Introducing nested virtualization, we have to identify the two main actors involved in the virtualization process: bare-metal hypervisor (L0) and nested hypervisor (L1).
-In addition, we will refer to the nested guest as L2 guest.
-However, virtualization introduces a computational overhead.
-It is normally acceptable when we are dealing with single layer virtualization, but it grows exponentially when we try to add further levels \cite{rec:virt}.
-This huge performance overhead has prevented nested virtualization adoption during last years and had been finally addressed by Ben Yehuda et Al \cite{turtle:ibm} proposing an architecture with acceptable overhead.
+Two layers should be distinguished when migrating to an NV architecture: (1) the hypervisor running above the hardware (\textit{L0 hypervisor}); and (2) the nested hypervisor (\textit{L1 hypervisor}).
+Nested guests are generally referred to as \textit{L2 guests}.
+Unfortunately, virtualization comes with a price due to the extra layer of indirection introduced: the computational overhead is generally acceptable when dealing with single-layer virtualization, but grows exponentially when adding further levels~\cite{rec:virt}.
+This huge performance overhead prevented NV adoption during many years.
+However, Ben Yehuda et al.
+finally proposed an architecture tending towards more reasonable overheads~\cite{turtle:ibm}.
 
-Nested virtualization feasibility opened a plenty of new possibilities and perspectives, that have been largely discussed in literature \cite{rec:virt,turtle:ibm,art:blan}.
-The first aspect in common among different works proposed in literature is to diversify hypervisor functions, introducing new features but keeping the existing ones.
-In a general analysis of existing works, it could be possible to identify the general role of each different layer.
-On one hand, L0 is charged of the vertical aspects of the platform.
-It has to achieve  the best possible performances and enforcing the strongest possible level of isolation.
-L0 represents the last line of defense of the platform and a privileged point of global status observation.
+NV feasibility opened plenty of new possibilities and avenues for research that have been largely discussed in the literature~\cite{turtle:ibm,rec:virt,art:blan}.
+Such works all tend to diversify hypervisor functionalities, introducing new features but also keeping existing ones.
+It turns out that each layer addresses clearly different sets of issues.
 
-On the other hand, L1 is responsible for horizontal feature set.
-It has to provide the widest possible support to different platforms and the capacity to virtualize adopting different techniques (para-virtualization, hardware-assisted, dynamic binary translation).
-According to these requirements, Williams et al.
-\cite{art:blan} propose a patched version of Xen, Xen Blanket, to accomplish a L1 virtualization layer, designed to run over different platforms.
-This concretely allows to run over different physical platform the same VM, and live-migrate it at the occurrence.
+L0 tackles vertical aspects of the platform.
+It aims to guarantee the best performance and strongest level of isolation possible.
+L0 represents the last line of defense of the platform and a privileged point for monitoring of its global status.
+
+L1 is responsible for horizontal feature set.
+It has to provide the widest possible support for different platforms and the ability to virtualize using different techniques e.g., para-virtualization, hardware-assisted virtualization, dynamic binary translation.
+{\sffamily To address such requirements, Williams et al.~\cite{art:blan} proposed a patched version of Xen, Xen Blanket, to realize an L1 virtualization layer designed to run over different hypervisors.
+This notably enables to run and to perform live migration of the same VM over different physical cloud infrastructures.} % A migrer dans le related work?
 

### 2014-03-23 paper/draft.tex
Updated section on NV.

@@
 
-\subsection{Limits}
+\subsection{Limitations}
 
-Nowadays, nested virtualization presents several limits related to the hardware support and software state of implementation.
-At first, to provide nested virtualization, hypervisors has to provide an exact copy of hardware virtualization facilities.
-The most exploited technique, full virtualization with hardware assistance (HV), relies on a quite complex dedicated support.
-This set of facilities assists instruction emulation, memory address translation and device management.
-Both Xen and KVM propose nested VMX/SVN features, that consists in the possibility to efficiently handle nested hardware virtualization, in presence of a single-layer hardware facilities.
-Full Virtualization is currently the only virtualization technique supported for L1 virtualization, due to its capacity of running unmodified guest OS and therefore hypervisors.
-On the contrary, PV would require an ad hoc modified hypervisor to act as L1 (not satisfying NDP1).
+\noindent Today, NV presents several limitations related to hardware support {\sffamily and software state of implementation}.
+
+First, to support NV, the hypervisor has to provide an exact copy of hardware virtualization primitives.
+Full virtualization with hardware assistance (HAV) is the most popular technique, but requires quite a complex dedicated support.
+HAV primitives assist instruction emulation, memory address translation, and management of devices.
+Both Xen and KVM propose {\sffamily nested VMX/SVN features} that allow to efficiently handle nested hardware virtualization on single-layer hardware primitives.
+Full virtualization is currently the only virtualization technique supported for L1 virtualization due to its ability to run unmodified guest OSes -- and thus hypervisors.
+Paravirtualization (PV) would on the contrary require an ad hoc modified hypervisor to act as L1 (thus violating the NDP1 principle).
 %To address this performance problems, several hardware producers proposed new facilities, like Intel VMCS shadowing, that reduces the amount of nested VMEXIT when handling a nested guest VMEXIT.
-Although nested virtualization performance are still not good and the feature is not yet stable on modern hypervisor \cite{xen:test}, the feature is under heavy development and it will be hopefully in a usable state soon.
+Despite still poor NV performance and not being yet stable on modern hypervisors~\cite{xen:test}, such feature is under heavy development and will hopefully be in a usable state soon.
 
-Secondly, due to the monolithic architecture, each feature introduced in GPH enlarges the TCB.
-Therefore, leveraging GPH, a huge amount of code is running in the most privileged mode and has to be trust.
-Several works discuss and analyze the overall TCB of modern GPH \cite{nova, xmhf}, considering GPH TCB always bigger than 100 KLoC.
-This approach directly collides with one design principles introduced in the introduction paragraph (DP3),
+Second, due to current monolithic GPH architectures, each feature directly introduced within the hypervisor such as security enhancements enlarges the TCB~\cite{cloudvisor:zhang}.
+Therefore, a huge amount of code runs in highly privileged mode and must be trusted.
+Several works discuss and analyze the overall TCB of modern GPHs, always bigger than 100 KLoC~\cite{nova, xmhf}.
+This approach directly violates the DP3 design principle.
 %that identified a small TCB as a major requirement for a solid architecture.
@@
 %More importanly, this is particularly true looking at the monolithic approach of commercial hyper-visors, where any addition to the hypervisor body enlarges the TCB.
+Moreover, the previous architecture does not improve overall stability of the infrastructure.
+The bottom-most hypervisor L0 faces the same issues as single-layer GPH architectures: a large TCB implies also that a failure experienced in one of its critical components may trigger a generalized failure of the whole platform (violating the DP2 principle).
+Besides, in standard GPH architectures, the administrator has the entire control of the platform and could potentially snoop and inspect user VM status and data (violating the DP1 principle).
+Such designs notably do not take advantage of the possibility {\sffamily to use a platform compatibility layer}, and more modular architectures with only a tiny security-critical kernel.
 
-Finally, the above architecture does not improve the overall stability of the infrastructure.
-The underlying core, L0, has the same problem of single-layer GPH architecture: the large TCB implies also that a failure experienced in one of the critical components will simply produce a generalized failure of the whole platform, not satisfying DP2.
-Moreover, in GPH standard architecture, the administrator has the entire control of the platform and could potentially snoop and inspect user VM status and data.
-
-This kind of architecture, in particular, does not leverage completely the possibility to hide behind a compatible layer, a really modular architecture with only a tiny critical core.
-To sum up, current nested virtualization limits are represented by a certain lack of dedicated feature in modern hypervisor and the impossibility to rely on a tiny TCB (DP3, DP4).
-The first one is being addressed by hypervisor developers. The second one could simply not be achieved with modern hypervisors.
+Overall, current NV architectures lack some of the dedicated features found in modern hypervisors and the impossibility to rely on a tiny TCB (violating principles DP3 and DP4).
+If the first one is currently being addressed by hypervisor developers, the second simply has not be achieved with modern hypervisors so far.
 

### 2014-03-24 paper/draft.tex
Tentative refonte Introduction
- Quelques corrections
- Deplacement plan papier

@@
 
+%% Motivation
+%Cloud computing platform are a solid reality nowadays. 
 \noindent Distributed cloud computing is now becoming a reality.
@@
 
+% Interop
+DCC does not handle heterogeneous infrastructures which secludes providers in non-evolving hardware.
+% Securite
+Also, security remains the last hurdle toward wide adoption of DCC.
+The user data are no more self contained but distributed among all parties without any proof of confidentiality, integrity or availability.
+
+%% Specific problem considered
+We classified problems related to DCC horizontally and vertically.
+The horizontal view gathers features related to interactions between IaaS stacks, such as live migrations.
+The vertical view addresses features close to the isolation of execution environment on the same physical machine.
+This classification provide a global coverage of problems to solve to deliver a secure, stable and robust architecture.
+
+%% "In this paper we show that"
+As a first element of answer, the nested virtualization~\cite{turtle:ibm} performed a breakthrough with enhanced isolation capabilities on the same physical machine.
+The hypervisor is now virtualized by a thin layer~\cite{cloudvisor:zhang} to counter failures.
+Furthermore, the horizontal challenges were also included in this tiny layer to deliver an homogeneous interface for commodity hypervisors~\cite{art:blan, xclo:blank}.
+However these classic approaches suffers from several shortcomings detailed in Section~\ref{ref:challenges}.
+Thus we propose a novel architectural approach to undertake remaining problems.
+
+%% Differences with others
+The previous works created both horizontal and vertical basic blocks from scratch.
+While being a necessary step, these approaches does not inherits the advances performed in terms of operating system security.
+The software architecture are mainly monolithic and yet-another-layer fends off security problems without
+figuring them out.
+Our approach learns from security issues fixed by OS architecure~\cite{} and apply these patterns to leverage nested virtualization.
+In this paper we show that component-based architecures are a promising design for the next generation of clouds.
+
 This paper is organized as follows.
+Section~\ref{ref:challenges} details the constraints of nested virtualization for the two dimension of our solution.
 In Section~\ref{sec:rw} we analyze the nested virtualization ecosystem and position existing solutions regarding our design principles.
@@
 \section{Challenges and Design Principles}
+\label{ref:challenges}
 

### 2014-03-24 paper/draft.tex
updated abstract

@@
 %In this paper, we will analyze the state of nested virtualization in modern hypervisor and evaulating the impact of utilization of microhypervisor to enforce a stronger security level.
-The problem of homogeneity over multiple IaaS providers is receiving considerable attention with the evolution toward nested virtualization.
+%The problem of homogeneity over multiple IaaS providers is receiving considerable attention with the evolution toward nested virtualization.
 These new architectures unveil user-centric opportunities such as efficient resources management and VM migration.
@@
 Thus, we argue that the promising architectural view of micro hypervisors approach is the key to the separation of hypervisor duties and reducing the attack surface.
-
-Furthermore, our architecture mushrooms commodity and micro hypervisors to leverage nested virtualization in terms of user-data privacy, modularity, TCB size and compatibility.
+%
+%Furthermore, our architecture mushrooms commodity and micro hypervisors to leverage nested virtualization in terms of user-data privacy, modularity, TCB size and compatibility.
 These design principles are evaluated for each architecture configuration and compared to establish a reference security architecture.
@@
 A reference implementation proven the architecture to be viable and will be extended to meet broader hypervisor scopes, enabling migration over heterogeneous cloud servers.
+Distributed cloud computing is now moving user-centric.
+The pro-mise of this evolution is: (1) unified resource management across IaaS providers, e.g., live migration independent from hypervisor technology; (2) increased customizability to choose the hypervisor services needed to realize fully à la carte, self-service clouds.
+However, a reality check shows that major interoperability and security barriers must be lifted, and complex trade-offs be found between functionalities provided and security assurance.
+Nested virtualization (NV) brings both a homogeneous interface for distributed hypervisor features and enhanced IaaS security, protecting VMs even in case of hypervisor compromise.
+However, practical NV adoption is still hindered by code additions to an already very large hypervisor TCB and by limited hardware support.
+New distributed IaaS architectures are therefore required.\\
+
+In this paper, we argue that micro-hypervisor (MH) architectures are the key to overcome such NV limitations. 
+From the previous challenges, we derive a set of design principles that a distributed IaaS architecture should satisfy. 
+We compare possible architectural designs combining NV with micro-hypervisor features to find the best trade-off between user data privacy, modularity, TCB size, and compatibility with multiple platforms.
+We sketch how the resulting architecture may serve as a basis for 2D security management of a cloud of clouds: vertically, across layers for hardened IaaS security, and horizontally spanning cloud domains for end-to-end security.
 }
+
+%The first challenge means that threats will be both vertical, spanning layers, and horizontal, spanning domains  the SUPERCLOUD being a multi-layer, multi-domain distributed cloud infrastructure.
+Vertically, attacks currently target several infrastructure layers (e.g., VMs, hypervisor).
+Unfortunately, most existing defenses are generally for single layers only.
+This makes it difficult to grasp the overall extent of an attack.
+Horizontally, threats may propagate through the network from cloud to cloud.
+Unfortunately, most existing defenses are for single clouds only.
+Lack of interoperability in security policies and mechanisms is thus a major barrier to  unified hybrid cloud security.
+The second challenge means that although many defense mechanisms may be leveraged in this 2D landscape, they remain difficult to configure and to reconcile.
+The current “by hand” approach becomes a nightmare for security administrators.
+Despite some automated solutions for managing security of single clouds, there is currently no solution truly applicable to federations of clouds in practice.
+
+%The resulting combination is built over common hypervisors empowered with micro hypervisors features. 
+%
+%rchitecture mushrooms commodity and micro hypervisors to leverage nested virtualization
+%
+%
+%
+%Furthermore, our architecture mushrooms commodity and micro hypervisors to leverage nested virtualization in terms of user-data privacy, modularity, TCB size and compatibility.
+These design principles are evaluated for each architecture configuration and compared to establish a reference security architecture.
+%
+%enhance NV architecture next generation of distributed clouds security and interoperabiltiyi issues.
+%
+%
+%failureverall, current NV architectures lack some of the dedicated features found in modern hypervisors and the impossibility to rely on a tiny TCB (violating principles DP3 and DP4).
+%If the first one is currently being addressed by hypervisor developers, the second simply has not be achieved with modern hypervisors so far.
+%
+%ypervisors have to find a complex trade-off between functionalities provided and security assurance. Those requirements are contrasting by design:
+%%adding new features will generally increases the LoC and consequently the attack surface. On the contrary, those code additions are required to properly fulfill market requests. 
+%
+%the inherent drawbacks hinder its global adoption with a larger TCB and the complexity to monitor another virtualization layer.
+%
+%First, independence from the provider resulting in lower infrastructure operation overheads, more rapid deployment of services, but also increased homogeneity the customer being free a particular hypervisor technology.
+Second, increased customizability, as the customer can choose which virtualized services (e.g., for security) may be deployed resulting in fully a la carte clouds.
+Third, new business opportunities, as SCs and their (virtual) apps might become new markets in the manner of ecosystems from the mobile computing world (e.g., iPhone, Android).
+%
+%User-centric: de DCC
+%vers plus d'homogeneity
+%problemes 
+%verticaux
+%horizontaux.
+%
+%NV: solutions pour résoudre certains de ces problèmes.
+%
+%VErticaux: ...
+%Horizontaux: ...
+%
+%Mais limites ... + ...
+%New architectural solutions must be found.
+%
+%In this paper explore potential of micro-hypervisor architectures to enhance NV architecture next generation of distributed clouds security and interoperabiltiyi issues.
+%
+%We explore design principles properties.
+%
+%We compare architecutral solutions combining NV+MH to find the best trade off between ..., ???; ... meilleure approche.
+%
+%We sketch finally: how this type of architecture could enable to perform 2D supervision xl alyer x domain.
 
@@
 
-\section{Micro-hypervisors}
+\section{Micro-Hypervisors}
 \label{sec:mh}
@@
 % Xen and KVM TCB
-Consequentially, several works focused this problem leveraging a micro-hypervisor architecture, with a traditional microkernel-like design that tries to consistently reduce the amount of code running with the highest privilege level.
-A micro-kernel traditionally consists in a tiny core running in Kernel Mode, offering only few fundamental services to other modules running (scheduling, IPC, memory management) in a less privileged domain.
-All other components, in particular device drivers, are moved out in user-space.
-Microkernels had been questioned because of the wasteful performance of first generations, due to IPC intrinsic cost.
+
+\subsection{Principle}
+
+\noindent To solve the TCB size issue, several works have introduced the idea of \textit{micro-hypervisor (MH)} architecture~\cite{nova, NoHype, xmhf}.
+To reduce the hypervisor size, such new hypervisor architectures draw inspiration from evolution in OS architecture (e.g., extensible, micro-, exo-, component-based kernels).
+Within the hypervisor are distinguished security-critical modules from non-sensitive code.
+The aim is to expel as much code as possible from the TCB and target ultra-thin hypervisors.
+For instance, some components of the hypervisor (e.g., device drivers, VM address space management) may be virtualized outside the micro-hypervisor, isolated, and restarted in case of compromise~\cite{nova}.
+The extreme case is to remove the virtualization layer altogether~\cite{NoHype}.
+
+with a traditional microkernel-like design that tries to consistently reduce the amount of code running with the highest privilege level.
+
+A micro-kernel traditionally consists in a minimal kernel running in privileged mode, offering only a few fundamental services to other modules (e.g, scheduling, IPC, memory management) running in a less privileged domain.
+All other components, such as device drivers, are moved out in user-space.
+While the benefits of such a design has been questioned due to the cost of IPCs in first generations of micro-kernels, such arguments have since then been completely dispelled~\cite{liedtke1993ImpIPCbyKerDes,liedtke1995micConnova}.
+%Microkernels had been questioned because of the wasteful performance of first generations, due to IPC intrinsic cost.
 This argument had been completed confuted by Liedtke in 1993: \cite{liedtke1993ImpIPCbyKerDes,liedtke1995micCon}.
 
-Microkernels have the capacity of easily redesign the hierarchy of privileges.
-A user application has normally to trust the tiny micro-kernel and only the set of user-space software it uses.
-However, this set of trusted set of drivers and tools have not to be trusted by every application executed in the system.
-Normally, this allows microkernels to implement different ``personalities''.
-User applications shares the system but could have to deal with completely different system services in a transparent way.
-However, ``personalities'' interest had been hugely reduced by the fact that system services, device drivers and even application had to be completely re-conceived to fit this new design pattern.
+%Architectural solutions were also created to make the hypervisor more minimal and flexible, targetting its two most vulnerable elements: (1) the core hypervisor itself; and (2) the management VM (e.g., Xen Dom0).
+%
+%A first class of solutions aims \textit{to reduce the hypervisor size}.
+The new hypervisor architecture should be more minimal, drawing inspiration from evolution in OS architecture (e.g., extensible, micro-, exo-, component-based kernels).
+Within the hypervisor are distinghuished security-critical modules from non-sensitive code.
+The aim is to expel as much code as possible from the TCB and target ultra-thin hypervisors, also known as \textit{micro-hypervisors}.
+For instance, some components of the hypervisor (e.g., device drivers, VM address space management) may be virtualized outside the micro-hypervisor, isolated, and restarted in case of compromise~\cite{NOVA}.
+The extreme case is to remove the virtualization layer altogether~\cite{NoHype}.
+%Unfortunately, although promising, such architectures usually require extensive code rewriting.
+They also provide a very limited number of resource management services unlike an operational IaaS infrastructure.
+They are thus hard to apply to mainstream hypervisors today.
+One of the benefits of micro-kernels is to enable to easily redesign a hierarchy of privileges.
+A user application normally trusts the micro-kernel and only the set of user-space services it requires.
+However, such drivers and libraries may not be trusted by every application run on the system.
+User applications share the system but may need to support completely different set of system services.
+This allows micro-kernels to implement different ``personalities''.
+However, the interest of such ``personalities'' has been largely reduced by the fact that system services, device drivers and even applications have to be completely re-designed to fit this new architectural pattern.
 Therefore, with the introduction of VMs, microkernels lost the battle in favor of hypervisors, in terms of popularity.
@@
 As we analyzed before, they present normally a monolithic architecture, like the great part of modern OS.
-This feature kept alive the interest in microkernels, in particular trying to couple microkernel benefits with virtualization.
+This feature kept alive the interest in microkernels, in particular trying to couple microkernel benefits with virtualization\cite{Heiser}.
 
@@
 
-\subsection {MH General Issues}
+\subsection {Limitations}
 

### 2014-03-26 paper/draft.tex
updated paper.

--- 2014-03-24 paper/draft.tex
@@
 
-The adoption of an extra layer of virtualization is capable to introduce a remarkable set of new features but it does not reduce the dimension of Trusted Code Base (TCB).
+\noindent The adoption of an extra layer of virtualization is capable to introduce a remarkable set of new features but it does not reduce the dimension of Trusted Code Base (TCB).
 MH could represent a new approach to improve the potential benefits provided by the adoption of nested virtualization.

### 2014-03-26 paper/draft.tex
Updated comparison section.

--- 2014-03-24 paper/draft.tex
@@
 
-\subsection {Limitations}
+\subsection{Limitations}
 
-Despite the positives, MH approaches had demonstrated some limitations.
+\noindent Despite the positives, MH approaches had demonstrated some limitations.
 Firstly, the straightforward one is the initial lack of hardware support, lack of device drivers are still a relevant Achilles' heel for this solutions, inherited from its ancestors microkernels.
@@
 
-The adoption of an extra layer of virtualization is capable to introduce a remarkable set of new features but it does not reduce the dimension of Trusted Code Base (TCB).
-MH could represent a new approach to improve the potential benefits provided by the adoption of nested virtualization.
-In the following subsections, it will presented the four possible architectures achievable ``nesting'' of GPH and MH, underlying potentialities, drawbacks and potential cost of implementation.
+\noindent Adding an extra layer of virtualization enables to introduce a wide range of new features in the IaaS infrastructure, but does not reduce the TCB size.
+As already discussed, introducing MH design within a nested IaaS architecture could represent a new approach to improve NV benefits by overcoming such limitations.
+In what follows, we analyze the four possible architectures to achieve nesting of GPH and MH, highlighting their potential, and assessing their benefits and weaknesses w.r.t.
+to the design principles of Section~\ref{ref:challenges}, as well as their cost of implementation.
 
@@
 
-The first analyzed architecture concerns only GPH.
+\noindent The first analyzed architecture concerns only GPH.
 At first, it is important to recall the most interesting achievable by such architectures.
@@
 
-MH on MH in reported only to see the orthogonality between this approach and GPH over GPH.
+\noindent MH on MH in reported only to see the orthogonality between this approach and GPH over GPH.
 A theoretical implementation of MH over MH will provide the vertical requirements that are missing in the previous approach.
@@
 
-The third architecture, proposed in order to reduce security concerns, is built combining an GPH at L0 and a MH at L1. 
+\noindent The third architecture, proposed in order to reduce security concerns, is built combining an GPH at L0 and a MH at L1. 
 %Several MH are proposed in literature focusing on different .

### 2014-03-28 paper/draft.tex
updated section on architecture comparison.

@@
 
-\noindent The first analyzed architecture concerns only GPH.
-At first, it is important to recall the most interesting achievable by such architectures.
-In particular, several remarkable features are live migration over different platform, transparent migration of virtualized clusters between different data-centers (horizontal requirements DP1).
-Supposing the completeness and the correctness of nested virtualization hardware facilities in widespread hypervisor, GPH over GPH satisfies all horizontal design principles, but it lacks in vertical DPs as the size of TCB is the same of currently deployed architectures.
+\noindent The first architecture only features GPHs in both layers to serve as point of comparison.
+This type of architecture enables several remarkable features related to VM management flexibility such as live migration over different platforms, and transparent migration of virtualized clusters across data centers, fulfilling principle DP4.
+More generally, assuming complete and correct support of virtualization primitives for nested hardware in mainstream hypervisors, this architecture satisfies all horizontal design principles.
 
-To sum up, it could be possible to achieve live-migration between different platform and other features related to VM management flexibility but it could not still be possible to reduce the TCB.
-Analyzing MH approaches, the aim is try at most to retain to features, trying to introduce isolation and privacy improvements due to the lower TCB promised.
+However, it fails to safisfy vertical design principles as TCB size is that of currently deployed NV architectures.
+The interest of other MH-based architectures will thus be to introduce isolation and privacy improvements by reducing the TCB, while still keeping horizontal features.
 
@@
 
-\noindent MH on MH in reported only to see the orthogonality between this approach and GPH over GPH.
-A theoretical implementation of MH over MH will provide the vertical requirements that are missing in the previous approach.
-The tiny core could be able to overcome monolithic problems due to privilege level of administration tasks (DP1), the possibility to improve the resilience of the platform to a large set of software failures (DP2) and finally reduce the possibility of the vulnerabilities of the most privilege part of the software (DP3).
-However, the interest on such architecture is limited because it is not able to overcome any cited issues of single-layer MH.
-Moreover, several horizontal properties are theoretically achievable but they will require a huge effort in developing functions of interest.
-Moreover, the lack of nested VMX functionalities prevents to concretely test this kind of architecture, because MH normally rely on VMX facilities in order to create a nested HVM guests.
-To conclude, nested virtualization is exploited in order to fill eco-systems gaps with GPH but MH over MH does not provide any improvement in this sense.
+\noindent This architecture is considered only to highlight improvements brought by a MH design compared to a GPH over GPH architecture.
+An MH over MH implementation would meet the vertical requirements missing in the previous approach.
+As in single-layer MH architectures, hypervisor disaggregation and minimization overcomes issues posed by monolithic system architectures and over-privileged administrative VMs (DP1).
+The smaller resulting TCB also improves resilience of the platform against a large set of software failures (DP2) and reduces the surface of attack of the most privileged part of the IaaS infrastructure (DP3).
+
+However, the interest this class of architecture remains limited as it cannot overcome any of the generally cited issues of single-layer MHs such as compatibility with legacy hypervisor architectures (DP5).
+While several horizontal properties are theoretically achievable, they would require a large development effort to implement the relevant features such as interoperable advanced resource management.
+Moreover, the lack of nested VMX support prevents testing in practice this type of architecture, as MH usually rely on VMX primitives to create nested HVM guests.
+
+Overall, with this architecture, if vertical challenges are well addressed as in single-layer MHs, it does not provide any improvement for horizontal challenges, despite NV being used to bridge gaps between IaaS platform eco-systems.
 
@@
 
-\noindent The third architecture, proposed in order to reduce security concerns, is built combining an GPH at L0 and a MH at L1. 
+\noindent The third architecture aims to alleviate security concerns. It combines a GPH as L0 and a MH as L1. 
 %Several MH are proposed in literature focusing on different .
@@
 The attacker would have to face with an essential virtualization layer and it would not be able to attack the back-end virtualization layer.
-An interesting point of this approach is the fact of being easily be deployed, without requiring any extra code addition to the GPH, L0.
-Existing GPH-based platforms could potentially exports nested VMX features to those guests.
-Even if this is not so usual in modern cloud platform, this lack is considered temporary and related only to the experimental status of those features.
+This approach is interesting as it may be easily deployed without requiring any extra code addition to the GPH.
+Hardware support is not a real barrier as existing GPH-based platforms could potentially export nested VMX features to guests.
+Even if this is not yet the case in modern cloud platforms, this lack may be considered temporary due to the experimental status of such features.
 
-However, the amount of code running in L0 kernel space is still as big as before and, therefore, no TCB reduction. 
-Finally, adopting a MH as L1 prevents to use inter-platform enhanced features.
-As analyzed when dealing with MH design, the price of having a tiny core is normally paid with a lack of certain functionalities.
-For example, XMHF only supports one single guest.
-It could be easily noticed that this approach is not architecturally able to provide any new "horizontal" features, leveraging on nested virtualization.
-Therefore, the architecture is only able to provide the same features that are provided by today deployed ones, with an enhanced but not sufficient level of isolation.
+However, there is no TCB reduction as the code size running in L0 kernel space is as big as before. 
+Besides, adopting a MH as L1 prevents using inter-platform enhanced features.
+As already seen for MH design, the price of having a thin hypervisor is normally paid with a lack of certain features.
+For example, XMHF only supports a single guest~\cite{xmhf}.
+Thus, this architectural approach cannot provide any new horizontal features leveraging NV.
+Therefore, this approach can only provide the same features as those deployed today, with an increased but not insufficient level of isolation.
 
@@
 
-In this configuration, the TCB is smaller than other architectures, due to the modular structure of micro-kernels., executed at L0.
+\noindent In this configuration, the TCB is smaller than for other architectures, due to the modular structure of micro-kernels., executed at L0.
 Leveraging as L1 a GPH, like Xen or KVM, provides to the user an environment apparently  identical to the one used in production today, satisfying DP5.
@@
 
-\section{Vertical Properties}
-
-\section{Implementation}
 In the previous sections we proposed a functional analysis of cloud architecture in order to respect design principles that we identified in the introduction.

### 2014-03-29 paper/draft.tex
Upd arch. comparison.

@@
 Thus, this architectural approach cannot provide any new horizontal features leveraging NV.
-Therefore, this approach can only provide the same features as those deployed today, with an increased but not insufficient level of isolation.
+Therefore, this approach can only provide the same features as those deployed today, with an increased but not sufficient level of isolation.
 
@@
 \label{par:gom}
+
+\noindent This approach achieves a smaller TCB than other architectures, due increased modularity brought by the micro-kernel-like design of the L0 layer -- satisfying DP3.
+Leveraging a GPH like Xen or KVM for the L1 layer provides the user with an environment looking identical to those used in production today -- satisfying DP5.
+The L1 GPH is normally integrated with modern cloud toolkits like OpenStack and will thus support third-party applications -- satisfying DP4.
+Therefore, this architecture achieves a stronger level of isolation due to vertical use of NV to introduce an additional hardware-protected security monitor.
+Thanks to the GPH, the MH layer remains transparent to the user and should not impact the functional properties of the IaaS platform.
+
+The first class of possible drawbacks of this architecture is represented by the necessity to expose to the L1 nested hypervisor the nested VMX extension to let him ``nest'' another HVM guest.
+This is experimentally done by GPH like Xen and KVM, but it is generally not included in the essential features provided by a minimal micro-hypervisor.
+As reported in XMHF paper \cite{xmhf}, the amount of lines of code for nested VMX is around 1KLOC.
+More generally, from an architectural point of view, the general MH minimalist attitude could represent a limitation to general preservation of existing platform functions.
+The constraint of privileged code size forces normally MH developers to drop some features, considered not essential.
+Therefore, several limitations could arise not only for nested virtualization proper features, but also for other class of features.
+For instance, XHMF is able only to virtualize one L1 instance at time.
+This is acceptable if the platform objective is limited to be provided of a secure core, but dramatically limits the possibility of having a platform model able to give the user the possibility to run its proper hypervisor as L1.
+
+Another important class is the necessity to explicit hardware devices support.
+Device drivers represents a general Achilles' heel of MH, due the general cost of development.
+However, this class of problem could be addressed leveraging device assignment.
+L1 GPH hypervisor will handle the real device, using the proper driver now demoted to run in a lower level of privilege.
+It has to be underlined that this solution could represent a feasible approach only if it possible to overcome the exclusive allocation trouble (SR-IOV).
 
@@
 \end{table*}
-
-\noindent In this configuration, the TCB is smaller than for other architectures, due to the modular structure of micro-kernels., executed at L0.
-Leveraging as L1 a GPH, like Xen or KVM, provides to the user an environment apparently  identical to the one used in production today, satisfying DP5.
-The L1 GPH is normally integrated with modern cloud toolkits like OpenStack and will support 3rd part applications.
-Therefore, the architecture presents a stronger isolation leveraging on nested virtualization, providing in particular a strong tiny core(DP3).
-The MH layer will be transparent from the user and it will not not any change in the functioning of the platform.
-
-The first class of possible drawbacks of this architecture is represented by the necessity to expose to the L1 nested hypervisor the nested VMX extension to let him ``nest'' another HVM guest.
-This is experimentally done by GPH like Xen and KVM, but it is generally not included in the essential features provided by a minimal micro-hypervisor.
-As reported in XMHF paper \cite{xmhf}, the amount of lines of code for nested VMX is around 1KLOC.
-More generally, from an architectural point of view, the general MH minimalist attitude could represent a limitation to general preservation of existing platform functions.
-The constraint of privileged code size forces normally MH developers to drop some features, considered not essential.
-Therefore, several limitations could arise not only for nested virtualization proper features, but also for other class of features.
-For instance, XHMF is able only to virtualize one L1 instance at time.
-This is acceptable if the platform objective is limited to be provided of a secure core, but dramatically limits the possibility of having a platform model able to give the user the possibility to run its proper hypervisor as L1.
-
-Another important class is the necessity to explicit hardware devices support.
-Device drivers represents a general Achilles' heel of MH, due the general cost of development.
-However, this class of problem could be addressed leveraging device assignment.
-L1 GPH hypervisor will handle the real device, using the proper driver now demoted to run in a lower level of privilege.
-It has to be underlined that this solution could represent a feasible approach only if it possible to overcome the exclusive allocation trouble (SR-IOV).
 

### 2014-03-29 paper/draft.tex
Upd. arch. comparison.

@@
 \end{center}
-\caption{}
+\caption{Different Types of Nested Architectures.}
 \label{fig:io}
@@
 As already discussed, introducing MH design within a nested IaaS architecture could represent a new approach to improve NV benefits by overcoming such limitations.
-In what follows, we analyze the four possible architectures to achieve nesting of GPH and MH, highlighting their potential, and assessing their benefits and weaknesses w.r.t.
+In what follows, we analyze the four possible architectures shown in Figure~\ref{fig:io} to achieve nesting of GPH and MH, highlighting their potential, and assessing their benefits and weaknesses w.r.t.
 to the design principles of Section~\ref{ref:challenges}, as well as their cost of implementation.
 
-\subsection{GPH Over GPH}
+\subsection{GPH/GPH}
 
@@
 
-\subsection{MH Over MH}
+\subsection{MH/MH}
 
-\noindent This architecture is considered only to highlight improvements brought by a MH design compared to a GPH over GPH architecture.
-An MH over MH implementation would meet the vertical requirements missing in the previous approach.
+\noindent This architecture is considered only to highlight improvements brought by a MH design compared to a GPH/GPH architecture.
+An MH/MH implementation would meet the vertical requirements missing in the previous approach.
 As in single-layer MH architectures, hypervisor disaggregation and minimization overcomes issues posed by monolithic system architectures and over-privileged administrative VMs (DP1).
@@
 
-\subsection{MH Over GPH}
+\subsection{MH/GPH}
 \label{par:mog}
@@
 
-\subsection{GPH Over MH}
+\subsection{GPH/MH}
 \label{par:gom}
 
-\noindent This approach achieves a smaller TCB than other architectures, due increased modularity brought by the micro-kernel-like design of the L0 layer -- satisfying DP3.
+\noindent This last approach achieves a smaller TCB than other architectures, due increased modularity brought by the micro-kernel-like design of the L0 layer -- satisfying DP3.
 Leveraging a GPH like Xen or KVM for the L1 layer provides the user with an environment looking identical to those used in production today -- satisfying DP5.
-The L1 GPH is normally integrated with modern cloud toolkits like OpenStack and will thus support third-party applications -- satisfying DP4.
+The L1 GPH is normally integrated with modern cloud toolkits like OpenStack, and will thus support third-party applications -- satisfying DP4.
 Therefore, this architecture achieves a stronger level of isolation due to vertical use of NV to introduce an additional hardware-protected security monitor.
@@
 
-The first class of possible drawbacks of this architecture is represented by the necessity to expose to the L1 nested hypervisor the nested VMX extension to let him ``nest'' another HVM guest.
-This is experimentally done by GPH like Xen and KVM, but it is generally not included in the essential features provided by a minimal micro-hypervisor.
-As reported in XMHF paper \cite{xmhf}, the amount of lines of code for nested VMX is around 1KLOC.
-More generally, from an architectural point of view, the general MH minimalist attitude could represent a limitation to general preservation of existing platform functions.
-The constraint of privileged code size forces normally MH developers to drop some features, considered not essential.
-Therefore, several limitations could arise not only for nested virtualization proper features, but also for other class of features.
-For instance, XHMF is able only to virtualize one L1 instance at time.
-This is acceptable if the platform objective is limited to be provided of a secure core, but dramatically limits the possibility of having a platform model able to give the user the possibility to run its proper hypervisor as L1.
+This class of architecture may suffer from two types of limitations.
+First, nested VMX extensions must be exposed to the L1 hypervisor to let him ``nest'' another HVM guest.
+This requirement is implemented by GPHs like Xen and KVM, but usually not included in the key features to be provided by a MH: implementing nested VMX may be estimated to around 1KLoC, to be compared with the size of a typical MH (2-20 KLoC)~\cite{xmhf}.
+More generally, the MH quest towards ever increasing minimalism could represent a serious limitation to preserving functions of existing platforms: the constraint of privileged code size may force MH developpers to drop some features considered as non-essential.
+This may apply to NV-related primitives, but also to other types of features.
+For instance, XMHF may only virtualize a single L1 instance at a time.
+While this may be acceptable if the platform objective to enforce strict resource access control in a single-provider private cloud, it hinders having a platform model enabling the user to run its own hypervisor as L1 to set up a self-service cloud independently from the underlying provider -- as in that case, multiple L1 hypervisors would need to run concurrently on the L0.
 
-Another important class is the necessity to explicit hardware devices support.
-Device drivers represents a general Achilles' heel of MH, due the general cost of development.
-However, this class of problem could be addressed leveraging device assignment.
-L1 GPH hypervisor will handle the real device, using the proper driver now demoted to run in a lower level of privilege.
-It has to be underlined that this solution could represent a feasible approach only if it possible to overcome the exclusive allocation trouble (SR-IOV).
+Second, device drivers should be correctly managed.
+Drivers are a major MH issue, due to costs of development or of adaptation to a new architecture.
+Such challenges may be addressed using device assignment: the L1 GPH hypervisor manage the physical device with the proper driver now running in a lower level of privilege.
+To be workable, this approach would require exclusive allocation to be simply implementable, e.g., using SR-IOV~\cite{sriov}.
 
 \begin{table*}
-\label{fin:conf}
+\caption{Architecture Comparison.\label{fin:conf}}
 \centering
@@
 \toprule
-Id & Functional design Principle & Class & GPH over GPH & MH over MH & MH over GPH & GPH over MH\\
+Id & Functional Design Principle & Class of Requirement & GPH/GPH & MH/MH & MH/GPH & GPH/MH\\
 \midrule
-DP1 & User-Data privacy & Vertical & $\times$  & $\checkmark$ & $\times $ & $\checkmark$ \\
+DP1  & User-Data Privacy               & Vertical   & $\times$      & $\checkmark$  & $\times$ & $\checkmark$ \\
 DP2  & Fail-Safe modular architecture  & Vertical   & $\times$      & $\checkmark$  & $\times$ & $\checkmark$ \\
-DP3 & Small TCB & Vertical &  $\times$& $\checkmark$ & $\times$ & $\checkmark$ \\
-DP4 & Inter-platform support & Horizontal & $\checkmark$  & $\times$ & $\times$ & $\checkmark$ \\
-DP5 & 3rd part compatibility & Horizontal & $\checkmark$  & $\times$ & $\times$ & $\checkmark$ \\
-NDP1 & Minimize code addition & - & $\checkmark$  & $\times$ & $\times$ & $\checkmark$ \\
+DP3  & Small TCB                       & Vertical   &  $\times$     & $\checkmark$  & $\times$ & $\checkmark$ \\
+DP4  & Inter-Platform support          & Horizontal & $\checkmark$  & $\times$      & $\times$ & $\checkmark$ \\
+DP5  & Compatibility with Legacy       & Horizontal & $\checkmark$  & $\times$      & $\times$ & $\checkmark$ \\
+NDP1 & Minimal Additional Code         & --         & $\checkmark$  & $\times$      & $\times$ & $\checkmark$ \\
 \bottomrule
@@
 
-\subsection{Lessons Learned}
+\subsection{Discussion}
 
-To conclude, this section provides a short comparison between different platforms, contrasting architecture features above the two main different types of benefits given by nested virtualization, as reported widely in literature \cite{art:blan, nested:berghmans,  cloudvisor:zhang, turtle:ibm}.
-The first could be identified in the augmented degree of isolation the architecture could enforce between host OS and the "customer" point of access, allowing to redesign the security architecture of the whole platform.
-The second one is represented by the possibility to leverage multiple virtualization layers to easily define inter-platform policy and operations between different platforms.
-In table \ref{fin:conf}, it is possible a summary of the previous analysis.
-To conclude, GPH over MH is the most interesting one in order to enlarge nested virtualization promised benefit.
-This architecture with a solid and secured MH as L0 is the only able to reduce the platform TCB and this could be a key point to obtain more protected infrastructure, .
-Firstly, adopting a MH as architecture trusted core, we will have a secure layer guaranteeing that administrator could not lose the control of the system.
-Klein et Al.
-\cite{Klein:2009} have demonstrated that a kernel with less than 10KLoC could be formally verified, under several conditions.
-This is far from be achievable today with standard GPH over GPH approach where the core is not tiny and the TCB is an order of
-magnitude bigger than MH \cite{nova}.
+\noindent Overall, how do the architectures above compare w.r.t.
+to the two main reported benefits of NV, i.e., vertical, through an increased level of isolation between hypervisor and customer VMs, and horizontal to easily achieve multiple provider platform interoperability at system level?
+Table~\ref{fin:conf} summarizes the previous analysis.
 
-However, the effort to set up a GPH over MH is strongly dependent on which technique leveraging for virtualizing L2.
-As stated in paragraph \ref{par:gom}, with L2 guest virtualized with the hardware assitance technique, the most popular one, the MH needs to export nVMX at least.
-On the contrary, adopting paravirtualization/Dynamic binary translation for L2 does not necessitate a particular hardware support and so it could not be necessary patch the MH to set up a platform. 
-However, it has not be underestimated, however, that para-virtualization supports only a subset of OS, and could not be considered a general solution and DBT has shown important overhead.
+The GPH/MH architecture appears as the most interesting to expand the NV promised benefits.
+Through a solid and secure MH as L0, it is the only one able to reduce the platform TCB.
+This would bring several benefits towards a more secure infrastructure.
+First, a guarantee that the platform administrator could not lose control of the system thanks to an extra security layer.
+Second, certification perspectives: Klein et al.~\cite{Klein:2009} showed that less than 10 KLoC kernels could be formally verified, under several conditions, which is far from be achievable today with standard GPH/GPH approach where the TCB is an order of
+magnitude bigger than MHs~\cite{nova}.
+
+However, the effort to set up a GPH/MH is strongly dependent on the technique for L2 virtualization.
+As already seen, to support L2 guests with hardware-assisted virtualization which is becoming mainstream, the MH needs to export nVMX (nested VMX) primitives.
+On the other end, adopting paravirtualization or dynamic binary translation for L2 does not requires particular hardware support, and thus would not require MH patching to set up a running platform.
+However, PV remains limited by supporting only a subset of system features which could be an issue as a general solution.
 %The last point of interest is related to device management.
@@
 
-To sum up,  the adoption of MH could  increase feature set introduced by nested virtualization adoption.
-The first impeding problem is the implementation of MH, able to propose to guest to virtualize with hardware assistance.
-In next section, we present an implementation of GPH over MH, that try to circumvent this limitation leveraging paravirtualization.
+Therefore, a MH-based architecture could increase the feature set of current NV architectures, mostly vertically.
+The main remaining barrier is to come up with an MH implementation supporting HAV nested guests.
+We are currently working on a GPH/MH implementation based on Xen over NOVA aiming to lift this limitation using PV.
 
@@
 In the previous sections we proposed a functional analysis of cloud architecture in order to respect design principles that we identified in the introduction.
-In this section, we propose a GPH over MH implementation, based on Xen over NOVA \cite{nova}.
+In this section, we propose a GPH over MH implementation, based on .
 

### 2014-03-29 paper/draft.tex
commit

@@
 \usepackage{times}
+\usepackage{url}
 %\pagestyle{fancy} % options: empty , plain , fancy
@@
 %\lfoot{}\cfoot{\thepage}\rfoot{}
+
+\hyphenation{he-terogeneous infra-structure lea-kage rea-lizing accep-table modula-rity tool-kits ho-rizontally me-chanisms}
 
@@
 
-%The first challenge means that threats will be both vertical, spanning layers, and horizontal, spanning domains  the SUPERCLOUD being a multi-layer, multi-domain distributed cloud infrastructure.
-Vertically, attacks currently target several infrastructure layers (e.g., VMs, hypervisor).
-Unfortunately, most existing defenses are generally for single layers only.
-This makes it difficult to grasp the overall extent of an attack.
-Horizontally, threats may propagate through the network from cloud to cloud.
-Unfortunately, most existing defenses are for single clouds only.
-Lack of interoperability in security policies and mechanisms is thus a major barrier to  unified hybrid cloud security.
-The second challenge means that although many defense mechanisms may be leveraged in this 2D landscape, they remain difficult to configure and to reconcile.
-The current “by hand” approach becomes a nightmare for security administrators.
-Despite some automated solutions for managing security of single clouds, there is currently no solution truly applicable to federations of clouds in practice.
-
 %The resulting combination is built over common hypervisors empowered with micro hypervisors features. 
@@
 %Cloud computing platform are a solid reality nowadays. 
-\noindent Distributed cloud computing is now becoming a reality.
-Modern cloud platforms are able to provide elastic use of resources, optimizing their allocation to meet user needs.
-However, currently deployed distributed cloud infrastructures present several limitations.
+% Modern cloud platforms are able to provide elastic use of resources, optimizing their allocation to meet user needs. 
+% la prochaine mutation majeure du cloud computing est celle du nuage hybride, qui fédère dans une même infrastructure nuages privés, publics, ou communautaires.
+L’intérêt est de s’affranchir d’un fournisseur de nuage dédié, ou de bénéficier de la sécurité d’un nuage privé, tout en profitant de la flexibilité d’un nuage public.
+% User-centric...
+% Cloud computing today is essentially still provider-centric (PC).
+The landscape is an increasing number of fiercely competing cloud service providers (CSP), operating multiple heterogeneous clouds.
+Each CSP offers its own, feature-rich offerings for customer VMs, with a great diversity of services in areas like security (e.g., system and network intrusion monitoring, firewalling, anti-malware, data security), resource management (e.g., elastic load balancing), or high availability (e.g., checkpointing/recovery).
+Such services usually come in the form of virtual appliances, often based on hypervisor support.
+% A strong push has thus recently been made towards a user-centric (UC) vision of the cloud  It might be described as a paradigm change in delivering cloud resources (e.g., compute, storage).
+To lift the interoperability barrier,  a resource distribution plane, a.k.a.
+super-clouds (SCs) is introduced, spanning CSPs to decouple resource production (CSPs) from consumption (customers).
+To tackle the control challenge, this plane is also an extensibility layer, where each new stakeholder operating the distribution layer (a.k.a.
+SC provider) may provide its own services enabling the customer to deploy self-service, customizable clouds ranging from the SaaS to the full IaaS level.
+Technically, the idea is to split the cloud infrastructure into provider- and user-controlled domains or modules, several design alternatives being available such extensible hypervisors, flexible hypervisor management interface (dom0), or nested virtualization.
+%This vision brings a number of benefits.
+First, independence from the provider resulting in lower infrastructure operation overheads, more rapid deployment of services, but also increased homogeneity the customer being free a particular hypervisor technology.
+Second, increased customizability, as the customer can choose which virtualized services (e.g., for security) may be deployed resulting in fully a la carte clouds.
+Third, new business opportunities, as SCs and their (virtual) apps might become new markets in the manner of ecosystems from the mobile computing world (e.g., iPhone, Android).
 
-% Interop
-DCC does not handle heterogeneous infrastructures which secludes providers in non-evolving hardware.
-% Securite
-Also, security remains the last hurdle toward wide adoption of DCC.
-The user data are no more self contained but distributed among all parties without any proof of confidentiality, integrity or availability.
+\noindent \textit{Distributed cloud computing (DCC) }is now turning into a reality: modern hybrid cloud platforms enable to benefit seamlessly both from private cloud security and from public cloud flexibility.
+DCC is also moving user-centric: customers demand greater empowerment to deploy self-service clouds really adapted to their needs, and to move away from a provider-dominated landscape of highly heterogeneous closed platforms and its infamous vendor lock-in~\cite{CSA:TT}.
 
-%% Specific problem considered
-We classified problems related to DCC horizontally and vertically.
-The horizontal view gathers features related to interactions between IaaS stacks, such as live migrations.
+% Mais pb:
+% We classified problems related to DCC horizontally and vertically.
+% %% Cclassification des problèmes: Specific problem considered
+%The horizontal view gathers features related to interactions between IaaS stacks, such as live migrations.
 The vertical view addresses features close to the isolation of execution environment on the same physical machine.
 This classification provide a global coverage of problems to solve to deliver a secure, stable and robust architecture.
+However, current \textit{distributed cloud infrastructures (DCI) }suffer from two sets of limitations. These may be broadly classified as \textit{horizontal} and \textit{vertical} challenges.
+% Problèmes horizontaux: Mais Interop
+%DCC does not handle heterogeneous infrastructures which secludes providers in non-evolving hardware.
+%2)	Un manque de vision globale et de flexibilité dans la gestion de la sécurité.
+En cause : l'hétérogénéité forte des politiques et mécanismes entre nuages élémentaires et leur manque d’interopérabilité.
+Les plates-formes d’interopérabilité actuelles sont de plus trop monolithiques : elles dépendent fortement de la volonté des fournisseurs de nuages de déployer certains services, et ne conviendront donc pas à tous les types de clients.
+% 
+%Unfortunately, this vision suffers from a lack of homogeneity.
+Many of those services have not been deployed due to deficiencies in: (1) unified control; and (2) interoperability.
+Lack of control is visible through vendor lock-in: such services are so tightly coupled with the CSP and on its willingness to deploy them, with little leeway for the customer.
+Control is also severely limited by monolithic infrastructures (e.g., non-flexible hypervisors hiding hardware capabilities) preventing fine-grained cloud customization by the customer.
+Lack of interoperability is mainly due to heterogeneity of services, hypervisor-specific and thus not compatible across CSPs.
+Despite much efforts, standardization might be too slow or too partial to solve such issues, as both customers and providers might want features that are not in the standard.
+%•	Lack of flexibility and control for overall security management: this refines the previous interoperability and control challenges for security.
+For interoperability, the problem comes from heterogeneity of security components and policies between CSPs and SCs.
+This has a strong security impact by introducing more vulnerabilities due to mismatching APIs.
+How also to manage interoperability and trust between SCs?
+For flexibility: how to clearly separate the infrastructure between UC and PC security management elements?
+how to make that frontier flexible to achieve a continuum between PC and UC security services?
+and how to define in that continuum security services that are flexibly composable with other types of services (e.g., replication, transactions)?
+Horizontally, DCIs lack interoperability, mainly due to heterogeneity of IaaS services, hypervisor-specific and not compatible across providers.
+Control is also severely limited by monolithic infrastructures preventing fine-grained cloud customization by users.
+% Problèmes verticaux: Securite
+%Also, security remains the last hurdle toward wide adoption of DCIs.
+The user data are no more self contained but distributed among all parties without any proof of confidentiality, integrity or availability.
+% 1)	Un manque de sécurité global.
+Dans chaque nuage élémentaire formant le nuage hybride, chaque couche logicielle (machines virtuelles, hyperviseur) s'avère extrêmement vulnérable aux attaques, rendant difficile une protection intégrée forte.
+% •	Security vulnerabilities in infrastructure layers:  For instance, the hypervisor and its over-privileged dom0 is a target of choice for attackers due to its complexity.
+This makes difficult an integrated protection.
+How to place / coordinate counter-measures in those layers that usually are not aware of one another, each with their own monitoring mechanisms, while attacks span several layers?
+How also to continue to guarantee VM protection when one or several of those layers are compromised?
+Vertically, security remains the last hurdle toward wide adoption of DCIs.
+Due to their complexity, infrastructure layers remain extremely vulnerable to attacks, making it difficult to achieve strong isolation and integrated protection.
 
-%% "In this paper we show that"
-As a first element of answer, the nested virtualization~\cite{turtle:ibm} performed a breakthrough with enhanced isolation capabilities on the same physical machine.
+\newpage
+%% La NV comme élément de réponse:
+% As a first element of answer, the nested virtualization~\cite{turtle:ibm} performed a breakthrough with enhanced isolation capabilities on the same physical machine.
 The hypervisor is now virtualized by a thin layer~\cite{cloudvisor:zhang} to counter failures.
 Furthermore, the horizontal challenges were also included in this tiny layer to deliver an homogeneous interface for commodity hypervisors~\cite{art:blan, xclo:blank}.
-However these classic approaches suffers from several shortcomings detailed in Section~\ref{ref:challenges}.
-Thus we propose a novel architectural approach to undertake remaining problems.
+% Nested virtualization (NV) brings both a homogeneous interface for distributed hypervisor features and enhanced IaaS security, protecting VMs even in case of hypervisor compromise. 
+% The hypervisor is now virtualized by a thin layer
+Advances in \textit{nested virtualization (NV)} were a major step forward to tackle such challenges~\cite{turtle:ibm}.
+By virtualizing the hypervisor through a thin software layer, NV brings enhanced isolation capabilities on the same physical machine, protecting VMs even in case of hypervisor compromise~\cite{cloudvisor:zhang}.
+NV also provides a distributed homogeneous interface for commodity hypervisors~\cite{art:blan,xclo:blank}.
+% Mais..
+% - Mais limites...
+% However these classic approaches suffers from several shortcomings detailed in Section~\ref{ref:challenges}. 
+Practical NV adoption is still hindered by code additions to an already large hypervisor TCB (Trusted Computing Base) and by limited hardware support.
+There is also currently no NV architecture solving both vertical and horizontal challenges simultaneously.
+% - New architectural challenges are required...
+New distributed IaaS architectures are therefore required.
 
-%% Differences with others
-The previous works created both horizontal and vertical basic blocks from scratch.
+%The previous works created both horizontal and vertical basic blocks from scratch.
 While being a necessary step, these approaches does not inherits the advances performed in terms of operating system security.
-The software architecture are mainly monolithic and yet-another-layer fends off security problems without
-figuring them out.
+The software architecture are mainly monolithic and yet-another-layer fends off security problems withoutfiguring them out.
 Our approach learns from security issues fixed by OS architecure~\cite{} and apply these patterns to leverage nested virtualization.
@@
 
+In this paper, we propose a new architectural approach for DCIs.
+We argue that advances in OS design bring significant improvements to traditional NV architectures: instead of relying on general-purpose hypervisors (GPHs) for both virtualization layers, new kernel designs allow to tackle more efficiently different sets of challenges separately in each layer.
+We make a position statement that \textit{micro-hypervisor (MH) architectures} combined with NV benefits are a promising design for the next generation of DCIs.
+
+% Contributions.
+%1)
+%2)
+%3)
+This paper makes the following contributions.
+From previous challenges, we derive a set of design principles that a DCI architecture should satisfy.
+We compare possible architectural designs combining NV with MH features to find the best trade-off between user data privacy, modularity, TCB size, and compatibility with multiple platforms.
+We finally sketch how the resulting architecture may serve as basis for 2D security management of DCIs: vertically, across layers for hardened IaaS security, and horizontally spanning cloud domains for end-to-end security.
+ %The first challenge means that threats will be both vertical, spanning layers, and horizontal, spanning domains  the SUPERCLOUD being a multi-layer, multi-domain distributed cloud infrastructure.
+Vertically, attacks currently target several infrastructure layers (e.g., VMs, hypervisor).
+Unfortunately, most existing defenses are generally for single layers only.
+This makes it difficult to grasp the overall extent of an attack.
+Horizontally, threats may propagate through the network from cloud to cloud.
+Unfortunately, most existing defenses are for single clouds only.
+Lack of interoperability in security policies and mechanisms is thus a major barrier to  unified hybrid cloud security.
+The second challenge means that although many defense mechanisms may be leveraged in this 2D landscape, they remain difficult to configure and to reconcile.
+The current “by hand” approach becomes a nightmare for security administrators.
+Despite some automated solutions for managing security of single clouds, there is currently no solution truly applicable to federations of clouds in practice.
+
 This paper is organized as follows.
-Section~\ref{ref:challenges} details the constraints of nested virtualization for the two dimension of our solution.
-In Section~\ref{sec:rw} we analyze the nested virtualization ecosystem and position existing solutions regarding our design principles.
-In Section~\ref{sec:classic}, we details limitations of classical nested virtualization approach and in Section~\ref{sec:mh} we present the microvisor approach.
-In Section~\ref{sec:archcomp} we compare cloud architectures propositions, combining nested virtualization and microvisors, leveraging the design principles grid introduced before.
+Section~\ref{ref:challenges} presents the DCI challenges and design principles.
+Sections~\ref{sec:rw} and~\ref{sec:classic} review the NV ecosystem and analyze the limitations of traditional NV design.
+Section~\ref{sec:mh} presents the MH approach.
+Section~\ref{sec:archcomp} compares possible DCI architectures combining NV and MH according to the design principles.
+Finally, Section~\ref{sec:arch} presents a preliminary DCI system architecture based on the best resulting design with promising properties for 2D management of DCI security.
 %Furthermore, in Section 6 we define a security architecture for multi-cloud scenario, with an analysis of necessary software components to satisfy vertical design principles.
-In Section~\ref{sec:conclu}, we conclude the paper with future works and our vision of the SuperCloud.
+%In Section~\ref{sec:conclu}, we conclude the paper with future works and our vision of the SuperCloud.
 % in section in section 6 we define a security architecture for multi-cloud scenario, with an analysis of necessary software components to satisfy vertical design principles.
@@
 
-First, deploying cloud applications able to transparently leverage resources of several heterogeneous platforms is still difficult.
-The concept of super-cloud~\cite{art:blan, xclo:blank} tries notably to overcome such issues.
-Multi-platform coordination is hard, primarily due to lock-in to different standards such as VM formats.
+\subsection{Challenges}
+
+\noindent Deploying cloud applications able to transparently use resources of several heterogeneous platforms is still difficult. 
+%The concept of super-cloud~\cite{art:blan, xclo:blank} tries notably to overcome such issues. 
+Multi-platform coordination is hard, primarily due to lock-in to different standards.
 Unfortunately, such lock-in is not only due to market strategic decisions, but also reflects heterogeneity of virtualization technologies.
-For example, each cloud provider may use different storage systems or different formats to store VMs images.
-Different hypervisors have nowadays a low degree of compatibility between them.
-This generally prevents leveraging some popular features like VM live migration between different platforms.
-Incompatibility is even greater if we consider that each platform may adopt several types of virtualization techniques.
+For example, each cloud provider may use different storage systems or formats for VM images.
+Different hypervisors have today a low degree of compatibility between them.
+This generally hinders using some popular features like VM live migration across different platforms.
+Incompatibility is even greater considering each platform may adopt several types of virtualization techniques.
 This situation induces a tight dependency between the customer and the cloud provider.
-It may prevent guaranteeing service availability by relying on alternative providers, or to set up geographical optimizations due to cost or origin of traffic.
-This has resulted in some dramatic data server failures in recent years, causing prolonged outage of service for many applications.
-In what follows, we refer to this class of multi-platform issues as \textit{horizontal} challenges.
+It may prevent guaranteeing service availability by relying on alternative providers, or setting up geographical optimizations according to cost or origin of traffic.
+This resulted in some dramatic data server failures in recent years, causing prolonged outage of service for many applications.
+We refer to this class of multi-platform issues as \textit{horizontal challenges}.
 
-Second, a number of issues have been raised about user data privacy and platform integrity.
-From the user perspective, users would like to preserve privacy of their data stored on the cloud platform.
-However, due to overprivileged hypervisors, a curious or malicious administrator could easily retrieve such data~\cite{cloudvisor:zhang}.
-In a multi-tenant cloud platform, physical resource sharing between users may lead not only to data leakage or but also attacks against the IaaS~\cite{bot:att, sec:you}.
-From the provider perspective, a key objective is to thus guarantee platform integrity against rogue VMs.
-However, commodity general-purpose hypervisors (GPHs) make it difficult to observe deeply all layers of a running IaaS platform.
-Indeed, running GPHs at the highest level of privilege hampers detection and reaction on the hypervisor itself in presence of faulty or malicious behavior -- unless to cause interruption of service for all hosted VMs.
-Thus, both user-centric and provider-centric visions of the cloud have a problem with overprivileged hypervisors.
+Several concerns were also raised about user data privacy and platform integrity.
+For users, preserving the privacy of their data stored on the cloud platform is important.
+However, due to overprivileged hypervisors, a curious or malicious administrator may easily retrieve such data~\cite{cloudvisor:zhang}.
+In a multi-tenant cloud, physical resource sharing between users may lead not only to VM data leakage but also to attacks against the IaaS~\cite{bot:att, sec:you}.
+For the provider, a primary objective is thus to guarantee platform integrity.
+Unfortunately, commodity GPHs are not well adapted to observe deeply all layers of a running IaaS platform: running a GPH at the highest privilege level hampers detection and reaction on the hypervisor itself in presence of faulty or malicious behavior.
+Thus, both users and providers have a problem with overprivileged hypervisors.
 %: the platform administrator cannot perform in-depth infrastructure monitoring nor can users concretely inspect administrator actions on their data. 
-A major challenge for new generations of cloud infrastructures is thus to find acceptable trade-offs between such conflicting requirements.
-In what follows, we refer to this class of layer-oriented issues as \textit{vertical} challenges.
+%A major challenge for new generations of DCIs will be to find acceptable trade-offs between such conflicting requirements. 
+We refer to this class of layer-oriented issues as \textit{vertical challenges}.
 
-We identified five key design principles (DP) that a IaaS architecture should implement to meet the previous challenges.
-In what follows, those principles will serve as reading template to analyze different architectures.
+\subsection{Design Principles}
+
+\noindent We identified the following key DCI design principles to meet the previous challenges: 
+%In what follows, those principles will serve as reading template to analyze different architectures.
 \begin{itemize}[]
-\item \textbf{DP1 User Data Privacy} A curious or malicious administrator (or another platform user) should not be authorized to inspect and analyze user data and VM instances without explicit user consent relying on mutually trusted services.
-A trade-off should be found to allow providers to detect and react to potential threats, but also to prevent unlimited platform and user data access in such inappropriate operator behaviors.
-Thus, a platform administrator should not manage the infrastructure directly using the highest level of privilege.
+\item \textbf{DP1 User Data Privacy} An administrator (or another platform user) should not be authorized to inspect and analyze user data and VM instances without explicit user consent.
+A trade-off should be found to allow providers to detect and react to potential threats, but also to prevent unlimited platform and user data access by curious or malicious operators.
+Thus, a platform administrator should not manage the infrastructure directly using the highest privilege level.
 Instead, his actions on user data objects should be validated by a mutually-trusted kernel run in the most privileged mode.
 This kernel should also enforce memory page encryption to prevent snooping when pages are swapped out.
-\item \textbf{DP2 Fail-Safe Modular Architecture} The architecture should be fail-safe: for instance, to enable automatically migration of instances or user-data in presence of a threat and to recover from software failures.
-To reach this goal, the amount of safety-critical components should be as low as possible to mimize the attack surface.
-Similarly, the overall design should be as modular as possible to enable platform reconfiguration.
-\item \textbf{DP3 Small TCB} The number of potential failures and bugs in the platform is directly linked with the code size executed at the highest level of privilege.
-Thus, an architecture with a tiny TCB (Trusted Computing Base) will improve safety and integrity by design.
-\item \textbf{DP4 Inter-Platform Support} The cloud architecture should allow supporting directly different platforms and service providers.
-The same user virtual instances should be allowed to be run (and potentially migrated) unmodified over different IaaS platforms.
+\item \textbf{DP2 Fail-Safe Modular Architecture} The architecture should be fail-safe, e.g., for automatically migration of VMs or user-data during an attack, or to recover from software failures.
+The number of safety-critical components should minimal to reach a low attack surface.
+The overall design should also be highly modular to enable platform reconfiguration.
+\item \textbf{DP3 Small TCB} The number of vulnerabilities and failures in the platform is directly linked with the code size run at the highest privilege level.
+A small TCB architecture will improve safety and integrity by design.
+\item \textbf{DP4 Inter-Platform Support} The cloud architecture should support different platforms and providers.
+The user VMs should run (and potentially be migrated) unmodified over different IaaS platforms.
 \item \textbf{DP5 Compatibility with Legacy} Evolution in the IaaS architecture should avoiding disrupting existing third party applications.
-Thus, platform interfaces should be kept as close as possible to existing ones, to prevent code rewriting notably for applications.
+Platform interfaces should be kept as close as possible to existing ones to prevent code rewriting notably for applications.
 \end{itemize}
-In addition to the above functional design principles that specify which features have to be achieved by the platform architecture, another set of non-design principles (NDP) should be considered regarding effective feasibility of implementing the architecture.
-We just give an example in terms of platform usability:
+In addition to the previous functional principles, another set of principles should be considered regarding effective feasibility of implementing the DCI architecture.
+For instance in terms of platform usability:
 \begin{itemize}[]
-\item \textbf{NDP1 Minimal Additional Code} A major issue could be represented by the amount of code to be added and maintained to an existing architecture to meet functional requirements.
-Despite not being a strictly a technical limitation of the architecture, it could become a show-stopper to its practical use.
+\item \textbf{DP* Minimal Additional Code} Despite not being a strictly a technical limitation, the amount of code to be added and maintained to an existing architecture to meet functional requirements could become a show-stopper to its practical use.
 \end{itemize}
 In our analysis, we will consider not only the overall suitability of the architecture, but also the amount of effort needed to develop and deploy it.
-Table \ref{int:des} summarizes the previous principles and which class of requirements they address.
+Table \ref{int:des} summarizes the previous principles and the class of challenges they address.
 
 \begin{table}
-\caption{Design Principles for a Distributed IaaS Platform.\label{int:des}}
+\caption{DCI Design Principles.\label{int:des}}
 \begin{tabular}{lll}
 \toprule
-\textbf{Id} & \textbf{Design Principle} & \textbf{Class of Requirement} \\
+\textbf{Id} & \textbf{Design Principle} & \textbf{Class of Challenges} \\
 \midrule
@@
 DP5 & Compatibility with Legacy & Horizontal \\
-NDP1 & Minimal Additional Code & -- \\
+DP* & Minimal Additional Code & -- \\
 \bottomrule
@@
 % Intro + performance issues
-Several works presented in literature integrates these a subset of this design principles in their architecture, leveraging nested virtualization \cite{turtle:ibm,art:blan, cloudvisor:zhang}
-Cloud computing feature set could be enlarged exploiting two layers of virtualization instead of just one.
-The presence of two different hypervisors allows the possibility to specialize hypervisors functions.
-Instead of having a general purpose hypervisor we could split the hypervisor role in two different layers, interconnected with a well defined interface.
-Each layer will have to care about a subset of features, simplifying overall development.
-However, nested virtualization on x86 architecture faced impeding performance issues \cite{rec:virt}.
+\noindent A growing body of research has investigated building architectures and infrastructures integrating a subset of those design principles leveraging NV~\cite{turtle:ibm, hvx:usenix, Inception, art:blan, cloudvisor:zhang}.
+The distributed cloud feature set may be expanded using two levels of virtualization instead of a single one.
+This design allows to specialize hypervisor functions per layer, each layer addressing a subset of features, simplifying overall development compared to a GPH.
+NV development on x86 architecture was initially hindered by severe performance issues~\cite{rec:virt}.
+% Perfomance issuess adressed
+However, several works have shown feasibility of realizing low-overhead NV using dedicated implementation architectures~\cite{turtle:ibm}, or to recover performance loss exploiting multi-level consolidation~\cite{art:blan}.
+NV slowdown can now be considered an acceptable price for advanced features -- hardware manufacturers are also developing new hardware primitives to reduce this overhead~\cite{vmcs:nakajima}.
 
-% Perfomance issuess adressed
-However, several works demonstrate that it is possible to put in place low overhead nested virtualization using special approaches \cite{turtle:ibm} .
-Moreover, other works try to recover performance loss exploiting multi-level consolidation, introduced in nested virtualization \cite{art:blan}.
-In conclusion, nested virtualization slowdown could be considered nowadays an acceptable price to achieve advanced features and, in addition, hardware manifacturers are developing and proposing new hardware facilities to reduce this overhead \cite{vmcs:nakajima}.
+However, so far no architecture has addressed both vertical and horizontal set of challenges simultaneously.
+XenBlanket~\cite{art:blan} does not investigate cloud architectural security concerns (DP2, DP3), limiting its focus to providing new features on a coordinated multi-provider cloud.
+Cloudvisor~\cite{cloudvisor:zhang} proposed a NV-based security architecture, but cannot run more than one nested hypervisor at a time, which is a critical requirement for a multi-provider cloud platform (DP4).
+In the super-cloud vision~\cite{xclo:blank}, each user runs its own L1 hypervisor instance to realize  a blanket layer overcoming incompatibilities between different providers.
+However, security issues have not been addressed.
 
-To sum up, no architecture had been presented to address vertical and horizontal class of problems at the same time until now.
-XenBlanket does not investigate cloud architectural security concerns (DP3,DP2), limiting the focus on new potential feature of a coordinated multi-provider cloud.
-Cloudvisor proposed a security architecture that leverages nested virtualization without allowing the possibility of executing more than one nested hypervisor at time, that is a crucial requirements of a multi-provider cloud platform (DP4).
-In a SuperCloud scenario, each user has to run its own L1 hypervisor instance to enforce a compatibility layer, overcoming incompatibilities between different providers.
-
-Other approaches leveraging only one layer of virtualization does not achieve the whole design-principle satisfaction.
-Self-service cloud approach (SSC) \cite{ssc:art} design a complete security architecture to overcome major security and privacy issues of currently deployed cloud infrastructures, in order to alleviate the different needing user-centric and provider-centric vision of cloud computing presented in \ref{sec:intro}.
-Even if this architecture reproduces an approach similar to the SuperCloud without leveraging an extra layer of virtualization (DP4), the approach changes the behavior where the user acts introducing new components that have to be directly managed by the user.
-Therefore, the main drawback of this solution is to change the interface with the user, preventing the possibility to straightforwardly move to the new platform without having to change its applications (DP5).
+Other approaches based on a single layer of virtualization do not fulfill all principles either.
+Self-Service Clouds (SSC)~\cite{ssc:art} introduces a complete security architecture to overcome major security and privacy issues of currently deployed cloud infrastructures.
+The aim is to reconcile user-centric and provider-centric requirements of the cloud.
+Despite being similar to the super-cloud approach (DP4), SSC introduces new components that have to be directly managed by the user.
+This change of interface with user VMs prevents moving straightforwardly to the new platform without having to change its applications (DP5).
 
 % Integration NDSS Taming
-Furthermore, DeHype \cite{wu2013TamHosHypMosDepExe} presented a dissected version of KVM hypervisor, trying to reduce the TCB of this popular hypervisor.
-The approach consisted in demoting a great part of the hypervisor codebase to a user-level execution.
-DeHype satisfies all vertical design principles listed before (DS1,DS2,DS3), reinterpreting the microkernel approach defining  the greatest part of code to userspace.
-DeHype increases a lot the robustness of a cloud platform leveraging KVM, limiting for example the possibility of attack in multi-tenancy platform, but it does not address horizontal design principles about a multi-provider service.
+To reduce the TCB size, DeHype~\cite{wu2013TamHosHypMosDepExe} presented a dissected version of KVM, moving a great part of the code to user space.
+DeHype satisfies all vertical design principles, re-interpreting a micro-kernel design in a KVM setting.
+It increases robustness and security of a cloud platform leveraging KVM (DP1, DP2, DP3), but does not address horizontal challenges for a multi-provider service.
 
@@
 
-\subsection{Principle}
+\subsection{Approach}
 
-\noindent \textit{Nested virtualization (NV)} refers to a system architecture with two layers of virtualization~\cite{turtle:ibm}.
-It consists in ``virtualized'' virtualization: the guest system virtualizes a nested guest.
+\noindent \textit{Nested virtualization (NV)} refers to a system architecture with two layers of virtualization: the guest OS virtualizes a nested guest.
 The concept may be generalized to an arbitrary number of nested guest layers, leading to recursive virtualization~\cite{rec:virt,Popek:1974:FRV:361011.361073}.
@@
 Unfortunately, virtualization comes with a price due to the extra layer of indirection introduced: the computational overhead is generally acceptable when dealing with single-layer virtualization, but grows exponentially when adding further levels~\cite{rec:virt}.
-This huge performance overhead prevented NV adoption during many years.
-However, Ben Yehuda et al.
-finally proposed an architecture tending towards more reasonable overheads~\cite{turtle:ibm}.
+This  performance overhead prevented NV adoption during many years until was finally proposed an architecture tending towards more reasonable overheads~\cite{turtle:ibm}.
 
-NV feasibility opened plenty of new possibilities and avenues for research that have been largely discussed in the literature~\cite{turtle:ibm,rec:virt,art:blan}.
+NV feasibility opened plenty of new possibilities and avenues for research that have been largely discussed in the literature.
 Such works all tend to diversify hypervisor functionalities, introducing new features but also keeping existing ones.
-It turns out that each layer addresses clearly different sets of issues.
-
-L0 tackles vertical aspects of the platform.
-It aims to guarantee the best performance and strongest level of isolation possible.
+Each layer clearly addresses different sets of issues:
+\begin{itemize}
+\item L0 deals with vertical aspects of the platform.
+It aims to guarantee the best performance and strongest level of isolation.
 L0 represents the last line of defense of the platform and a privileged point for monitoring of its global status.
-
-L1 is responsible for horizontal feature set.
-It has to provide the widest possible support for different platforms and the ability to virtualize using different techniques e.g., para-virtualization, hardware-assisted virtualization, dynamic binary translation.
-{\sffamily To address such requirements, Williams et al.~\cite{art:blan} proposed a patched version of Xen, Xen Blanket, to realize an L1 virtualization layer designed to run over different hypervisors.
+\item L1 deals with the horizontal feature set.
+It should provide the widest possible support for different platforms and the ability to virtualize using different techniques e.g., para-virtualization, hardware-assisted virtualization, dynamic binary translation.
+\end{itemize}
+%{\sffamily To address such requirements, Williams et al.~\cite{art:blan} proposed a patched version of Xen, Xen Blanket, to realize an L1 virtualization layer designed to run over different hypervisors.
 This notably enables to run and to perform live migration of the same VM over different physical cloud infrastructures.} % A migrer dans le related work?
@@
 
-\noindent Today, NV presents several limitations related to hardware support {\sffamily and software state of implementation}.
+\noindent Today, NV presents several limitations related to hardware support and TCB size.
 
@@
 HAV primitives assist instruction emulation, memory address translation, and management of devices.
-Both Xen and KVM propose {\sffamily nested VMX/SVN features} that allow to efficiently handle nested hardware virtualization on single-layer hardware primitives.
+Both Xen and KVM propose nested VMX/SVM features that allow to efficiently handle nested hardware virtualization on single-layer hardware primitives.
 Full virtualization is currently the only virtualization technique supported for L1 virtualization due to its ability to run unmodified guest OSes -- and thus hypervisors.
-Paravirtualization (PV) would on the contrary require an ad hoc modified hypervisor to act as L1 (thus violating the NDP1 principle).
+Paravirtualization would on the contrary require an ad hoc modified hypervisor to act as L1 (thus violating DP*).
 %To address this performance problems, several hardware producers proposed new facilities, like Intel VMCS shadowing, that reduces the amount of nested VMEXIT when handling a nested guest VMEXIT.
@@
 Several works discuss and analyze the overall TCB of modern GPHs, always bigger than 100 KLoC~\cite{nova, xmhf}.
-This approach directly violates the DP3 design principle.
+This approach directly violates the DP3 principle.
 %that identified a small TCB as a major requirement for a solid architecture.
@@
 Moreover, the previous architecture does not improve overall stability of the infrastructure.
-The bottom-most hypervisor L0 faces the same issues as single-layer GPH architectures: a large TCB implies also that a failure experienced in one of its critical components may trigger a generalized failure of the whole platform (violating the DP2 principle).
+The L0 hypervisor faces the same issues as single-layer GPH architectures: a large TCB implies also that a failure experienced in one of its critical components may trigger a generalized failure of the whole platform (violating the DP2 principle).
 Besides, in standard GPH architectures, the administrator has the entire control of the platform and could potentially snoop and inspect user VM status and data (violating the DP1 principle).
-Such designs notably do not take advantage of the possibility {\sffamily to use a platform compatibility layer}, and more modular architectures with only a tiny security-critical kernel.
+Such designs notably do not take advantage of the possibility  to use more modular architectures with only a tiny security-critical kernel.
 
-Overall, current NV architectures lack some of the dedicated features found in modern hypervisors and the impossibility to rely on a tiny TCB (violating principles DP3 and DP4).
+Overall, current NV architectures lack some of the dedicated features found in modern hypervisors and the impossibility to rely on a tiny TCB (violating principles DP4 and DP3).
 If the first one is currently being addressed by hypervisor developers, the second simply has not be achieved with modern hypervisors so far.
@@
 
-\subsection{Principle}
+\subsection{Approach}
 
-\noindent To solve the TCB size issue, several works have introduced the idea of \textit{micro-hypervisor (MH)} architecture~\cite{nova, NoHype, xmhf}.
-To reduce the hypervisor size, such new hypervisor architectures draw inspiration from evolution in OS architecture (e.g., extensible, micro-, exo-, component-based kernels).
+\noindent To solve the TCB size issue, the idea of \textit{micro-hypervisor (MH)} architecture was introduced~\cite{nova, NoHype, xmhf}.
+Such hypervisor design draws inspiration from evolution in OS architecture (e.g., extensible, micro-, exo-, component-based kernels).
 Within the hypervisor are distinguished security-critical modules from non-sensitive code.
-The aim is to expel as much code as possible from the TCB and target ultra-thin hypervisors.
-For instance, some components of the hypervisor (e.g., device drivers, VM address space management) may be virtualized outside the micro-hypervisor, isolated, and restarted in case of compromise~\cite{nova}.
-The extreme case is to remove the virtualization layer altogether~\cite{NoHype}.
+The aim is to expel as much code as possible from the TCB make the hypervisor ultra-thin.
+Some hypervisor components (e.g., device drivers, VM address space management) may be virtualized outside the MH, isolated, and restarted in case of compromise~\cite{nova}.
+The utmost MH removes the virtualization layer altogether~\cite{NoHype}.
 
-with a traditional microkernel-like design that tries to consistently reduce the amount of code running with the highest privilege level.
-
-A micro-kernel traditionally consists in a minimal kernel running in privileged mode, offering only a few fundamental services to other modules (e.g, scheduling, IPC, memory management) running in a less privileged domain.
+%MH = MK+ H.
+MH architectures may be seen as convergence of traditional hypervisor and micro-kernels (MK).
+%1) MK
+%Perspective: minimalisme: 
+%with a traditional microkernel-like design that tries to consistently reduce the amount of code running with the highest privilege level.
+From the MK design principles legacy, the MH is guided by TCB minimization. 
+%-- Définition du mk
+%A micro-kernel traditionally consists in a minimal kernel running in privileged mode, offering only a few fundamental services to other modules (e.g, scheduling, IPC, memory management) running in a less privileged domain.
 All other components, such as device drivers, are moved out in user-space.
-While the benefits of such a design has been questioned due to the cost of IPCs in first generations of micro-kernels, such arguments have since then been completely dispelled~\cite{liedtke1993ImpIPCbyKerDes,liedtke1995micConnova}.
+%-- Définition du mh:
+%Therefore, leveraging this concept of microkernels, a micro hypervisor (MH) is composed from a small kernel providing only essential functions, like scheduling, IPC, memory management and VMX extension.
+Thus, it features a minimal kernel running in privileged mode, offering only a few fundamental services (e.g, scheduling, IPCs, memory management, VMX extensions) to other modules running in less privileged domains.
+All other components, such as device drivers, are moved out in user-space.
+% Skip inconvénients du MK
+%1) Inconvénients du mk: performance.
 %Microkernels had been questioned because of the wasteful performance of first generations, due to IPC intrinsic cost.
 This argument had been completed confuted by Liedtke in 1993: \cite{liedtke1993ImpIPCbyKerDes,liedtke1995micCon}.
+%While the benefits of such a design has been questioned due to the cost of IPCs in first generations of micro-kernels, such arguments have since then been completely dispelled~\cite{liedtke1993ImpIPCbyKerDes,liedtke1995micConnova}.
+%-- Bénéfices du mk
+The main benefit of such increased modularity is
+% a) Bénéfices du mk: flexibility in privilege
+% One of the benefits of micro-kernels is to enable to easily redesign a hierarchy of privileges. 
+%
+%The different groups may hold correspond to different trust relationships of principals running Control is finer than in the hypervisor, both by the granularity of isolation (lightweight threads), and by a unique control point of RT behavior in the scheduler.
+Device drivers are run in , allowing great flexibility for device sharing patterns.
+to easily define groups of user-space services, each group sharing a common a set of privileges.
+Trust relationships between groups may be arbitrarily defined, mirroring those between sets of principals or applications sharing the system.
+%For instance, a user application normally trusts the micro-kernel and only the set of user-space services it requires.
+However, such drivers and libraries may not be trusted by every application run on the system.
+User applications share the system but may need to support completely different set of system services.
+% -- Les personalités:
+% In addition, MH inherits the concept of micro-kernel to implement different ``personalities''. 
+This allows MK to implement different ``personalities'': device drivers being run in user mode, such fine-grained control allows great flexibility to implement device sharing patterns.
+%
+%2) H
+%
+%Plus facile dans un contexte d'hyperviseurs:
+%However, the interest of such ``personalities'' has been largely reduced by the fact that system services, device drivers and even applications have to be completely re-designed to fit this new architectural pattern.
+%
+%Therefore, with the introduction of VMs, microkernels lost the battle in favor of hypervisors, in terms of popularity.
+While the interest of realizing such ``personalities'' in traditional MKs somewhat lessened as system services, drivers, and even applications had to be completely re-designed (hypervisors offering better primitives for abstraction with the notion of VM),
+%
+%
+%3) MH.
+%
+such flexibility still remains a major benefit of coupling micro-kernel design with virtualization~\cite{OKL4}.
 
+%1) Pas de tiny TCB
+%However, the great advantage of tiny TCB, provided by microkernel, was not provided also by hypervisors. 
+%2) Architecture monilithic
+%As we analyzed before, they present normally a monolithic architecture, like the great part of modern OS. 
+%This feature kept alive the interest in microkernels, in particular trying to couple microkernel benefits with virtualization
+
+%-- Exemples: Several micro-hypervisor have been proposed in literature \cite{xmhf, nova}. 
+%
+%=> -- MH = MK+ virtualization. They rely on the micro-kernel paradigm, extending it to virtualization. MV
+
+%-- Relation entre MH et principes: 
+%* MH naturally satisfy vertical design principles (DP1,DP3, DP4). 
+MHs are particularly well adapted to sastisfy vertical design principles.
+% Security:
+%Pourquoi: 
+For instance, MH design increases reliability as services crashing in user space may be restarted without bringing down the entire system (DP2).
+%Fail-safe principle:
+%Moving Finally, this extraction process drives to a more solid platform, considering that user-space crashes are generally not fatal for the entire system integrity. 
+The attack surface is also lowered by the small TCB size, the minimal kernel only containing security-critical components and forming a last line of defense (DP3).
+%The tiny core could be considered the only system-critical componente of the architecture and, due to its relative small dimension and the consequent minor attitude to vulnerabilities,it could represent the ``last line of defense''.
+MH might also address some horizontal principles. 
+%Interop principle:
+%Qu'est ce que c'est:
+``Personalities'' are concretely a set of processes handling specific functions running in user-space.
+Implementing such features outside the kernel makes it easier for each user to rely on different instances of such personalities running concurrently, possibly in a distributed manner (DP4).
+%This process separation between users could enforce a straightforward mechanism to neutralize VM-to-VM attack from a malicious user.
+Moreover, user-space processes are easier to debug with commodity tools (GDB, valgrind).
+
+% Enlever les exemples qui ne sont pas significatifs.
+% A possible example, that will be analyzed in next section, is the user-space VMM component of Nova, Vancouver, that resides in user-space \cite{nova}.
+% In addition, this approach is not far from the one proposed in DeHype \cite{wu2013TamHosHypMosDepExe}, where a consistent part of the KVM hypervisor has been demoted in user-space.
+
+% Texte additionnel sur les MH:
 %Architectural solutions were also created to make the hypervisor more minimal and flexible, targetting its two most vulnerable elements: (1) the core hypervisor itself; and (2) the management VM (e.g., Xen Dom0).
@@
 They are thus hard to apply to mainstream hypervisors today.
-One of the benefits of micro-kernels is to enable to easily redesign a hierarchy of privileges.
-A user application normally trusts the micro-kernel and only the set of user-space services it requires.
-However, such drivers and libraries may not be trusted by every application run on the system.
-User applications share the system but may need to support completely different set of system services.
-This allows micro-kernels to implement different ``personalities''.
-However, the interest of such ``personalities'' has been largely reduced by the fact that system services, device drivers and even applications have to be completely re-designed to fit this new architectural pattern.
-Therefore, with the introduction of VMs, microkernels lost the battle in favor of hypervisors, in terms of popularity.
-However, the great advantage of tiny TCB, provided by microkernel, was not provided also by hypervisors.
-As we analyzed before, they present normally a monolithic architecture, like the great part of modern OS.
-This feature kept alive the interest in microkernels, in particular trying to couple microkernel benefits with virtualization\cite{Heiser}.
 
-Therefore, leveraging this concept of microkernels, a micro hypervisor (MH) is composed from a small kernel providing only essential functions, like scheduling, IPC, memory management and VMX extension.
-Several micro-hypervisor have been proposed in literature \cite{xmhf, nova}.
-They rely on the micro-kernel paradigm, extending it to virtualization.
-MH naturally satisfy vertical design principles (DP1,DP3, DP4).
-The tiny core could be considered the only system-critical componente of the architecture and, due to its relative small dimension and the consequent minor attitude to vulnerabilities,it could represent the ``last line of defense''.
-
-In addition, MH inherits the concept of micro-kernel to implement different ``personalities''.
-Those ``personalities'' are concretely a set of ``processes'' handling specific functions running in user-space.
-As those functions are implemented outside the kernel, each user could rely on a different instances of the hypervisor user-space process, that could coexist at the same time.
-This process separation between user could enforce a straightforward mechanism to neutralize VM-to-VM attack from a malicious user.
-Moreover, user-space processes are easier to debug with commodity tools (GDB, valgrind).
-Finally, this extraction process drives to a more solid platform, considering that user-space crashes are generally not fatal for the entire system integrity.
-A possible example, that will be analyzed in next section, is the user-space VMM component of Nova, Vancouver, that resides in user-space \cite{nova}.
-In addition, this approach is not far from the one proposed in DeHype \cite{wu2013TamHosHypMosDepExe}, where a consistent part of the KVM hypervisor has been demoted in user-space.
-
-\subsection{Limitations}
-
-\noindent Despite the positives, MH approaches had demonstrated some limitations.
-Firstly, the straightforward one is the initial lack of hardware support, lack of device drivers are still a relevant Achilles' heel for this solutions, inherited from its ancestors microkernels.
-Fortunately, this kind of issue could be addressed leveraging I/O passthrough techniques, although it will cause exclusive allocation issues.
-
-Moreover, as reported in previous sections, several cloud computing issues are simply not addressable with general purpose hypervisor (GPH).
-However, they are commonly widely adopted and represent the bedrock of existing platform.
-Therefore, GPH exports wide and solid interfaces leveraged by a plenty of different software that is able to properly communicate and act through them on the virtualization layer.
-On the contrary, MH lack of cloud toolkit integration, requiring an explicit support to be written, not satysfing 3rd part compatibility requirements (DP5).
-GPH are tightly connected with cloud toolkit and could present solid interfaces to communicate with different tools.
-
-Finally, microkernel modularization introduce a major challenge in architecture design.
-Due the increased complexity, for sake of kernel size, MH could generally lack of several virtualization features, commonly available on modern GPH hypervisor.
-For instance, XMHF does not support multiple guests.
-This kind of drawbacks could limits the potential interest in MH, even they are able to introduce several important features.
-
-\section{Architecture Comparison}
-\label{sec:archcomp}
+% Texte sur les MK, H, et MV
+%Embedded virtualization brings benefits to achieve several such properties, for instance abstraction and isolation by multiplexing and isolating VMs.
+However, current hypervisor architectures show strong limitations for many others.
+A main flaw is the huge TCB size, leading to reliability, security, and administration issues.
+Flexibility is rather poor, as for device sharing.
+RT guarantees are also difficult due to two-level system scheduling, i.e., threads at VM-level, and VMs at hypervisor-level.
+Finally, resource control remains coarse-grained through heavyweight VMs.
+%
+%\textit{Micro-kernel} design has thus been increasingly popular to meet such challenges.
+Reducing the kernel to its fundamental elements, and making the rest of the OS deprivileged outside results in an extremely minimal TCB that may be formally verified.
+Strong isolation and high-perfromance communication between address spaces is also achieved through very efficient IPCs.
+Micro-kernels increasingly support virtualization: privileged instructions are intercepted and emulated in a dedicated system service.
+Control is finer than in the hypervisor, both by the granularity of isolation (lightweight threads), and by a unique control point of RT behavior in the scheduler.
+Device drivers are run in user space above the micro-kernel, allowing great flexibility for device sharing patterns.
+%
+%Such architectures are very similar, although the hypervisor is driven by abstraction, to match closely virtual to real resources, while micro-kernels are guided by TCB minimization.
+%They tend to converge: the \textit{micro-visor} allows to run both VMs and threads over a lightweight execution environment~\cite{OKL4}.
+Like micro-kernels, it has a minimal TCB, and high-performance asynchronous IPCs between VMs, applications, or user-mode drivers.
+Like hypervisors, it allows virtualization of CPU, memory, or I/O through a hardware abstraction layer (HAL).
+Arbitrary control over resource partitioning in security domains or RT scheduling of threads or VMs is also possible.
+Finally, the micro-visor offers great architectural flexibility: drivers may be virtualized in a VM, or executed as system services in their own address space.
 
@@
 
+\subsection{Limitations}
+
+\noindent The MH approach also showed several limitations.
+First, support for hardware is lacking, notably in terms of device drivers.
+Fortunately, this issue may be addressed leveraging I/O passthrough techniques, despite some resulting exclusive allocation challenges.
+Second, several cloud issues are simply not addressable with GPHs.
+However, GPHs are now widely adopted and foundation of existing cloud platforms.
+While a GPH exports mature APIs to upper software layers and is also tightly integrated with cloud toolkits, MHs lack such API stability and integration, requiring explicit dedicated code to be written -- thus violating DP5.
+Third, for the sake of kernel size, MHs could lack several virtualization features commonly available on modern GPH hypervisors.
+For instance, XMHF does not support multiple guests.
+Such limitations applicable for MH architectures alone may be lifted partially or totally when applying MH design to NV architectures.
+
+\section{Architecture Comparison}
+\label{sec:archcomp}
+
+\begin{table*}[htbp]
+\caption{Architecture Comparison.\label{fin:conf}}
+\centering
+\begin{tabular}{llcllll}
+\toprule
+Id & Functional Design Principle & Class of Requirement & GPH/GPH & MH/MH & MH/GPH & GPH/MH\\
+\midrule
+DP1  & User-Data Privacy               & Vertical   & $\times$      & $\checkmark$  & $\times$/$\checkmark$ & $\checkmark$ \\
+DP2  & Fail-Safe modular architecture  & Vertical   & $\times$      & $\checkmark$  & $\times$ & $\checkmark$ \\
+DP3  & Small TCB                       & Vertical   &  $\times$     & $\checkmark$  & $\times$ & $\checkmark$ \\
+DP4  & Inter-Platform support          & Horizontal & $\checkmark$  & $\times$      & $\times$ & $\checkmark$ \\
+DP5  & Compatibility with Legacy       & Horizontal & $\checkmark$  & $\times$      & $\times$ & $\checkmark$ \\
+DP* & Minimal Additional Code         & --         & $\checkmark$  & $\times$      & $\checkmark$ & $\checkmark$ \\
+\bottomrule
+\end{tabular}
+\end{table*}
+
 \noindent Adding an extra layer of virtualization enables to introduce a wide range of new features in the IaaS infrastructure, but does not reduce the TCB size.
-As already discussed, introducing MH design within a nested IaaS architecture could represent a new approach to improve NV benefits by overcoming such limitations.
+Introducing MH design within a nested IaaS architecture could represent a new approach to improve NV benefits by overcoming such limitations.
 In what follows, we analyze the four possible architectures shown in Figure~\ref{fig:io} to achieve nesting of GPH and MH, highlighting their potential, and assessing their benefits and weaknesses w.r.t.
@@
 
-\noindent This architecture is considered only to highlight improvements brought by a MH design compared to a GPH/GPH architecture.
+\noindent This architecture is considered only to highlight the improvements of a MH design compared to a GPH/GPH architecture.
 An MH/MH implementation would meet the vertical requirements missing in the previous approach.
@@
 
-\noindent The third architecture aims to alleviate security concerns. It combines a GPH as L0 and a MH as L1. 
+\noindent The third architecture aims to address security concerns. It combines a GPH as L0 and a MH as L1. 
 %Several MH are proposed in literature focusing on different .
@@
 However, there is no TCB reduction as the code size running in L0 kernel space is as big as before. 
-Besides, adopting a MH as L1 prevents using inter-platform enhanced features.
-As already seen for MH design, the price of having a thin hypervisor is normally paid with a lack of certain features.
-For example, XMHF only supports a single guest~\cite{xmhf}.
-Thus, this architectural approach cannot provide any new horizontal features leveraging NV.
+Besides, adopting a MH as L1 prevents using inter-platform enhanced features. As already seen for MH design, the price of having a thin hypervisor is normally paid with a lack of certain features. 
+%For example, XMHF only supports a single guest~\cite{xmhf}. 
+Thus, this architecture cannot provide any new horizontal features leveraging NV.
 Therefore, this approach can only provide the same features as those deployed today, with an increased but not sufficient level of isolation.
@@
 \noindent This last approach achieves a smaller TCB than other architectures, due increased modularity brought by the micro-kernel-like design of the L0 layer -- satisfying DP3.
-Leveraging a GPH like Xen or KVM for the L1 layer provides the user with an environment looking identical to those used in production today -- satisfying DP5.
+Leveraging a GPH like Xen or KVM for the L1 layer provides the user with an environment seeming identical to those used in production today -- satisfying DP5.
 The L1 GPH is normally integrated with modern cloud toolkits like OpenStack, and will thus support third-party applications -- satisfying DP4.
@@
 Drivers are a major MH issue, due to costs of development or of adaptation to a new architecture.
-Such challenges may be addressed using device assignment: the L1 GPH hypervisor manage the physical device with the proper driver now running in a lower level of privilege.
+Such challenges may be addressed using device assignment: the L1 GPH hypervisor manages the physical device with the proper driver now running in a lower level of privilege.
 To be workable, this approach would require exclusive allocation to be simply implementable, e.g., using SR-IOV~\cite{sriov}.
-
-\begin{table*}
-\caption{Architecture Comparison.\label{fin:conf}}
-\centering
-\begin{tabular}{llcllll}
-\toprule
-Id & Functional Design Principle & Class of Requirement & GPH/GPH & MH/MH & MH/GPH & GPH/MH\\
-\midrule
-DP1  & User-Data Privacy               & Vertical   & $\times$      & $\checkmark$  & $\times$ & $\checkmark$ \\
-DP2  & Fail-Safe modular architecture  & Vertical   & $\times$      & $\checkmark$  & $\times$ & $\checkmark$ \\
-DP3  & Small TCB                       & Vertical   &  $\times$     & $\checkmark$  & $\times$ & $\checkmark$ \\
-DP4  & Inter-Platform support          & Horizontal & $\checkmark$  & $\times$      & $\times$ & $\checkmark$ \\
-DP5  & Compatibility with Legacy       & Horizontal & $\checkmark$  & $\times$      & $\times$ & $\checkmark$ \\
-NDP1 & Minimal Additional Code         & --         & $\checkmark$  & $\times$      & $\times$ & $\checkmark$ \\
-\bottomrule
-\end{tabular}
-\end{table*}
 
@@
 
-\noindent Overall, how do the architectures above compare w.r.t.
-to the two main reported benefits of NV, i.e., vertical, through an increased level of isolation between hypervisor and customer VMs, and horizontal to easily achieve multiple provider platform interoperability at system level?
+\noindent Overall, how do such architectures compare to the two main reported benefits of NV, i.e., vertical, through an increased level of isolation between hypervisor and customer VMs, and horizontal to easily achieve multiple provider platform interoperability at system level?
 Table~\ref{fin:conf} summarizes the previous analysis.
@@
 However, the effort to set up a GPH/MH is strongly dependent on the technique for L2 virtualization.
-As already seen, to support L2 guests with hardware-assisted virtualization which is becoming mainstream, the MH needs to export nVMX (nested VMX) primitives.
+As already seen, to support L2 guests with hardware-assisted virtualization which is becoming mainstream, the MH needs to export nested VMX primitives.
 On the other end, adopting paravirtualization or dynamic binary translation for L2 does not requires particular hardware support, and thus would not require MH patching to set up a running platform.
-However, PV remains limited by supporting only a subset of system features which could be an issue as a general solution.
+However, paravirtualization supports only a subset of system features, which could be an issue as general solution.
 %The last point of interest is related to device management.
@@
 The main remaining barrier is to come up with an MH implementation supporting HAV nested guests.
-We are currently working on a GPH/MH implementation based on Xen over NOVA aiming to lift this limitation using PV.
+We are working on a GPH/MH implementation aiming to lift this limitation using paravirtualization.
 
-\section{Security Architecture}
-\label{ref:arch}
+\section{Towards 2D Security Management}
+\label{sec:arch}
 
-In the previous sections we proposed a functional analysis of cloud architecture in order to respect design principles that we identified in the introduction.
-In this section, we propose a GPH over MH implementation, based on .
+\begin{figure*}[htbp]
+\begin{center}
+\includegraphics[width=10cm]{DCCArchitecture.png}
+\vspace*{-0.5cm}
+\end{center}
+\caption{DCI System Architecture.}
+\vspace*{-0.5cm}
+\label{fig:archisecu}
+\end{figure*}
 
-\begin{itemize}[]
-\item \textbf{Inter-platform support} Executed as L1, Xen would virtualize its nested guest initially using PV delivering the compatibility layer.
-\item \textbf{Tiny TCB} The NOVA kernel will represent the only part of the platform executed at the highest level of privilege.
-Even the VMM, Vancouver, will be executed in the L0 user-space.
-The micro-kernel in L0 enforces a strictly modular approach to L0.
-Therefore, the VMM will present distinct processes instances for each L1 guest, enforcing a context division.
-\item \textbf{3rd part compatibility} From a user point of view the architecture still exports the same interfaces provided by the Xen hypervisor, allowing the user to keep its applications unchanged.
-\item \textbf{Fail-safe components architecture} Moreover, the tiny core is the only safety-critical software component.
-The system could potentially be able to recover any crash in different user-space components.
+\noindent We now sketch how the previous enhanced-NV design enables to build a DCI security architecture meeting both vertical and horizontal requirements.
+Such DCIs will face 2D threats, vertical spanning layers, and horizontal, spanning domains -- DCIs being both multi-layer and multi-domain.
+Vertically, attacks currently target several infrastructure layers (e.g., VMs, hypervisor).
+Unfortunately, most existing defenses are generally for single layers only.
+This makes it difficult to grasp the overall extent of an attack.
+Horizontally, threats may propagate through the network from cloud to cloud.
+Unfortunately, most existing defenses are for single clouds only.
+Lack of interoperability in security policies and mechanisms is thus a major barrier to  unified hybrid cloud security.
+Overall, DCI security requires 2D protection.
+
+%3 Layer infrastructure
+We consider the 3-layered DCI architecture shown in Figure~\ref{fig:archisecu}. The bottom and top layers are similar to a single-cloud IaaS.
+\begin{itemize}
+\item
+%Provider:
+At the bottom is the \textit{provider-centric} view of the DCI: each cloud provider contributing to the DCI is in control of compute, storage, and network resource provisioning.  
+%Top User-VM
+% of the cloud each provider providing layer and t
+% On the top are the VM built self-service security where security is personalized/customized.
+\item At the top is the \textit{user-centric} view of the DCI: customer VMs are arranged in a set of \textit{self-service clouds} (SSCs) where security and other services may be fully customized by users, independently from underlying providers.
+%Middlelayer: Supercloud
+\item The middle layer is a distributed resource abstraction and security management plane we call the \textit{security super-cloud}.
 \end{itemize}
+
+%Super-cloud 
+%The SUPERCLOUD will fulfill that vision with a two-level infrastructure:
+The super-cloud has two main functions.
+% Fonction North
+% Fonctions 1: SUPERCLOUD Layer built SS cloud
+% implementing the self-service security to build SSChe lower part of the SUPERCLOUD is be a distributed infrastructure will realize the distribute resource abstraction layer  and .
+It will implement the northbound, user-centric super-cloud API.
+%The northbound, user-centric super-cloud API will implement self service security and dependability, allowing to build self-service clouds (SSC) where VM, network, and storage security may be personalized independently from the underlying provider.
+Such clouds may be defined at the service level to the full IaaS, relationships between SSC being flat or nested, also with complex trust relationships between sets of SSC according to the ecosystem relationships of the super-cloud providers providing the SSC.
+Its northbound interface federates cloud resources in a provider-independent manner and enables to customize their security to realize the SSCs.
+While many orchestration architectures have been proposed at service or middleware levels (e.g., multi-Iaas platforms), our architecture aims to achieve federation at system-level to fully take advantage of underlying provider platform openness (if available) for finer-grained control and increased flexibility.
+% Fonction South
+% Fonction  2:
+% The southbound, provider-centric super-cloud API will define a unified security and dependenability management plane of the compute, data, and network resources independently from providers.
+This may apply for B2B use cases to management of clouds of clouds but also to device-to-cloud scenarios if devices are considered, with possible extensions to the personal cloud for device-to-device scenarios, the SSC taking the form of personal cloud-based digital infospheres where security and dependability may be totally user-controlled.
+% Also security management both vertically and horizontally.
+% This means: providing vertically a cross-layer integrated vision of security of the cloud and SUPERCLOUD infrastructure; defining horizontally a security abstraction layer to achieve a homogeneous level of security across clouds independently from underlying provider mechanisms; and integrating both dimensions smoothly.
+This approach is key to providing a unified view of security of the SUPERCLOUD.
+
+% bove, the upper part of the SUPERCLOUD will be a supervision infrastructure providing a set of security and resilience services to guarantee the security of the overall SUPERCLOUD architecture.
+It will realizing self-service security, end-to-end security, and resilience.
+It will implement the southbound, provider-centric super-cloud API.
+
+The super-cloud southbound interface is in charge of 2D DCI security management.
+It provides a unified vision of security across layers of the DCI, guarantees a homogeneous level of security across clouds independently from underlying provider mechanisms, and integrates both dimensions smoothly.
+% Automation
+% Pour l'intégration: requirement: autonomous
+% Automation.
+%Despite some automated solutions for managing security of single clouds, there is currently no solution truly applicable to federations of clouds in practice.
+% and security management complexity. Th
+% 360 Atuomation
+Despite many defense mechanisms being available in this 2D landscape, they remain difficult to configure and to reconcile, the current ``by hand'' approach for supervision being a nightmare for security administrators.
+To overcome such complexity, we consider here a security monitoring infrastructure providing 2D detection and reaction to threats with a high degree of automation.
+
+%OS:
+In terms of system architecture, we consider a NV-based design to realize the super-cloud above the DCI as in~\cite{xclo:blank}. 
+%Layer 1: MH Ce que cela permet
+% Provides enhanced security
+% Achieve Vertical security Cross Layer security
+We assume the provider-layer to be MH-based. 
+This design choice provides a solid foundation to meet all vertical features, notably smaller TCB size and enhanced DCI security such as guaranteeing VM security even if intermediate layers are compromised.
+Note that this assumption does not seem unreasonable for the coming years as: 
+%Provider MH: Assumption min V, XenDisagregated...
+(1) despite most current IaaS platforms still being GPH-based, there is a strong trend towards making the hypervisor more minimal and flexible, as witnessed by disaggregation of hypervisors of the mainstream IaaS platforms~\cite{XOAR,MinV};
+% Interoperability betwene differnet component-based design such as XMHF.
+(2) MHs themselves are becoming component-based, with possibility of interoperability to avoid any further risk of IaaS lock-in as shown by first multi-MH  OS frameworks~\cite{xmhf}. 
+% Above GPH: 
+% 
+% GPH allows also to layer of interoperability horizontal blanket layer above the different MH to enalbe self-service security.
+% Layer 2: GPH Ce que cela permet. Verticalement et horizontalement.
+% % Blanket layer .
+
+The super-cloud layer is GPH-based, both to be able to realize the SSCs by defining blanket layers and to guarantee security interoperability and other horizontal features.
+
+%Framework VESPA horizontal UCC
+% Cross-Domain Framework:
+% ou VERTICAL ICAC.
+% Cross-Layer Framewokr: VESPA
+To implement 2D security management proper, we intend to rely on self-protection frameworks which have been defined for both cross-layer~\cite{VESPA} and cross-domain~\cite{VESPA:UCC} IaaS and multi-IaaS integrated security monitoring.
+We are currently investigating how they may be integrated within the previous system architecture, notably to guarantee security SLAs~\cite{CUMULUS}.
+
+%The SUPERCLOUD project aims to define the architecture and build the infrastructure to fulfill the vision of user-centric secure and dependable management of a clouds of clouds.
+Our approach will be to define a new distributed architectural plane, the super-cloud providing an end-to-end interface both between user-centric and provider-centric views of multiple clouds.
+Its role will be both to provide a distributed resource abstraction and flexible but unified control for management of security and resilience.
+%From the user perspective, the super-cloud will enable user-centric security and dependability through a unified distribution layer for cloud resources, independently from their type (computation, storage, and network) and from underlying cloud providers.
+It thus forms an intermediation layer for resources between their production by providers and their consumption by users.
+This in line with the vision introduced by IBM [X] as well as with the utility computing paradigm [X], but will go well beyond, leveraging this distribution layer to explore high customizability of user cloud security and dependability according to well-defined SLAs.
+%From the provider perspective, the super-cloud will also enable provider-independent control over security and resilience over the overall distributed cloud infrastructure, security being autonomous managed and guaranteed end-to-end.
+This will also make the resource distribution layer both secure and resilient.
+%The level of abstraction, from system to middleware interoperability of the super-cloud be also be tuned to allow trade-offs between user-centric and provider-centric control for security services. 
+%In terms of architecture:
+%The northbound, user-centric super-cloud API will implement self service security and dependability, allowing to build self-service clouds (SSC) where VM, network, and storage security may be personalized independently from the underlying provider.
+Such clouds may be defined at the service level to the full IaaS, relationships between SSC being flat or nested, also with complex trust relationships between sets of SSC according to the ecosystem relationships of the super-cloud providers providing the SSC.
+%The southbound, provider-centric super-cloud API will define a unified security and dependenability management plane of the compute, data, and network resources independently from providers.
+This may apply for B2B use cases to management of clouds of clouds but also to device-to-cloud scenarios if devices are considered, with possible extensions to the personal cloud for device-to-device scenarios, the SSC taking the form of personal cloud-based digital infospheres where security and dependability may be totally user-controlled.
+%...
+%The SUPERCLOUD will fulfill that vision with a two-level infrastructure:
+%•	The lower part of the SUPERCLOUD is be a distributed infrastructure will realize the distribute resource abstraction layer federating distributed cloud resources and implementing the self-service security to build SSC.
+It will implement the northbound, user-centric super-cloud API.
+%•	Above, the upper part of the SUPERCLOUD will be a supervision infrastructure providing a set of security and resilience services to guarantee the security of the overall SUPERCLOUD architecture.
+It will realizing self-service security, end-to-end security, and resilience.
+It will implement the southbound, provider-centric super-cloud API.
+%In what follows, we describe the SUPECLOUD vision for each self-service security, self-managed security, end-to-end security and resilience properties. 
+%
+%For computing workloads, user-centric cloud security management faces two challenges.
+%•	Lack of IaaS infrastructure openness: provider-controlled immutable hypervisors (or that mask access to the hardware) prevent users from taking advantage of hypervisor-level advanced resource management and security services such as high availability or VM introspection for fine-grained VM misbehavior detection and reaction.
+Such examples of vendor lock-in are already a notorious issue for single provider infrastructures [CSA Guidance].
+They are a real barrier for hybrid clouds: the varying degree of openness of the underlying IaaS forces to achieve security interoperability in upper software layers instead of at system level as in multi-IaaS orchestration platforms, highly reducing user control over security management.
+Such limitations also prevent fine-grained trade-offs between provider and user control over security : among security services, users cannot choose between those to customize, and those to delegate to the provider infrastructure.
+%•	Security threats for customer VMs: large, complex, overprivileged hypervisors are a serious risk for VMs [CSATT].
+The assumption of a trusted hypervisor is no more, illustrated through many examples of VM escapes, breaking hypervisor isolation through exploitaiton of device drivers vulnerabilities [Virtunoid, CloudBurst].
+A multi-layer approach to IaaS security is needed considering lower layers, to guarantee VM security even if the hypervisor layer is compromised.
+This challenge is intrisically linked with openness and flexibility issues: from a user perspective, to protect their VMs from an untrusted provider infrastructure, customers may wish adding their own security modules (or layer) in the infrastructure, provided it is open enough.
+Conversely, a provider may also wish to benefit from an open hypervisor architecture to harden its IaaS against rogue VM behaviors.
+%Earlier IaaS architectures such as disaggregated hypervisors or nested virtualization provided partial answers to either of those issues but not both.
+Hypervisor disaggregation (for the Dom0 or the core hypervisor) achieves flexibility through the definition of clearly separated access control domains between provider-centric and user-centric security management planes where fine-grained trade-offs may be reached [SSC].
+This approach is compatible with legacy hypervisors (Dom0 case) or requires new architectures (core hypervisor case).
+However, it only addresses security and flexibility challenges in a single cloud context, but not for interoperability of multiple independent hypervisors.
+Nested virtualization addresses the (vertical) security challenge by placing an additional hardware-protected security monitor beneath the untrusted virtualization layer [Turtles, CloudVisor] -- requiring a sufficiently open IaaS.
+It also features (horizontal) flexibility and interoperability benefits by defining a blanket layer on top of existing provider hypervisors in a hybrid cloud setting, enabler for self-service security for clouds of clouds [XenBlanket].
+However, such horizontal and vertical benefits have not yet been combined in a unified security architecture.
+Moreover, the additional layer separation might also make user vs.
+provider trade-offs more coarse-grained than the previous approach.
+%Hence, in SUPERCLOUD, we propose to define a new IaaS security architecture for distributed clouds that may combine the benefits of the two previous architectural approaches in terms of both layered design and modularity to achieve self-service cloud security with tuneable, fine-grained trade-offs with provider-centric security to protect customer VMs.
+The benefits of this hybrid approach will be to lift both barriers simultaneously, but also to be applicable in a large variety of underlying provider infrastructures.
+For instance, if the underlying hypervisor is not open, leveraging the benefits of nested virtualization for hypervisor interoperability; and otherwise, using either approach, depending on desired security vs.
+performance trade-offs, and compatibility with legacy systems or deployment of new security hardened IaaS architectures.
+%
+%
 
@@
 
-In previous sections we provided a global analysis of modern challenges of cloud computing  and how those could be globally addressed addressed leveraging nested virtualization and different hypervisor architectures.
+\noindent This paper reviewed the main DCI design challenges and how they could be addressed by combining NV with different hypervisor architectures.
+We also presented a preliminary system architecture combining GPHs and MHs for 2D security management of DCIs.
 
-Future works will be focused on two complementary parts.
-Firstly, they will concern the completion of the architecture of security in order to concretely meet horizontal design principles, identifying key components, similarly as done for the vertical properties.
-At the same time, a major effort will be directed on enlarging the initial implementation.
+Ongoing work focuses on refining this architecture to add self-protection features based on the VESPA framework.
+The objective is to determine where to place detection, decision, and reaction components for the different autonomic security loops, both within layers and across clouds.
+We are also implementing a proof-of-concept prototype of this security architecture using Xen as L1 GPH and NOVA as L0 MH.
+Xen seemed the most suitable hypervisor for L1, notably due to its support of different virtualization techniques and ability to run without HAV support.
+NOVA appears promising for L0 due to its clean design and code availability.
+With this platform, we are currently focusing on implementing vertical properties (e.g., cross-layer self-protection) in a single cloud setting, with extension to horizontal properties (e.g., inter-cloud self-protection) in a distributed cloud setting.
+
+%On the other hand, L1 choice is much simpler. 
+%
+%, focusing first on vertical properties in a single cloud setting, and extending it to horizontal properties in a distributed setting.
+%
+%
+%\begin{itemize}[]
+%\item \textbf{Inter-platform support} Executed as L1, Xen would virtualize its nested guest initially using PV delivering the compatibility layer.
+%\item \textbf{Tiny TCB} The NOVA kernel will represent the only part of the platform executed at the highest level of privilege.
+Even the VMM, Vancouver, will be executed in the L0 user-space.
+The micro-kernel in L0 enforces a strictly modular approach to L0.
+Therefore, the VMM will present distinct processes instances for each L1 guest, enforcing a context division.
+%\item \textbf{3rd part compatibility} From a user point of view the architecture still exports the same interfaces provided by the Xen hypervisor, allowing the user to keep its applications unchanged.
+%\item \textbf{Fail-safe components architecture} Moreover, the tiny core is the only safety-critical software component.
+The system could potentially be able to recover any crash in different user-space components.
+%\end{itemize}
+%
+ %
+%
+%After this initial study, the first step will be deploying a minimal nested architecture testing ``out-of-the-box'' features and limitations.
+The following part will concern the conception of patch and extensions in order to adapt the general design of MH to properly act as L0.
+%The final goal is to provide an architecture able to leverage nested virtualization cloud platforms improvements, having a smaller TCB and allowing customers to leverage flexible migrations and scheduling.
+
+%
+% Our vision is that the security of the SUPERCLOUD should be both self-managed and 360.
+
+% Our ambition is to define a security supervision architecture and framework for the SUPERCLOUD enabling this 360 view of security management with a high level of automation.
+Thus, the SUPERCLOUD will bring unified, simple management of security for cloud of clouds.
+
+%In the previous sections we proposed a functional analysis of cloud architecture in order to respect design principles that we identified in the introduction.
+In this section, we propose a GPH over MH implementation, based on .
+%
+%\begin{itemize}[]
+%\item \textbf{Inter-platform support} Executed as L1, Xen would virtualize its nested guest initially using PV delivering the compatibility layer.
+%\item \textbf{Tiny TCB} The NOVA kernel will represent the only part of the platform executed at the highest level of privilege.
+Even the VMM, Vancouver, will be executed in the L0 user-space.
+The micro-kernel in L0 enforces a strictly modular approach to L0.
+Therefore, the VMM will present distinct processes instances for each L1 guest, enforcing a context division.
+%\item \textbf{3rd part compatibility} From a user point of view the architecture still exports the same interfaces provided by the Xen hypervisor, allowing the user to keep its applications unchanged.
+%\item \textbf{Fail-safe components architecture} Moreover, the tiny core is the only safety-critical software component.
+The system could potentially be able to recover any crash in different user-space components.
+%\end{itemize}
 

### 2014-03-29 paper/draft.tex
commit

--- 2014-03-29 paper/draft.tex
@@
 % Intro + performance issues
-Several works presented in literature integrates these a subset of this design principles in their architecture, leveraging nested virtualization \cite{turtle:ibm,art:blan, cloudvisor:zhang}
-Cloud computing feature set could be enlarged exploiting two layers of virtualization instead of just one.
-The presence of two different hypervisors allows the possibility to specialize hypervisors functions.
-Instead of having a general purpose hypervisor we could split the hypervisor role in two different layers, interconnected with a well defined interface.
-Each layer will have to care about a subset of features, simplifying overall development.
-However, nested virtualization on x86 architecture faced impeding performance issues \cite{rec:virt}.
+\noindent A growing body of research has investigated building architectures and infrastructures integrating a subset of those design principles leveraging nested virtualization (NV)~\cite{turtle:ibm,art:blan, cloudvisor:zhang, Inception, hvx:usenix}.
+The distributed cloud feature set may be expanded using two levels of virtualization instead of a single one.
+This design allows to specialize hypervisor functions per layer, each layer addressing a subset of features, simplifying overall development compared to a general-purpose hypervisor.
+However, NV development on x86 architecture was initially hindered by severe performance issues~\cite{rec:virt}.
+% Perfomance issuess adressed
+However, several works have shown it possible to realize low-overhead NV using dedicated implementation architecture~\cite{turtle:ibm}, or to recover performance loss exploiting multi-level consolidation~\cite{art:blan}.
+NV slowdown can now be considered an acceptable price to achieve advanced features -- hardware manufacturers are also developing new hardware primitives to reduce this overhead~\cite{vmcs:nakajima}.
 
-% Perfomance issuess adressed
-However, several works demonstrate that it is possible to put in place low overhead nested virtualization using special approaches \cite{turtle:ibm} .
-Moreover, other works try to recover performance loss exploiting multi-level consolidation, introduced in nested virtualization \cite{art:blan}.
-In conclusion, nested virtualization slowdown could be considered nowadays an acceptable price to achieve advanced features and, in addition, hardware manifacturers are developing and proposing new hardware facilities to reduce this overhead \cite{vmcs:nakajima}.
+However, so far no architecture has currently addresses both vertical and horizontal set of challenges simultaneously.
+XenBlanket~\cite{art:blan} does not investigate cloud architectural security concerns (DP2, DP3), limiting its focus to providing new features on a coordinated multi-provider cloud.
+Cloudvisor~\cite{cloudvisor:zhang} proposed a NV-based security architecture, but cannot run more than one nested hypervisor at time, which is a criticial requirement for multi-provider cloud platform (DP4).
+In the super-cloud vision~\cite{xclo:blank}, each user runs its own L1 hypervisor instance to realize  a blanket layer overcoming incompatibilities between different providers.
+However, security issues have not been addressed.
 
-To sum up, no architecture had been presented to address vertical and horizontal class of problems at the same time until now.
-XenBlanket does not investigate cloud architectural security concerns (DP3,DP2), limiting the focus on new potential feature of a coordinated multi-provider cloud.
-Cloudvisor proposed a security architecture that leverages nested virtualization without allowing the possibility of executing more than one nested hypervisor at time, that is a crucial requirements of a multi-provider cloud platform (DP4).
-In a SuperCloud scenario, each user has to run its own L1 hypervisor instance to enforce a compatibility layer, overcoming incompatibilities between different providers.
-
-Other approaches leveraging only one layer of virtualization does not achieve the whole design-principle satisfaction.
-Self-service cloud approach (SSC) \cite{ssc:art} design a complete security architecture to overcome major security and privacy issues of currently deployed cloud infrastructures, in order to alleviate the different needing user-centric and provider-centric vision of cloud computing presented in \ref{sec:intro}.
-Even if this architecture reproduces an approach similar to the SuperCloud without leveraging an extra layer of virtualization (DP4), the approach changes the behavior where the user acts introducing new components that have to be directly managed by the user.
-Therefore, the main drawback of this solution is to change the interface with the user, preventing the possibility to straightforwardly move to the new platform without having to change its applications (DP5).
+Other approaches based on a single layer of virtualization do not fulfill all principles either.
+Self-Service Clouds (SSC)~\cite{ssc:art} introduces a complete security architecture to overcome major security and privacy issues of currently deployed cloud infrastructures.
+The aim is to reconcile user-centric and provider-centric requirements of the cloud.
+Despite being similar to the super-cloud approach (DP4), SSCs introduces new components that have to be directly managed by the user.
+This change of interface with user VMs prevents moving straightforwardly to the new platform without having to change its  applications (DP5).
 
 % Integration NDSS Taming
-Furthermore, DeHype \cite{wu2013TamHosHypMosDepExe} presented a dissected version of KVM hypervisor, trying to reduce the TCB of this popular hypervisor.
-The approach consisted in demoting a great part of the hypervisor codebase to a user-level execution.
-DeHype satisfies all vertical design principles listed before (DS1,DS2,DS3), reinterpreting the microkernel approach defining  the greatest part of code to userspace.
-DeHype increases a lot the robustness of a cloud platform leveraging KVM, limiting for example the possibility of attack in multi-tenancy platform, but it does not address horizontal design principles about a multi-provider service.
+To reduce the TCB size, DeHype~\cite{wu2013TamHosHypMosDepExe} presented a dissected version of KVM, moving a great part of the code to user space.
+DeHype satisfies all vertical design principles, re-interpreting the microkernel approach in a KVM setting.
+It increases a lot the robustness and security of a cloud platform leveraging KVM (DP1), but does not address horizontal challenges for a multi-provider service.
 
@@
 
-\noindent To solve the TCB size issue, several works have introduced the idea of \textit{micro-hypervisor (MH)} architecture~\cite{nova, NoHype, xmhf}.
-To reduce the hypervisor size, such new hypervisor architectures draw inspiration from evolution in OS architecture (e.g., extensible, micro-, exo-, component-based kernels).
+\noindent To solve the TCB size issue, the idea of \textit{micro-hypervisor (MH)} architecture was introduced~\cite{nova, NoHype, xmhf, mccune2010TruEffTCBRedAtt}.
+Such hypervisor design draws inspiration from evolution in OS architecture (e.g., extensible, micro-, exo-, component-based kernels).
 Within the hypervisor are distinguished security-critical modules from non-sensitive code.
-The aim is to expel as much code as possible from the TCB and target ultra-thin hypervisors.
-For instance, some components of the hypervisor (e.g., device drivers, VM address space management) may be virtualized outside the micro-hypervisor, isolated, and restarted in case of compromise~\cite{nova}.
-The extreme case is to remove the virtualization layer altogether~\cite{NoHype}.
+The aim is to expel as much code as possible from the TCB make the hypervisor ultra-thin.
+Some hypervisor components (e.g., device drivers, VM address space management) may be virtualized outside the MH, isolated, and restarted in case of compromise~\cite{nova}.
+The utmost MH removes the virtualization layer altogether~\cite{NoHype}.
 
-with a traditional microkernel-like design that tries to consistently reduce the amount of code running with the highest privilege level.
-
-A micro-kernel traditionally consists in a minimal kernel running in privileged mode, offering only a few fundamental services to other modules (e.g, scheduling, IPC, memory management) running in a less privileged domain.
+%MH = MK+ H.
+MH architectures may be seen as convergence of traditional hypervisor and micro-kernels (MK).
+%1) MK
+%Perspective: minimalisme: 
+%with a traditional microkernel-like design that tries to consistently reduce the amount of code running with the highest privilege level.
+From the MK design principles legacy, the MH is guided by TCB minimization. 
+%-- Définition du mk
+%A micro-kernel traditionally consists in a minimal kernel running in privileged mode, offering only a few fundamental services to other modules (e.g, scheduling, IPC, memory management) running in a less privileged domain.
 All other components, such as device drivers, are moved out in user-space.
-While the benefits of such a design has been questioned due to the cost of IPCs in first generations of micro-kernels, such arguments have since then been completely dispelled~\cite{liedtke1993ImpIPCbyKerDes,liedtke1995micConnova}.
+%-- Définition du mh:
+%Therefore, leveraging this concept of microkernels, a micro hypervisor (MH) is composed from a small kernel providing only essential functions, like scheduling, IPC, memory management and VMX extension.
+Thus, it features a minimal kernel running in privileged mode, offering only a few fundamental services (e.g, scheduling, IPCs, memory management, VMX extensions) to other modules running in less privileged domains.
+All other components, such as device drivers, are moved out in user-space.
+% Skip inconvénients du MK
+%1) Inconvénients du mk: performance.
 %Microkernels had been questioned because of the wasteful performance of first generations, due to IPC intrinsic cost.
 This argument had been completed confuted by Liedtke in 1993: \cite{liedtke1993ImpIPCbyKerDes,liedtke1995micCon}.
+%While the benefits of such a design has been questioned due to the cost of IPCs in first generations of micro-kernels, such arguments have since then been completely dispelled~\cite{liedtke1993ImpIPCbyKerDes,liedtke1995micConnova}.
+%-- Bénéfices du mk
+The main benefit of such increased modularity is
+% a) Bénéfices du mk: flexibility in privilege
+% One of the benefits of micro-kernels is to enable to easily redesign a hierarchy of privileges. 
+%
+%The different groups may hold correspond to different trust relationships of principals running Control is finer than in the hypervisor, both by the granularity of isolation (lightweight threads), and by a unique control point of RT behavior in the scheduler.
+Device drivers are run in , allowing great flexibility for device sharing patterns.
+to easily define groups of user-space services, each group sharing a common a set of privileges.
+Trust relationships between groups may be arbitrarily defined, mirroring those between sets of principals or applications sharing the system.
+%For instance, a user application normally trusts the micro-kernel and only the set of user-space services it requires.
+However, such drivers and libraries may not be trusted by every application run on the system.
+User applications share the system but may need to support completely different set of system services.
+% -- Les personalités:
+% In addition, MH inherits the concept of micro-kernel to implement different ``personalities''. 
+This allows MK to implement different ``personalities'': device drivers being run in user mode, such fine-grained control allows great flexibility to implement device sharing patterns.
+%
+%2) H
+%
+%Plus facile dans un contexte d'hyperviseurs:
+%However, the interest of such ``personalities'' has been largely reduced by the fact that system services, device drivers and even applications have to be completely re-designed to fit this new architectural pattern.
+%
+%Therefore, with the introduction of VMs, microkernels lost the battle in favor of hypervisors, in terms of popularity.
+While the interest of realizing such ``personalities'' in traditional MKs somewhat lessened as system services, drivers, and even applications had to be completely re-designed (hypervisors offering better primitives for abstraction with the notion of VM),
+%
+%
+%3) MH.
+%
+such flexibility still remains a major benefit of coupling micro-kernel design with virtualization~\cite{OKL4}.
 
+%1) Pas de tiny TCB
+%However, the great advantage of tiny TCB, provided by microkernel, was not provided also by hypervisors. 
+%2) Architecture monilithic
+%As we analyzed before, they present normally a monolithic architecture, like the great part of modern OS. 
+%This feature kept alive the interest in microkernels, in particular trying to couple microkernel benefits with virtualization
+
+%-- Exemples: Several micro-hypervisor have been proposed in literature \cite{xmhf, nova}. 
+%
+%=> -- MH = MK+ virtualization. They rely on the micro-kernel paradigm, extending it to virtualization. MV
+
+%-- Relation entre MH et principes: 
+%* MH naturally satisfy vertical design principles (DP1,DP3, DP4). 
+MHs are particularly well adapted to sastisfy vertical design principles.
+% Security:
+%Pourquoi: 
+For instance, MH design increases reliability as services crashing in user space may be restarted without bringing down the entire system (DP2).
+%Fail-safe principle:
+%Moving Finally, this extraction process drives to a more solid platform, considering that user-space crashes are generally not fatal for the entire system integrity. 
+The attack surface is also lowered by the small TCB size, the minimal kernel only containing security-critical components and forming a last line of defense (DP3).
+%The tiny core could be considered the only system-critical componente of the architecture and, due to its relative small dimension and the consequent minor attitude to vulnerabilities,it could represent the ``last line of defense''.
+MH might also address some horizontal principles. 
+%Interop principle:
+%Qu'est ce que c'est:
+``Personalities'' are concretely a set of processes handling specific functions running in user-space.
+Implementing such features outside the kernel makes it easier for each user to rely on different instances of such personalities running concurrently, possibly in a distributed manner (DP4).
+%This process separation between users could enforce a straightforward mechanism to neutralize VM-to-VM attack from a malicious user.
+Moreover, user-space processes are easier to debug with commodity tools (GDB, valgrind).
+
+% Enlever les exemples qui ne sont pas significatifs.
+% A possible example, that will be analyzed in next section, is the user-space VMM component of Nova, Vancouver, that resides in user-space \cite{nova}.
+% In addition, this approach is not far from the one proposed in DeHype \cite{wu2013TamHosHypMosDepExe}, where a consistent part of the KVM hypervisor has been demoted in user-space.
+
+% Texte additionnel sur les MH:
 %Architectural solutions were also created to make the hypervisor more minimal and flexible, targetting its two most vulnerable elements: (1) the core hypervisor itself; and (2) the management VM (e.g., Xen Dom0).
@@
 They are thus hard to apply to mainstream hypervisors today.
-One of the benefits of micro-kernels is to enable to easily redesign a hierarchy of privileges.
-A user application normally trusts the micro-kernel and only the set of user-space services it requires.
-However, such drivers and libraries may not be trusted by every application run on the system.
-User applications share the system but may need to support completely different set of system services.
-This allows micro-kernels to implement different ``personalities''.
-However, the interest of such ``personalities'' has been largely reduced by the fact that system services, device drivers and even applications have to be completely re-designed to fit this new architectural pattern.
-Therefore, with the introduction of VMs, microkernels lost the battle in favor of hypervisors, in terms of popularity.
-However, the great advantage of tiny TCB, provided by microkernel, was not provided also by hypervisors.
-As we analyzed before, they present normally a monolithic architecture, like the great part of modern OS.
-This feature kept alive the interest in microkernels, in particular trying to couple microkernel benefits with virtualization\cite{Heiser}.
 
-Therefore, leveraging this concept of microkernels, a micro hypervisor (MH) is composed from a small kernel providing only essential functions, like scheduling, IPC, memory management and VMX extension.
-Several micro-hypervisor have been proposed in literature \cite{xmhf, nova}.
-They rely on the micro-kernel paradigm, extending it to virtualization.
-MH naturally satisfy vertical design principles (DP1,DP3, DP4).
-The tiny core could be considered the only system-critical componente of the architecture and, due to its relative small dimension and the consequent minor attitude to vulnerabilities,it could represent the ``last line of defense''.
-
-In addition, MH inherits the concept of micro-kernel to implement different ``personalities''.
-Those ``personalities'' are concretely a set of ``processes'' handling specific functions running in user-space.
-As those functions are implemented outside the kernel, each user could rely on a different instances of the hypervisor user-space process, that could coexist at the same time.
-This process separation between user could enforce a straightforward mechanism to neutralize VM-to-VM attack from a malicious user.
-Moreover, user-space processes are easier to debug with commodity tools (GDB, valgrind).
-Finally, this extraction process drives to a more solid platform, considering that user-space crashes are generally not fatal for the entire system integrity.
-A possible example, that will be analyzed in next section, is the user-space VMM component of Nova, Vancouver, that resides in user-space \cite{nova}.
-In addition, this approach is not far from the one proposed in DeHype \cite{wu2013TamHosHypMosDepExe}, where a consistent part of the KVM hypervisor has been demoted in user-space.
+% Texte sur les MK, H, et MV
+%Embedded virtualization brings benefits to achieve several such properties, for instance abstraction and isolation by multiplexing and isolating VMs.
+However, current hypervisor architectures show strong limitations for many others.
+A main flaw is the huge TCB size, leading to reliability, security, and administration issues.
+Flexibility is rather poor, as for device sharing.
+RT guarantees are also difficult due to two-level system scheduling, i.e., threads at VM-level, and VMs at hypervisor-level.
+Finally, resource control remains coarse-grained through heavyweight VMs.
+%
+%\textit{Micro-kernel} design has thus been increasingly popular to meet such challenges.
+Reducing the kernel to its fundamental elements, and making the rest of the OS deprivileged outside results in an extremely minimal TCB that may be formally verified.
+Strong isolation and high-perfromance communication between address spaces is also achieved through very efficient IPCs.
+Micro-kernels increasingly support virtualization: privileged instructions are intercepted and emulated in a dedicated system service.
+Control is finer than in the hypervisor, both by the granularity of isolation (lightweight threads), and by a unique control point of RT behavior in the scheduler.
+Device drivers are run in user space above the micro-kernel, allowing great flexibility for device sharing patterns.
+%
+%Such architectures are very similar, although the hypervisor is driven by abstraction, to match closely virtual to real resources, while micro-kernels are guided by TCB minimization.
+%They tend to converge: the \textit{micro-visor} allows to run both VMs and threads over a lightweight execution environment~\cite{OKL4}.
+Like micro-kernels, it has a minimal TCB, and high-performance asynchronous IPCs between VMs, applications, or user-mode drivers.
+Like hypervisors, it allows virtualization of CPU, memory, or I/O through a hardware abstraction layer (HAL).
+Arbitrary control over resource partitioning in security domains or RT scheduling of threads or VMs is also possible.
+Finally, the micro-visor offers great architectural flexibility: drivers may be virtualized in a VM, or executed as system services in their own address space.
 
@@
 
-\noindent Despite the positives, MH approaches had demonstrated some limitations.
-Firstly, the straightforward one is the initial lack of hardware support, lack of device drivers are still a relevant Achilles' heel for this solutions, inherited from its ancestors microkernels.
-Fortunately, this kind of issue could be addressed leveraging I/O passthrough techniques, although it will cause exclusive allocation issues.
-
-Moreover, as reported in previous sections, several cloud computing issues are simply not addressable with general purpose hypervisor (GPH).
-However, they are commonly widely adopted and represent the bedrock of existing platform.
-Therefore, GPH exports wide and solid interfaces leveraged by a plenty of different software that is able to properly communicate and act through them on the virtualization layer.
-On the contrary, MH lack of cloud toolkit integration, requiring an explicit support to be written, not satysfing 3rd part compatibility requirements (DP5).
-GPH are tightly connected with cloud toolkit and could present solid interfaces to communicate with different tools.
-
-Finally, microkernel modularization introduce a major challenge in architecture design.
-Due the increased complexity, for sake of kernel size, MH could generally lack of several virtualization features, commonly available on modern GPH hypervisor.
+\noindent The MH approach also showed several limitations.
+First, support for hardware is lacking, notably in terms of device drivers.
+Fortunately, this issue may be addressed leveraging I/O passthrough techniques, despite some resulting exclusive allocation challenges.
+Second, several cloud issues are simply not addressable with GPHs.
+However, GPHs are now widely adopted and foundation of existing cloud platforms.
+While a GPH exports mature APIs to upper software layers and is also tightly integrated with cloud toolkits, MHs lack such API stability and integration, requiring explicit dedicated code to be written -- thus violating DP5.
+Third, for the sake of kernel size, MHs could lack several virtualization features commonly available on modern GPH hypervisors.
 For instance, XMHF does not support multiple guests.
-This kind of drawbacks could limits the potential interest in MH, even they are able to introduce several important features.
+Such limitations applicable for MH architectures alone may be lifted partially or totally when applying MH design to NV architectures.
 
@@
 
-\section{Security Architecture}
+\section{Towards 2D Security Management}
 \label{ref:arch}
+
+\indent We now sketch how the previous enhanced-NV design could enable to build a security architecture for a distributed cloud infrastructure  (DCI) satisfying both vertical and horizontal requirements.
+Such DCIs will face 2D threats, vertical spanning layers, and horizontal, spanning domains DCIs being multi-layer, multi-domain distributed cloud infrastructures.
+Vertically, attacks currently target several infrastructure layers (e.g., VMs, hypervisor).
+Unfortunately, most existing defenses are generally for single layers only.
+This makes it difficult to grasp the overall extent of an attack.
+Horizontally, threats may propagate through the network from cloud to cloud.
+Unfortunately, most existing defenses are for single clouds only.
+Lack of interoperability in security policies and mechanisms is thus a major barrier to  unified hybrid cloud security.
+DCI security thus requires 2D protection.
+
+We consider the security architecture shown in Figure~\ref{fig:archisecu}.
+
+Supercloud
+SUPERCLOUD Layer built SS cloud
+Also security management both vertically and horizontally.
+
+OS:
+
+Provider MH: Assumption min V, XenDisagregated...
+Above GPH: 
+Achieve Vertical security Cross Layer security
+
+GPH allows also to layer of interoperability horizontal blanket layer above the different MH to enalbe self-service security.
+
+On the top are the VM built self-service security where security is personalized/customized.
+Automation.
+Cross-Layer Framewokr: VESPA
+Cross-Domain Framework:
+360 Atuomation
+
+ This means: providing vertically a cross-layer integrated vision of security of the cloud and SUPERCLOUD infrastructure; defining horizontally a security abstraction layer to achieve a homogeneous level of security across clouds independently from underlying provider mechanisms; and integrating both dimensions smoothly.
+This approach is key to providing a unified view of security of the SUPERCLOUD.
+Our ambition is to define a security supervision architecture and framework for the SUPERCLOUD enabling this 360 view of security management with a high level of automation.
+Thus, the SUPERCLOUD will bring unified, simple management of security for cloud of clouds.
+
+ Our vision is that the security of the SUPERCLOUD should be both self-managed and 360.
+
+The second challenge means that although many defense mechanisms may be leveraged in this 2D landscape, they remain difficult to configure and to reconcile.
+The current “by hand” approach becomes a nightmare for security administrators.
+Despite some automated solutions for managing security of single clouds, there is currently no solution truly applicable to federations of clouds in practice.
+
+two major security challenges:
+
+In terms of architecture:
+The northbound, user-centric super-cloud API will implement self service security and dependability, allowing to build self-service clouds (SSC) where VM, network, and storage security may be personalized independently from the underlying provider.
+Such clouds may be defined at the service level to the full IaaS, relationships between SSC being flat or nested, also with complex trust relationships between sets of SSC according to the ecosystem relationships of the super-cloud providers providing the SSC.
+The southbound, provider-centric super-cloud API will define a unified security and dependenability management plane of the compute, data, and network resources independently from providers.
+This may apply for B2B use cases to management of clouds of clouds but also to device-to-cloud scenarios if devices are considered, with possible extensions to the personal cloud for device-to-device scenarios, the SSC taking the form of personal cloud-based digital infospheres where security and dependability may be totally user-controlled.
+...
+The SUPERCLOUD will fulfill that vision with a two-level infrastructure:
+•	The lower part of the SUPERCLOUD is be a distributed infrastructure will realize the distribute resource abstraction layer federating distributed cloud resources and implementing the self-service security to build SSC.
+It will implement the northbound, user-centric super-cloud API.
+•	Above, the upper part of the SUPERCLOUD will be a supervision infrastructure providing a set of security and resilience services to guarantee the security of the overall SUPERCLOUD architecture.
+It will realizing self-service security, end-to-end security, and resilience.
+It will implement the southbound, provider-centric super-cloud API.
+
+and security management complexity. Th
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=10cm]{DCCArchitecture.png}
+\end{center}
+\caption{Security Architecture.}
+\label{fig:archisecu}
+\end{figure}
+
+A DC 
+%Our vision is that the security of the SUPERCLOUD should be both self-managed and 360°.
+Security will be autonomically managed through orchestration of a number of self-protection loops governing interactions between detection, decision, and reaction components.
+This approach enables simpler, stronger, and more efficient security management to detect and respond to threats [CPW03].
+Security monitoring will also be 360°.
+This means: providing vertically a cross-layer integrated vision of security of the cloud and SUPERCLOUD infrastructure; defining horizontally a security abstraction layer to achieve a homogeneous level of security across clouds independently from underlying provider mechanisms; and integrating both dimensions smoothly.
+This approach is key to providing a unified view of security of the SUPERCLOUD.
+%Our ambition is to define a security supervision architecture and framework for the SUPERCLOUD enabling this 360° view of security management with a high level of automation.
+Thus, the SUPERCLOUD will bring unified, simple management of security for cloud of clouds.
+
+Autonomous
+
+Supercloud architecture
+
+Proposition of Architecture
+
+Base Layer
+
+Supercloud layer
+
+VESPA.
 
@@
 \end{itemize}
+
+%The SUPERCLOUD project aims to define the architecture and build the infrastructure to fulfill the vision of user-centric secure and dependable management of a clouds of clouds.
+Our approach will be to define a new distributed architectural plane, the super-cloud providing an end-to-end interface both between user-centric and provider-centric views of multiple clouds.
+Its role will be both to provide a distributed resource abstraction and flexible but unified control for management of security and resilience.
+%From the user perspective, the super-cloud will enable user-centric security and dependability through a unified distribution layer for cloud resources, independently from their type (computation, storage, and network) and from underlying cloud providers.
+It thus forms an intermediation layer for resources between their production by providers and their consumption by users.
+This in line with the vision introduced by IBM [X] as well as with the utility computing paradigm [X], but will go well beyond, leveraging this distribution layer to explore high customizability of user cloud security and dependability according to well-defined SLAs.
+%From the provider perspective, the super-cloud will also enable provider-independent control over security and resilience over the overall distributed cloud infrastructure, security being autonomous managed and guaranteed end-to-end.
+This will also make the resource distribution layer both secure and resilient.
+%The level of abstraction, from system to middleware interoperability of the super-cloud be also be tuned to allow trade-offs between user-centric and provider-centric control for security services. 
+%In terms of architecture:
+%The northbound, user-centric super-cloud API will implement self service security and dependability, allowing to build self-service clouds (SSC) where VM, network, and storage security may be personalized independently from the underlying provider.
+Such clouds may be defined at the service level to the full IaaS, relationships between SSC being flat or nested, also with complex trust relationships between sets of SSC according to the ecosystem relationships of the super-cloud providers providing the SSC.
+%The southbound, provider-centric super-cloud API will define a unified security and dependenability management plane of the compute, data, and network resources independently from providers.
+This may apply for B2B use cases to management of clouds of clouds but also to device-to-cloud scenarios if devices are considered, with possible extensions to the personal cloud for device-to-device scenarios, the SSC taking the form of personal cloud-based digital infospheres where security and dependability may be totally user-controlled.
+%...
+%The SUPERCLOUD will fulfill that vision with a two-level infrastructure:
+%•	The lower part of the SUPERCLOUD is be a distributed infrastructure will realize the distribute resource abstraction layer federating distributed cloud resources and implementing the self-service security to build SSC.
+It will implement the northbound, user-centric super-cloud API.
+%•	Above, the upper part of the SUPERCLOUD will be a supervision infrastructure providing a set of security and resilience services to guarantee the security of the overall SUPERCLOUD architecture.
+It will realizing self-service security, end-to-end security, and resilience.
+It will implement the southbound, provider-centric super-cloud API.
+%In what follows, we describe the SUPECLOUD vision for each self-service security, self-managed security, end-to-end security and resilience properties. 
+%
+%For computing workloads, user-centric cloud security management faces two challenges.
+%•	Lack of IaaS infrastructure openness: provider-controlled immutable hypervisors (or that mask access to the hardware) prevent users from taking advantage of hypervisor-level advanced resource management and security services such as high availability or VM introspection for fine-grained VM misbehavior detection and reaction.
+Such examples of vendor lock-in are already a notorious issue for single provider infrastructures [CSA Guidance].
+They are a real barrier for hybrid clouds: the varying degree of openness of the underlying IaaS forces to achieve security interoperability in upper software layers instead of at system level as in multi-IaaS orchestration platforms, highly reducing user control over security management.
+Such limitations also prevent fine-grained trade-offs between provider and user control over security : among security services, users cannot choose between those to customize, and those to delegate to the provider infrastructure.
+%•	Security threats for customer VMs: large, complex, overprivileged hypervisors are a serious risk for VMs [CSATT].
+The assumption of a trusted hypervisor is no more, illustrated through many examples of VM escapes, breaking hypervisor isolation through exploitaiton of device drivers vulnerabilities [Virtunoid, CloudBurst].
+A multi-layer approach to IaaS security is needed considering lower layers, to guarantee VM security even if the hypervisor layer is compromised.
+This challenge is intrisically linked with openness and flexibility issues: from a user perspective, to protect their VMs from an untrusted provider infrastructure, customers may wish adding their own security modules (or layer) in the infrastructure, provided it is open enough.
+Conversely, a provider may also wish to benefit from an open hypervisor architecture to harden its IaaS against rogue VM behaviors.
+%Earlier IaaS architectures such as disaggregated hypervisors or nested virtualization provided partial answers to either of those issues but not both.
+Hypervisor disaggregation (for the Dom0 or the core hypervisor) achieves flexibility through the definition of clearly separated access control domains between provider-centric and user-centric security management planes where fine-grained trade-offs may be reached [SSC].
+This approach is compatible with legacy hypervisors (Dom0 case) or requires new architectures (core hypervisor case).
+However, it only addresses security and flexibility challenges in a single cloud context, but not for interoperability of multiple independent hypervisors.
+Nested virtualization addresses the (vertical) security challenge by placing an additional hardware-protected security monitor beneath the untrusted virtualization layer [Turtles, CloudVisor] -- requiring a sufficiently open IaaS.
+It also features (horizontal) flexibility and interoperability benefits by defining a blanket layer on top of existing provider hypervisors in a hybrid cloud setting, enabler for self-service security for clouds of clouds [XenBlanket].
+However, such horizontal and vertical benefits have not yet been combined in a unified security architecture.
+Moreover, the additional layer separation might also make user vs.
+provider trade-offs more coarse-grained than the previous approach.
+%Hence, in SUPERCLOUD, we propose to define a new IaaS security architecture for distributed clouds that may combine the benefits of the two previous architectural approaches in terms of both layered design and modularity to achieve self-service cloud security with tuneable, fine-grained trade-offs with provider-centric security to protect customer VMs.
+The benefits of this hybrid approach will be to lift both barriers simultaneously, but also to be applicable in a large variety of underlying provider infrastructures.
+For instance, if the underlying hypervisor is not open, leveraging the benefits of nested virtualization for hypervisor interoperability; and otherwise, using either approach, depending on desired security vs.
+performance trade-offs, and compatibility with legacy systems or deployment of new security hardened IaaS architectures.
+%
+%
 

### 2014-03-30 paper/draft.tex
Update cite uK
Qques corrections

--- 2014-03-29 paper/draft.tex
@@
 %* MH naturally satisfy vertical design principles (DP1,DP3, DP4). 
-MHs are particularly well adapted to sastisfy vertical design principles.
+MHs are particularly well adapted to satisfy vertical design principles.
 % Security:
@@
 
-However, it fails to safisfy vertical design principles as TCB size is that of currently deployed NV architectures.
+However, it fails to satisfy vertical design principles as TCB size is that of currently deployed NV architectures.
 The interest of other MH-based architectures will thus be to introduce isolation and privacy improvements by reducing the TCB, while still keeping horizontal features.
@@
 This requirement is implemented by GPHs like Xen and KVM, but usually not included in the key features to be provided by a MH: implementing nested VMX may be estimated to around 1KLoC, to be compared with the size of a typical MH (2-20 KLoC)~\cite{xmhf}.
-More generally, the MH quest towards ever increasing minimalism could represent a serious limitation to preserving functions of existing platforms: the constraint of privileged code size may force MH developpers to drop some features considered as non-essential.
+More generally, the MH quest towards ever increasing minimalism could represent a serious limitation to preserving functions of existing platforms: the constraint of privileged code size may force MH developers to drop some features considered as non-essential.
 This may apply to NV-related primitives, but also to other types of features.
@@
 Its northbound interface federates cloud resources in a provider-independent manner and enables to customize their security to realize the SSCs.
-While many orchestration architectures have been proposed at service or middleware levels (e.g., multi-Iaas platforms), our architecture aims to achieve federation at system-level to fully take advantage of underlying provider platform openness (if available) for finer-grained control and increased flexibility.
+While many orchestration architectures have been proposed at service or middleware levels (e.g., multi-IaaS platforms), our architecture aims to achieve federation at system-level to fully take advantage of underlying provider platform openness (if available) for finer-grained control and increased flexibility.
 % Fonction South
